### 2022-09-19 paper/paper.tex
paper

\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}

\title{Semilattice Graphs for Distributed Databases}
\author{Jessica Taylor}
\date{\today}

\begin{document}
    \maketitle

    \section{Introduction}
        A \emph{semilattice} is a set $L$ with a binary operation $\vee$ (``join'') that is commutative ($x \vee y = y \vee x$), associative ($(x \vee y) \vee z = x \vee (y \vee z)$), and idempotent ($x \vee x = x$).

        In databases, lattices are useful for defining a notion of a most updated value. For example, with $L$ as a set of sets of messages and $\vee = \cup$, the lattice represents sets of messages that are updated to include more messages as they come in.

    \section{Semilattice Graphs}

        A common problem in networked computing is dependency managament. A software package may depend on a set of other packages, each of which depends on another set of packages. We consider extending semilattices to semilattice graphs to handle dependency management.

        A \emph{semilattice graph} consists of:

        \begin{itemize}
            \item A set of keys $K$.
            \item A set of values $V$.
            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a set of keys, its dependencies.
            \item A join function $j : (\Sigma (k: K, x : V, y : V), d(k, x) \cup d(k, y) \rightarrow V) \rightarrow V$ that joins two values given their dependencies. (The sigma set $\Sigma (a: A), B(a)$ denotes a set of pairs $(a, b)$ where $a \in A$ and $b \in B(a)$.)
        \end{itemize}

        And satisfies the following properties:

        \begin{itemize}
            \item $d$ must be well-founded; that is, there are no infinite sequences $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, \ldots$
            \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
            \item $j$ must not grow dependencies; that is, for any $k : K$, $g: K \rightarrow V$, $x, y \in V$, we must have $d(k, x \vee_{k, g} y) \subseteq d(k, x) \cup d(k, y)$.
        \end{itemize}

        A semilattice graph \emph{has defaults} if there is a function $z : K \rightarrow V$ such that $d(k, z(k)) = \{\}$ for each $k \in K$.

    \section{Joins on semilattice graphs}

        It is now possible to define a semilattice over a semilattice graph. This lattice's items are functions $K \rightarrow V$ (called \emph{mappings}) and the join operation is defined as follows:

        $$ (g_1 \vee g_2)(k) := g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) $$

        where the recursion is well-founded due to $d$ not containing any cycles.

        Computationally, a mapping may be represented as a set of $K \times V$ pairs (perhaps in a database), with the remaining keys being mapped to their default values; such a mapping is called \emph{finite}. The join of two finite mappings is finite, since the join of two identical default values is itself a default value (due to $\vee_{k, g}$ being idempotent). Alternatively, an infinite mapping may be represented as a function.

        Merging two mappings can be done recursively. Given mappings $g_1$ and $g_2$ and a key $k$, we compute the values $g_1(k), g_2(k)$ and their dependency sets $d(k, g_1(k)), d(k, g_2(k))$. We recursively merge $g_1$ and $g_2$ on each key in either dependency set, caching intermediate values to avoid repetated recursive calls. Finally, we compute the join and cache it in the database.

        We may consider merging a mapping with a singleton mapping, which specifies the value of a single key and otherwise contains only default values. In this case, merging will change this value and all its direct or indirect dependencies. Merging with a singleton is similar to a ``database insert'', updating the value of a given key.

\end{document}

### 2022-09-19 paper/paper.tex
autofill

@@
 
+\usepackage{amsthm}
 \usepackage{amsmath}
 \usepackage{amssymb}
+\newtheorem{theorem}{Theorem}
 
@@
             \item A set of values $V$.
-            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a set of keys, its dependencies.
+            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a finite set of keys, its dependencies.
             \item A join function $j : (\Sigma (k: K, x : V, y : V), d(k, x) \cup d(k, y) \rightarrow V) \rightarrow V$ that joins two values given their dependencies.
@@
             \item $d$ must be well-founded; that is, there are no infinite sequences $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, \ldots$
-            \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
+            \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g|_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
             \item $j$ must not grow dependencies; that is, for any $k : K$, $g: K \rightarrow V$, $x, y \in V$, we must have $d(k, x \vee_{k, g} y) \subseteq d(k, x) \cup d(k, y)$.
         \end{itemize}
-
-        A semilattice graph \emph{has defaults} if there is a function $z : K \rightarrow V$ such that $d(k, z(k)) = \{\}$ for each $k \in K$.
-
-    \section{Joins on semilattice graphs}
 
@@
 
-        where the recursion is well-founded due to $d$ not containing any cycles.
+        where the recursion is well-founded due to $d$ being well-founded.
 
-        Computationally, a mapping may be represented as a set of $K \times V$ pairs (perhaps in a database), with the remaining keys being mapped to their default values; such a mapping is called \emph{finite}.
-The join of two finite mappings is finite, since the join of two identical default values is itself a default value (due to $\vee_{k, g}$ being idempotent).
-Alternatively, an infinite mapping may be represented as a function.
+    \section{Autofills}
+
+        Mappings may be infinite in general, and even finite ones contain a large number of keys.
+It is therefore computationally infeasible to store a mapping as a set of $K \times V$ pairs.
+Accordingly, we define \emph{autofills} to fill in lattice values automatically from a set of other values.
+
+        Define $d^*(k, g) := \{k\} \cup \bigcup_{k' \in d(k, g(k))} d^*(k', g)$, the transitive closure of $d$, where $g$ is a mapping.
+The recursion is well-founded (and $d^*(k, g)$ is finite) because $d$ is well-founded.
+
+        Note that $d^*(k, g_1 \vee g_2) \subseteq d^*(k, g_1) \cup d^*(k, g_2)$.
+
+        An \emph{autofill} for a semilattice graph is a function $a : \forall (k:K, g: V \rightarrow K), V$ that satisfies:
+
+        \begin{itemize}
+          \item
+            $a(k, g_1) = a(k, g_2)$ if $g_1(k) = g_2(k)$ for each $k \in d^*(k, g_1)$; that is, $a(k, g)$ must only depend on the values of $g$ on $d^*(k, g)$.
+          \item
+            $a(k, g_1) \vee_{k, g_1 \vee g_2} a(k, g_2) = a(k, g_1 \vee g_2)$; that is, the join of two autofill values on two mappings must be an autofill value on the joined mapping.
+        \end{itemize}
+
+        A mapping $g : K \rightarrow V$ is an \emph{autofill mapping} if:
+
+        \begin{itemize}
+          \item $g(k) = a(k, g)$ for all but a finite number of $k$ values.
+          \item $g(k) \vee_{k, g} a(k, g) = g(k)$ for all $k \in K$.
+        \end{itemize}
+
+        An autofill mapping can be represented as a finite set of $K \times V$ pairs; the remaining values are filled in automatically.
+
+        \begin{theorem}
+          Let $g_1, g_2 : K \rightarrow V$ be autofill mappings. Then $g_1 \vee g_2$ is an autofill mapping.
+        \end{theorem}
+
+        \begin{proof}
+
+          Whenever $g_1(k) = a(k, g_1)$ and $g_2(k) = a(k, g_2)$ (which is true for all but a finite number of $k$ values):
+           $$(g_1 \vee g_2)(k) = g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) = a(k, g_1) \vee_{k, (g_1 \vee g_2)} a(k, g_2) = a(k, g_1 \vee g_2)$$
+
+          Additionally, for each $k \in K$:
+          \begin{align*}
+            (g_1 \vee g_2)(k) \vee_{k, (g_1 \vee g_2)} a(k, g) 
+            &= (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee g_2)} a(k, g) \\
+            &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2(k) \vee_{k, (g_1 \vee g_2)} a(k, g)) \\
+            &= g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) \\
+            &= (g_1 \vee g_2)(k)
+          \end{align*}
+        \end{proof}
 

### 2022-09-19 src/lattice.rs

@@ -17,7 +17,7 @@ pub trait LatGraph : Send + Sync {
     /// Joins two values given dependencies.
     fn join(&self, key: &Self::K, v1: &Self::V, v2: &Self::V, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
 
-    /// The default element of the lattice.
+    /// The default element of the lattice. Must have no dependencies.
     fn default(&self, key: &Self::K) -> Result<Self::V, String>;
 }
 

### 2022-09-19 src/raw_db.rs

@@ -6,88 +6,83 @@ use async_trait::async_trait;
 
 use super::lattice::{LatGraph, LatticeReadDB, LatticeWriteDB, DependencyLatticeDB};
 
-pub trait LatticeRawWriteDB<LID, Value> {
-    fn get_lattice_max(&self, lid: &LID) -> Result<Option<Value>, String>;
-    fn set_lattice_max(&mut self, lid: &LID, value: Option<Value>) -> Result<(), String>;
+pub trait LatticeRawMappingDB<K, V> {
+    fn get_lattice_max(&self, key: &K) -> Result<Option<V>, String>;
+    fn set_lattice_max(&mut self, key: &K, value: Option<V>, deps: BTreeSet<K>) -> Result<(), String>;
 
-    fn is_dirty(&self, lid: &LID) -> Result<bool, String>;
-    fn set_dirty(&mut self, lid: &LID) -> Result<(), String>;
-    fn set_undirty(&mut self, lid: &LID) -> Result<(), String>;
+    fn is_dirty(&self, key: &K) -> Result<bool, String>;
+    fn set_dirty(&mut self, key: &K) -> Result<(), String>;
+    fn set_undirty(&mut self, key: &K) -> Result<(), String>;
     fn get_some_dirty(&mut self) -> Result<Option<LID>, String>;
 
-    fn get_dependencies(&self, lid: &LID) -> Result<BTreeSet<LID>, String>;
-    fn set_dependencies(&mut self, lid: &LID, deps: BTreeSet<LID>) -> Result<(), String>;
-
-    fn get_dependents(&self, lid: &LID) -> Result<BTreeSet<LID>, String>;
-    fn set_dependents(&mut self, lid: &LID) -> Result<BTreeSet<LID>, String>;
-    fn remove_dependent(&mut self, lid: &LID, dep: &LID) -> Result<(), String>;
-    fn add_dependent(&mut self, lid: &LID, dep: &LID) -> Result<(), String>;
+    fn get_dependencies(&self, key: &K) -> Result<BTreeSet<LID>, String>;
+    fn get_dependents(&self, key: &K) -> Result<BTreeSet<LID>, String>;
 }
 
-pub struct LatticeWrapperWriteDB<L: LatGraph> {
-    latgraph: Arc<L>,
-    raw_db: Mutex<Box<dyn Send + LatticeRawWriteDB<L::LID, L::Value>>>,
-}
-
[…]

### 2022-09-19 src/lattice.rs
default to autofill

@@ -17,8 +17,8 @@ pub trait LatGraph : Send + Sync {
     /// Joins two values given dependencies.
     fn join(&self, key: &Self::K, v1: &Self::V, v2: &Self::V, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
 
-    /// The default element of the lattice. Must have no dependencies.
-    fn default(&self, key: &Self::K) -> Result<Self::V, String>;
+    /// The default element of the lattice.
+    fn autofill(&self, key: &Self::K, vals: &mut dyn FnMut(&Vec<Self::K>) -> Result<Self::V, String>) -> Result<Self::V, String>;
 }
 
 
@@ -61,9 +61,16 @@ impl<'a, L: LatGraph + 'static> LatGraph for SerializeLatGraph<'a, L> {
         rmp_serde::to_vec_named(&v).map_err(|e| format!("failed to serialize value: {}", e))
     }
 
-    fn default(&self, key: &Self::K) -> Result<Self::V, String> {
+    fn autofill(&self, key: &Self::K, vs: &mut dyn FnMut(&Vec<Self::K>) -> Result<Self::V, String>) -> Result<Self::V, String> {
         let key = rmp_serde::from_slice(key).map_err(|e| format!("failed to deserialize key: {}", e))?;
-        let v = self.0.default(&key)?;
+        let v = self.0.autofill(&key, &mut move |ks| {
+            let mut ks2 = Vec::new();
+            for k in ks {
+                ks2.push(rmp_serde::to_vec(&k).map_err(|e| format!("failed to serialize key: {}", e))?);
+            }
+            let res = vs(&ks2)?;
+            rmp_serde::from_slice(&res).map_err(|e| format!("failed to deserialize value: {}", e))
+        })?;
         rmp_serde::to_vec_named(&v).map_err(|e| format!("failed to serialize value: {}", e))
     }
 }
@@ -112,11 +119,11 @@ where L::K: IsEnum, L::V: IsEnum {
         self.0[branch].join(key, v1, v2, deps)
     }
 
-    fn default(&self, key: &Self::K) -> Result<Self::V, String> {
+    fn autofill(&self, key: &Self::K, vals: &mut dyn FnMut(&Vec<Self::K>) -> Result<Self::V, String>) -> Result<Self::V, String> {
         let branch = key.get_branch();
         if branch >= self.0.len() {
             return Err(format!("branch {} out of range", branch));
         }
[…]

### 2022-09-19 src/lattice.rs
autofill async

@@ -1,8 +1,13 @@
+use std::sync::Arc;
 use std::collections::{BTreeSet, BTreeMap};
+use std::marker::PhantomData;
+
+use async_trait::async_trait;
 
 use serde::{Serialize, de::DeserializeOwned};
 
 /// A graph where each item is a semilattice, which may depend on other lattice max values.
+#[async_trait]
 pub trait LatGraph : Send + Sync {
 
     /// Keys identifying lattice value keys.
@@ -18,7 +23,12 @@ pub trait LatGraph : Send + Sync {
     fn join(&self, key: &Self::K, v1: &Self::V, v2: &Self::V, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
 
     /// The default element of the lattice.
-    fn autofill(&self, key: &Self::K, vals: &mut dyn FnMut(&Vec<Self::K>) -> Result<Self::V, String>) -> Result<Self::V, String>;
+    async fn autofill(&self, key: &Self::K, vals: Arc<dyn AutofillDatabase<Self::K, Self::V>>) -> Result<Self::V, String>;
+}
+
+#[async_trait]
+pub trait AutofillDatabase<K, V> : Send + Sync {
+    async fn get_value(self: Arc<Self>, keys: Vec<K>) -> Result<V, String>;
 }
 
 
@@ -31,6 +41,7 @@ impl<'a, L: LatGraph> SerializeLatGraph<'a, L> {
     }
 }
 
+#[async_trait]
 impl<'a, L: LatGraph + 'static> LatGraph for SerializeLatGraph<'a, L> {
 
     type K = Vec<u8>;
@@ -61,17 +72,33 @@ impl<'a, L: LatGraph + 'static> LatGraph for SerializeLatGraph<'a, L> {
         rmp_serde::to_vec_named(&v).map_err(|e| format!("failed to serialize value: {}", e))
     }
 
-    fn autofill(&self, key: &Self::K, vs: &mut dyn FnMut(&Vec<Self::K>) -> Result<Self::V, String>) -> Result<Self::V, String> {
+    async fn autofill(&self, key: &Self::K, vals: Arc<dyn AutofillDatabase<Self::K, Self::V>>) -> Result<Self::V, String> {
         let key = rmp_serde::from_slice(key).map_err(|e| format!("failed to deserialize key: {}", e))?;
-        let v = self.0.autofill(&key, &mut move |ks| {
-            let mut ks2 = Vec::new();
[…]

### 2022-09-19 src/mapping.rs
mapping autofill

@@ -3,7 +3,7 @@ use std::sync::{Arc, Mutex};
 
 use async_trait::async_trait;
 
-use super::lattice::LatGraph;
+use super::lattice::{LatGraph, AutofillDatabase};
 
 #[async_trait]
 pub trait LatMapping<L: LatGraph> : Send + Sync {
@@ -13,7 +13,18 @@ pub trait LatMapping<L: LatGraph> : Send + Sync {
     async fn get_lattice_max(self: Arc<Self>, key: L::K) -> Result<L::V, String>;
 }
 
-impl<L: LatGraph> dyn LatMapping<L> {
+#[async_trait]
+pub trait LatMappingExt<L: LatGraph> : LatMapping<L> {
+
+    async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String>;
+
+    async fn join(self: Arc<Self>, key: L::K, v1: L::V, v2: L::V) -> Result<L::V, String>;
+
+    async fn autofill(self: Arc<Self>, key: L::K) -> Result<L::V, String>;
+}
+
+#[async_trait]
+impl<L: LatGraph + 'static, M: LatMapping<L> + 'static> LatMappingExt<L> for M {
     async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String> {
         let value = self.clone().get_lattice_max(key.clone()).await?;
         self.get_latgraph().dependencies(&key, &value)
@@ -27,6 +38,41 @@ impl<L: LatGraph> dyn LatMapping<L> {
             deps.insert(dep.clone(), self.clone().get_lattice_max(dep.clone()).await?);
         }
         self.get_latgraph().join(&key, &v1, &v2, &deps)
+    }
+
+    async fn autofill(self: Arc<Self>, key: L::K) -> Result<L::V, String> {
+        let db = Arc::new(MappingAutofillDatabase::new(key.clone(), self.clone()));
+        self.get_latgraph().autofill(&key, db).await
+    }
+}
+
+struct MappingAutofillDatabase<L: LatGraph, M: LatMapping<L> + ?Sized> {
+    key: L::K,
+    mapping: Arc<M>
+}
+
+impl <L: LatGraph, M: LatMapping<L>> MappingAutofillDatabase<L, M> {
+    fn new(key: L::K, mapping: Arc<M>) -> Self {
+        MappingAutofillDatabase {key, mapping}
+    }
+}
+
+#[async_trait]
[…]

### 2022-09-19 paper/paper.tex
simplify paper autofill dependencies

@@
 
-        Define $d^*(k, g) := \{k\} \cup \bigcup_{k' \in d(k, g(k))} d^*(k', g)$, the transitive closure of $d$, where $g$ is a mapping.
+        % Define $d^*(k, g) := \{k\} \cup \bigcup_{k' \in d(k, g(k))} d^*(k', g)$, the transitive closure of $d$, where $g$ is a mapping.
 The recursion is well-founded (and $d^*(k, g)$ is finite) because $d$ is well-founded.
 
-        Note that $d^*(k, g_1 \vee g_2) \subseteq d^*(k, g_1) \cup d^*(k, g_2)$.
+        % Note that $d^*(k, g_1 \vee g_2) \subseteq d^*(k, g_1) \cup d^*(k, g_2)$.
 
-        An \emph{autofill} for a semilattice graph is a function $a : \forall (k:K, g: V \rightarrow K), V$ that satisfies:
+        An \emph{autofill} for a semilattice graph consists of a dependency function $\hat{d}: K \rightarrow 2^V$ and an autofill function $a : \forall (k:K), \hat{d}(k) \rightarrow V$ that satisfy:
 
@@
           \item
-            $a(k, g_1) = a(k, g_2)$ if $g_1(k) = g_2(k)$ for each $k \in d^*(k, g_1)$; that is, $a(k, g)$ must only depend on the values of $g$ on $d^*(k, g)$.
+            For each $k \in K, v \in V$, $\hat{d}(k) \subseteq d(k, v)$.
           \item
@@
         \end{itemize}
+
+        where we write $a(k, g) := a(k, g|_{\hat{d}(k)})$ for brevity.
 
@@
           Whenever $g_1(k) = a(k, g_1)$ and $g_2(k) = a(k, g_2)$ (which is true for all but a finite number of $k$ values):
-           $$(g_1 \vee g_2)(k) = g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) = a(k, g_1) \vee_{k, (g_1 \vee g_2)} a(k, g_2) = a(k, g_1 \vee g_2)$$
+           $$(g_1 \vee g_2)(k) = g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) = a(k, g_1) \vee_{k, g_1 \vee g_2} a(k, g_2) = a(k, g_1 \vee g_2)$$
 
@@
 
+        For an autofill mapping $g$, we can define the \emph{concrete dependencies} of a key $k \in K$, which are a set of non-autofill keys of $g$ that are sufficient to derive $k$'s dependencies:
+
+        $$c(k, g) := \bigcup_{k' \in d(k, g(k))} \begin{cases}
+          \{k'\} & \text{if } g(k') \neq a(k, g) \\
+          c(k', g) & \text{otherwise}
+        \end{cases}  $$
+
         Merging two mappings can be done recursively.

### 2022-09-19 src/lattice.rs
simpler autofill!

@@ -1,13 +1,8 @@
-use std::sync::Arc;
 use std::collections::{BTreeSet, BTreeMap};
-use std::marker::PhantomData;
-
-use async_trait::async_trait;
 
 use serde::{Serialize, de::DeserializeOwned};
 
 /// A graph where each item is a semilattice, which may depend on other lattice max values.
-#[async_trait]
 pub trait LatGraph : Send + Sync {
 
     /// Keys identifying lattice value keys.
@@ -17,21 +12,14 @@ pub trait LatGraph : Send + Sync {
     type V : Eq + Clone + Serialize + DeserializeOwned + Send + Sync;
 
     /// Dependencies of a given key.
-    fn dependencies(&self, key: &Self::K, value: &Self::V) -> Result<BTreeSet<Self::K>, String>;
+    fn dependencies(&self, key: &Self::K, value: Option<&Self::V>) -> Result<BTreeSet<Self::K>, String>;
 
     /// Joins two values given dependencies.
     fn join(&self, key: &Self::K, v1: &Self::V, v2: &Self::V, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
 
     /// The default element of the lattice.
-    async fn autofill(&self, key: &Self::K, vals: Arc<dyn AutofillDatabase<Self::K, Self::V>>) -> Result<Self::V, String>;
+    fn autofill(&self, key: &Self::K, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
 }
-
-#[async_trait]
-pub trait AutofillDatabase<K, V> : Send + Sync {
-    async fn get_value(self: Arc<Self>, keys: Vec<K>) -> Result<V, String>;
-}
-
-
 
 pub struct SerializeLatGraph<'a, L: LatGraph>(&'a L);
 
@@ -41,18 +29,23 @@ impl<'a, L: LatGraph> SerializeLatGraph<'a, L> {
     }
 }
 
-#[async_trait]
 impl<'a, L: LatGraph + 'static> LatGraph for SerializeLatGraph<'a, L> {
 
     type K = Vec<u8>;
 
     type V = Vec<u8>;
 
-    fn dependencies(&self, key: &Vec<u8>, value: &Vec<u8>) -> Result<BTreeSet<Vec<u8>>, String> {
+    fn dependencies(&self, key: &Vec<u8>, value: Option<&Vec<u8>>) -> Result<BTreeSet<Vec<u8>>, String> {
[…]

### 2022-09-19 src/mapping.rs

@@ -3,7 +3,7 @@ use std::sync::{Arc, Mutex};
 
 use async_trait::async_trait;
 
-use super::lattice::{LatGraph, AutofillDatabase};
+use super::lattice::LatGraph;
 
 #[async_trait]
 pub trait LatMapping<L: LatGraph> : Send + Sync {
@@ -27,13 +27,13 @@ pub trait LatMappingExt<L: LatGraph> : LatMapping<L> {
 impl<L: LatGraph + 'static, M: LatMapping<L> + 'static> LatMappingExt<L> for M {
     async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String> {
         let value = self.clone().get_lattice_max(key.clone()).await?;
-        self.get_latgraph().dependencies(&key, &value)
+        self.get_latgraph().dependencies(&key, Some(&value))
     }
 
     async fn join(self: Arc<Self>, key: L::K, v1: L::V, v2: L::V) -> Result<L::V, String> {
         let mut deps = BTreeMap::new();
-        let deps1 = self.get_latgraph().dependencies(&key, &v1)?;
-        let deps2 = self.get_latgraph().dependencies(&key, &v2)?;
+        let deps1 = self.get_latgraph().dependencies(&key, Some(&v1))?;
+        let deps2 = self.get_latgraph().dependencies(&key, Some(&v2))?;
         for dep in deps1.union(&deps2) {
             deps.insert(dep.clone(), self.clone().get_lattice_max(dep.clone()).await?);
         }
@@ -41,38 +41,11 @@ impl<L: LatGraph + 'static, M: LatMapping<L> + 'static> LatMappingExt<L> for M {
     }
 
     async fn autofill(self: Arc<Self>, key: L::K) -> Result<L::V, String> {
-        let db = Arc::new(MappingAutofillDatabase::new(key.clone(), self.clone()));
-        self.get_latgraph().autofill(&key, db).await
-    }
-}
-
-struct MappingAutofillDatabase<L: LatGraph, M: LatMapping<L> + ?Sized> {
-    key: L::K,
-    mapping: Arc<M>
-}
-
-impl <L: LatGraph, M: LatMapping<L>> MappingAutofillDatabase<L, M> {
-    fn new(key: L::K, mapping: Arc<M>) -> Self {
-        MappingAutofillDatabase {key, mapping}
-    }
-}
-
-#[async_trait]
[…]

### 2022-09-19 paper/paper.tex
gitignore

@@
         \begin{itemize}
-            \item $d$ must be well-founded; that is, there are no infinite sequences $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, \ldots$
+            \item $d$ must be well-founded; that is, there are no infinite sequences $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, 2, \ldots$
             \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g|_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.

### 2022-09-30 paper/paper.tex
about to refactor a lot!

--- 2022-09-19 paper/paper.tex
@@
             \item A set of keys $K$.
-            \item A set of values $V$.
+            \item A nonempty set of values $V$.
             \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a finite set of keys, its dependencies.
@@
         \begin{itemize}
-            \item $d$ must be well-founded; that is, there are no infinite sequences $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, \ldots$
+            \item $d$ must be well-founded; that is, there are no infinite dependency chains $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, 2, \ldots$
             \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g|_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
@@
         where the recursion is well-founded due to $d$ being well-founded.
+
+        We must now show that $\vee$ is commutative, associative, and idempotent, inductively.
+For any natural $n$ we say $\vee$ (applied to semilattice graphs) forms a semilattice up to $n$ if, for any $k : K$ whose longest dependency chain is no longer than $n$, we have:
+
+        \begin{itemize}
+          \item $(g_1 \vee g_2)(k) = (g_2 \vee g_1)(k)$
+          \item $((g_1 \vee g_2) \vee g_3)(k) = (g_1 \vee (g_2 \vee g_3))(k)$
+          \item $(g \vee g)(k) = g(k)$
+        \end{itemize}
+
+        We note that if $\vee$ forms a semilattice up to all $n$, then it forms a semilattice.
+We now note that if $\vee$ forms a semilattice up to $n$, then it forms a semilattice up to $n+1$, inductively:
+
+        % WANT: (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k)
+        %     = g_1(k) \vee_{k, g_1 \vee g_2 \vee g_3} (g_2(k) \vee_{k, g_2 \vee g_3} g_3(k))
+        % HAVE: (g_1(k) \vee_{k, (g_1 \vee g_2 \vee g_3)}) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k)
+        %     = g_1(k) \vee_{k, g_1 \vee g_2 \vee g_3} (g_2(k) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k))
+
+        \begin{itemize}
+          \item $(g_1 \vee g_2)(k) = g_2(k) \vee_{k, (g_1 \vee g_2)} g_1(k) = g_2(k) \vee_{k, (g_2 \vee g_1)} g_1(k) = (g_2 \vee g_1)(k)$
+          \item
+            \begin{align*}
+              ((g_1 \vee g_2) \vee g_3))(k) &= (g_1 \vee g_2)(k) \vee_{k, ((g_1 \vee g_2) \vee g_3)} g_3(k) \\
+              &= (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee (g_2 \vee g_3))} g_3(k) \\
+              &= g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) \vee_{k, (g_1 \vee g_2)} g_3(k) \\
+              &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2(k) \vee_{k, (g_1 \vee g_2)} g_3(k)) \\
+              &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2 \vee g_3)(k) \\
+              &= (g_1 \vee (g_2 \vee g_3))(k)
+            \end{align*}
+          \item $(g_1 \vee g_1)(k) = g_1(k) \vee_{k, (g_1 \vee g_1)} g_1(k) = g_1(k)$
+        \end{itemize}
 

### 2022-10-01 paper/paper.tex
categories!

@@
 
+        We write $x \leq y$ to mean $y = x \vee y$.
+
         In databases, lattices are useful for defining a notion of a most updated value.
 For example, with $L$ as a set of sets of messages and $\vee = \cup$, the lattice represents sets of messages that are updated to include more messages as they come in.
+
+    \section{Semilattices as a category}
+
+      We discuss category theoretic language for semilattices to derive a notion of dependency between semilattices.
+In the next section, we drop the category-theoretic language, so it is not strictly necessary to understand category theory to understand the result, although it helps with the derivation.
+
+      A semilattice can be interpreted as a category whose objects are the members of $L$, and where a morphism exists between $x$ and $y$ if and only if $x \leq y$; in this case there must be exactly one morphism of this form denoted $x \leq y$.
+To be a semilattice, the category must have coproducts (corresponding to joins).
+
+      Note that if $x \leq y$ and $y \leq x$, then $x = x \vee y = y$.
+
+      Semilattices form a category $\mathrm{Semilattice}$.
+The objects are semilattices and the morphisms are coproduct-preserving functors between semilattices.
+A functor between two semilattices $L_1, L_2$ consists of a function $f: L_1 \rightarrow L_2$ mapping elements of $L_1$ to elements of $L_2$, such that if $x \leq y$, then $f(x) \leq f(y)$.
+The functor is said to preserve coproducts/joins if $f(x \vee y) = f(x) \vee f(y)$ for all $x, y : L_1$.
+The $\mathrm{Semilattice}$ category has products in the obvious way.
+
+      A semilattice fibration on a semilattice $L$ is a functor $g : L \rightarrow \mathrm{Semilattice}$, which assigns to each element $x$ of $L$ a semilattice $g(x)$, and assigns to each ordering $x \leq y$ a semilattice functor $g(x \leq y)$ from $g(x)$ to $g(y)$.
+
+      The total space's elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
+We consider $(x, x') \leq (y, y')$ when $x \leq y$ and $g(x \leq y)(x') \leq y'$.
+The total space is a semilattice, as shown:
+
+      \begin{theorem}
+        The total space of a semilattice fibration $g$ is a semilattice.
+      \end{theorem}
+      \begin{proof}
+        Consider two elements of the total space, $(x, x'), (y, y')$.
+Their join is defined as $(x, x') \vee (y, y') := (x \vee y', g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y'))$.
+We must show that this join is commutative, associative, and idempotent.
+        \begin{itemize}
+          \item \emph{Commutative}:
+            \begin{align*}
+              (x, x') \vee (y, y') &= (x \vee y, g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y')) \\
+              &= (y \vee x, g(y \leq y \vee x)(y') \vee g(x \leq y \vee x)(x')) \\
+              &= (y, y') \vee (x, x')
+            \end{align*}
+          \item \emph{Associative}:
+            \begin{align*} &((x, x') \vee (y, y')) \vee (z, z')  \\
+              =& (x \vee y \vee z, g(x \vee y \leq x \vee y \vee z)(g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y')) \vee g(z \leq x \vee y \vee z)(z')) \\
+              =& (x \vee y \vee z, g(x \leq x \vee y \vee z)(x') \vee g(y \leq x \vee y \vee z)(y') \vee g(z \leq x \vee y \vee z)(z') \\
+              =& (x \vee y \vee z, g(x \leq x \vee y \vee z)(x') \vee g(y \vee z \leq x \vee y \vee z)(g(y \leq y \vee z)(y') \vee g(z \leq y \vee z)(z')) \\
+              =& (x, x') \vee ((y, y') \vee (z, z'))
+            \end{align*}
+          \item \emph{Idempotent}: 
+            \begin{align*}
+              (x, x') \vee (x, x') &= (x \vee x, g(x \leq x \vee x)(x') \vee g(x \leq x \vee x)(x')) \\
+              &= (x, g(x \leq x)(x')) \\
+              &= (x, x')
+            \end{align*}
+        \end{itemize}
+      \end{proof}
+
+    \section{De-categorizing semilattice fibrations}
+
+      To define a semilattice fibration more directly, we can remove the category-theoretic language, achieving a more direct presentation.
+
+      A semilattice fibration consists of a function $g_l : L \rightarrow \mathrm{Semilattice}$, assigning to each element of $L$ a semilattice, and a function $g_t : \Pi (x, y : L, x \leq y, x' : g(x)), g(y)$ (a ``transport function''), satisfying:
+
+      \begin{itemize}
+        \item For any $x : L, x' : g_l(x)$: $$g_t(x, x, x') = x'$$.
+        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : g_l(x)$: $$g_t(y, z, g_t(x, y, x')) = g_t(x, z, x')$$.
+        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : g_l(x)$ such that $x'_1 \leq x'_2$: $$g_t(x, y, x'_1) \leq g_t(x, y, x'_2)$$.
+      \end{itemize}
+
+      This is equivalent to the above category-theoretic definition.
+$g_l$ corresponds to the semilattice fibration functor's action on objects, and $g_t$ to its action on morphisms (itself a functor).
+The first two conditions ensure that $(g_l, g_t)$ is a functor, while the third ensures that $g_t(x, y, \cdot)$ is a semilattice functor.
+
+      Intuitively, the first condition states that transporting a value $x'$ when not changing its dependency $x$ doesn't change $x'$; the second condition states dependency updates aren't path-dependent; and the third states that transports preserve ordering.
+
+      The total space is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : g_l(x)$.
+The join is defined as $(x, x') \vee (y, y') := (x \vee y, g_t(x, x \vee y, x') \vee g_t(y, x \vee y, y'))$.
+This is equivalent to the above category-theoretic definition, proving it is a semilattice.
 
@@
             \item A nonempty set of values $V$.
-            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key to a finite set of keys, its dependencies.
-            \item A join function $j : (\Sigma (k: K, x : V, y : V), d(k, x) \cup d(k, y) \rightarrow V) \rightarrow V$ that joins two values given their dependencies.
-(The sigma set $\Sigma (a: A), B(a)$ denotes a set of pairs $(a, b)$ where $a \in A$ and $b \in B(a)$.)
+            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key and value to a finite set of keys, its dependencies.
+            \item A transport function $t : \Pi (k : K, v : V), (d(k, v) \rightarrow V)^2 \rightarrow V$ that transports a value when its dependencies change.
+            \item A join function $j : \Pi (k : K, v_1, v_2 : V), (d(k, v_1) \cup d(k, v_2) \rightarrow V) \rightarrow V$ that maps two values to a value given their dependencies.
         \end{itemize}
 
-        And satisfies the following properties:
+        We write $v_1 \vee_{k, g} v_2 := j(k, v_1, v_2, g|_{d(k, v_1) \cup d(k, v_2)})$, and $t_{k, g_1, g_2}(v) := t(k, v, g_1|_{d(k, v)}, g_2|_{d(k, v)})$.
+
+        The semilattice graph must satisfy the following properties:
 
@@
             \item $d$ must be well-founded; that is, there are no infinite dependency chains $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, 2, \ldots$
-            \item $j$ must be a valid join function; for any $k : K$, $g: K \rightarrow V$, defining $x \vee_{k,g} y := j(k, x, y, g|_{d(k, x) \cup d(k, y)})$, then $(V, \vee_{k,g})$ must be a semilattice.
-            \item $j$ must not grow dependencies; that is, for any $k : K$, $g: K \rightarrow V$, $x, y \in V$, we must have $d(k, x \vee_{k, g} y) \subseteq d(k, x) \cup d(k, y)$.
+            \item $t$ must not grow dependencies: for $k : K, v : V, g_1, g_2 : d(k, v) \rightarrow V$, we must have $d(k, t(k, v, g_1, g_2)) \subseteq d(k, v)$.
+            \item $j$ must not grow dependencies; that is, for any $k : K, v_1 : V, v_2: V, g: (d(k, v_1) \cup d(k, v_2)) \rightarrow V$, we must have $d(k, j(k, v_1, v_2, g)) \subseteq d(k, v_1) \cup d(k, v_2)$.
+            \item $t$ must be an identity when dependencies are equal: for $k : K, v : V, g : d(k, v) \rightarrow V$, we must have $t(k, v, g, g) = v$.
+            \item $t$ must not be path-dependent: for $k : K, v : V, g_1, g_2, g_3 : d(k, v) \rightarrow V$, we must have $t(k, t(k, v, g_1, g_2), g_2, g_3) = t(k, v, g_1, g_3)$.
+            \item $t$ must preserve orderings: for $k : K, v_1, v_2 : V, g_1, g_2 : d(k, v) \rightarrow V$, such that $j(k, v_1, v_2, g_1) = v_2$, we must have $j(k, t(k, v_1, g_1, g_2), t(k, v_2, g_1, g_2)) = t(k, v_2, g_1, g_2)$.
+            \item For any $k: K, g : K \rightarrow V$, it must be that $\vee_{k, g}$ is commutative, associative, and idempotent.
         \end{itemize}
 
-        It is now possible to define a semilattice over a semilattice graph. This lattice's items are functions $K \rightarrow V$ (called \emph{mappings}) and the join operation is defined as follows:
+        It is now possible to define a semilattice over a semilattice graph.
+This lattice's items are functions $g : K \rightarrow V$ (called \emph{mappings}).
+The join operation is defined as follows:
 
-        $$ (g_1 \vee g_2)(k) := g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) $$
-
-        where the recursion is well-founded due to $d$ being well-founded.
+        $$ (g_1 \vee g_2)(k) := t_{k, g_1, (g_1 \vee g_2)}(g_1(k)) \vee_{k, (g_1 \vee g_2)} t_{k, g_2, (g_1 \vee g_2)}(g_2(k))$$.
 
@@
 
-        % WANT: (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k)
-        %     = g_1(k) \vee_{k, g_1 \vee g_2 \vee g_3} (g_2(k) \vee_{k, g_2 \vee g_3} g_3(k))
-        % HAVE: (g_1(k) \vee_{k, (g_1 \vee g_2 \vee g_3)}) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k)
-        %     = g_1(k) \vee_{k, g_1 \vee g_2 \vee g_3} (g_2(k) \vee_{k, (g_1 \vee g_2 \vee g_3)} g_3(k))
-
         \begin{itemize}
-          \item $(g_1 \vee g_2)(k) = g_2(k) \vee_{k, (g_1 \vee g_2)} g_1(k) = g_2(k) \vee_{k, (g_2 \vee g_1)} g_1(k) = (g_2 \vee g_1)(k)$
-          \item
+          \item \emph{Commutative}:
             \begin{align*}
-              ((g_1 \vee g_2) \vee g_3))(k) &= (g_1 \vee g_2)(k) \vee_{k, ((g_1 \vee g_2) \vee g_3)} g_3(k) \\
-              &= (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee (g_2 \vee g_3))} g_3(k) \\
-              &= g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) \vee_{k, (g_1 \vee g_2)} g_3(k) \\
-              &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2(k) \vee_{k, (g_1 \vee g_2)} g_3(k)) \\
-              &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2 \vee g_3)(k) \\
-              &= (g_1 \vee (g_2 \vee g_3))(k)
+              (g_1 \vee g_2)(k) &= t_{k, g_1, (g_1 \vee g_2)}(g_1(k)) \vee_{k, (g_1 \vee g_2)} t_{k, g_2, (g_1 \vee g_2)}(g_2(k)) \\
+              &= t_{k, g_2, (g_2 \vee g_1)}(g_2(k)) \vee_{k, (g_2 \vee g_1)} t_{k, g_1, (g_2 \vee g_1)}(g_1(k)) \\
+              &= (g_2 \vee g_1)(k)
             \end{align*}
-          \item $(g_1 \vee g_1)(k) = g_1(k) \vee_{k, (g_1 \vee g_1)} g_1(k) = g_1(k)$
+          \item \emph{Associative}:
+            \begin{align*}
+              &((g_1 \vee g_2) \vee g_3))(k) \\
+              =& t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}((g_1 \vee g_2)(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
+              =& t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}(t_{k, g_1, (g_1 \vee g_2)}(g_1(k)) \vee_{k, (g_1 \vee g_2)} t_{k, g_2, (g_1 \vee g_2)}(g_2(k))) \\
+               & \vee_{k, (g_1 \vee g_2 \vee g_3)} t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
+              =& t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}(t_{k, g_1, (g_1 \vee g_2)}(g_1(k))) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
+               & t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}(t_{k, g_2, (g_1 \vee g_2)}(g_2(k))) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
+               & t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
+              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)}
+                 t_{k, g_2, (g_1 \vee g_2 \vee g_3)}(g_2(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)}
+                 t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
+              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
+               & t_{k, (g_2 \vee g_3), (g_1 \vee g_2 \vee g_3)}(t_{k, g_2, (g_2 \vee g_3)}(g_2(k))) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
+               & t_{k, (g_2 \vee g_3), (g_1 \vee g_2 \vee g_3)}(t_{k, g_3, (g_2 \vee g_3)}(g_3(k))) \\
+              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
+               & t_{k, (g_2 \vee g_3)}(t_{k, g_2, (g_2 \vee g_3)}(g_2(k)) \vee_{k, (g_2 \vee g_3)} t_{k, g_3, (g_2 \vee g_3)}(g_3(k))) \\
+              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} t_{k, (g_2 \vee g_3)}((g_2 \vee g_3)(k)) \\
+              =& (g_1 \vee (g_2 \vee g_3))(k)
+            \end{align*}
+          \item \emph{Idempotent}:
+            \begin{align*}
+              (g \vee g)(k) &= t_{k, g, (g \vee g)}(g(k)) \vee_{k, (g \vee g)} t_{k, g, (g \vee g)}(g(k)) \\
+              &= t_{k, g, g}(g(k)) \vee_{k, g} t_{k, g, g}(g(k)) \\
+              &= g(k)
+            \end{align*}
         \end{itemize}

### 2022-10-01 paper/paper.tex
transport distributes over joins

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem{lemma}{Lemma}
 
@@
 The objects are semilattices and the morphisms are coproduct-preserving functors between semilattices.
-A functor between two semilattices $L_1, L_2$ consists of a function $f: L_1 \rightarrow L_2$ mapping elements of $L_1$ to elements of $L_2$, such that if $x \leq y$, then $f(x) \leq f(y)$.
-The functor is said to preserve coproducts/joins if $f(x \vee y) = f(x) \vee f(y)$ for all $x, y : L_1$.
+A functor between two semilattices $L_1, L_2$ consists of a function $f: L_1 \rightarrow L_2$ mapping elements of $L_1$ to elements of $L_2$ that preserves joins in the sense that for $x_1 : L_1, x_2 : L_2$, $f(x_1 \vee x_2) = f(x_1) \vee f(x_2)$.
+From this condition it follows that if $x \leq y$, then $f(x) \leq f(x) \vee f(y) = f(x \vee y) = f(y)$.
 The $\mathrm{Semilattice}$ category has products in the obvious way.
@@
         \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : g_l(x)$: $$g_t(y, z, g_t(x, y, x')) = g_t(x, z, x')$$.
-        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : g_l(x)$ such that $x'_1 \leq x'_2$: $$g_t(x, y, x'_1) \leq g_t(x, y, x'_2)$$.
+        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : g_l(x)$: $$g_t(x, y, x'_1 \vee x'_2) = g_t(x, y, x'_1) \vee g_t(x, y, x'_2)$$.
       \end{itemize}
@@
 
-      Intuitively, the first condition states that transporting a value $x'$ when not changing its dependency $x$ doesn't change $x'$; the second condition states dependency updates aren't path-dependent; and the third states that transports preserve ordering.
+      Intuitively, the first condition states that transporting a value $x'$ when not changing its dependency $x$ doesn't change $x'$; the second condition states dependency updates aren't path-dependent; and the third states transports distribute over joins.
 
@@
             \item $t$ must not be path-dependent: for $k : K, v : V, g_1, g_2, g_3 : d(k, v) \rightarrow V$, we must have $t(k, t(k, v, g_1, g_2), g_2, g_3) = t(k, v, g_1, g_3)$.
-            \item $t$ must preserve orderings: for $k : K, v_1, v_2 : V, g_1, g_2 : d(k, v) \rightarrow V$, such that $j(k, v_1, v_2, g_1) = v_2$, we must have $j(k, t(k, v_1, g_1, g_2), t(k, v_2, g_1, g_2)) = t(k, v_2, g_1, g_2)$.
-            \item For any $k: K, g : K \rightarrow V$, it must be that $\vee_{k, g}$ is commutative, associative, and idempotent.
+            \item $t$ must distribute over $j$: for $k : K, v_1, v_2 : V, g_1, g_2 : d(k, v_1) \cup d(k, v_2) \rightarrow V$, we must have $t_{k, g_1, g_2}(v_1 \vee_{k, g_1} v_2) = t_{k, g_1, g_2}(v_1) \vee_{k, g_2} t_{k, g_1, g_2}(v_2)$.
+            \item For any $k: K, g : K \rightarrow V$, we must have that $\vee_{k, g}$ is commutative, associative, and idempotent.
         \end{itemize}

### 2022-10-01 paper/paper.tex
about to remove parens

@@
             \item $t$ must not grow dependencies: for $k : K, v : V, g_1, g_2 : d(k, v) \rightarrow V$, we must have $d(k, t(k, v, g_1, g_2)) \subseteq d(k, v)$.
-            \item $j$ must not grow dependencies; that is, for any $k : K, v_1 : V, v_2: V, g: (d(k, v_1) \cup d(k, v_2)) \rightarrow V$, we must have $d(k, j(k, v_1, v_2, g)) \subseteq d(k, v_1) \cup d(k, v_2)$.
+            \item $j$ must not grow dependencies; for any $k : K, v_1 : V, v_2: V, g: d(k, v_1) \cup d(k, v_2) \rightarrow V$, we must have $d(k, j(k, v_1, v_2, g)) \subseteq d(k, v_1) \cup d(k, v_2)$.
             \item $t$ must be an identity when dependencies are equal: for $k : K, v : V, g : d(k, v) \rightarrow V$, we must have $t(k, v, g, g) = v$.
-            \item $t$ must not be path-dependent: for $k : K, v : V, g_1, g_2, g_3 : d(k, v) \rightarrow V$, we must have $t(k, t(k, v, g_1, g_2), g_2, g_3) = t(k, v, g_1, g_3)$.
+            \item $t$ must not be path-dependent: for $k : K, v : V, g_1, g_2, g_3: d(k, v) \rightarrow V$, we must have $t_{k, g_2, g_3} \circ t_{k, g_1, g_2} = t_{k, g_1, g_3}$.
             \item $t$ must distribute over $j$: for $k : K, v_1, v_2 : V, g_1, g_2 : d(k, v_1) \cup d(k, v_2) \rightarrow V$, we must have $t_{k, g_1, g_2}(v_1 \vee_{k, g_1} v_2) = t_{k, g_1, g_2}(v_1) \vee_{k, g_2} t_{k, g_1, g_2}(v_2)$.
@@
 
-        $$ (g_1 \vee g_2)(k) := t_{k, g_1, (g_1 \vee g_2)}(g_1(k)) \vee_{k, (g_1 \vee g_2)} t_{k, g_2, (g_1 \vee g_2)}(g_2(k))$$.
+        $$ (g_1 \vee g_2)(k) := t_{k, g_1, g_1 \vee g_2}(g_1(k)) \vee_{k, g_1 \vee g_2} t_{k, g_2, g_1 \vee g_2}(g_2(k))$$.
 
@@
         \begin{itemize}
-          \item $(g_1 \vee g_2)(k) = (g_2 \vee g_1)(k)$
-          \item $((g_1 \vee g_2) \vee g_3)(k) = (g_1 \vee (g_2 \vee g_3))(k)$
-          \item $(g \vee g)(k) = g(k)$
+          \item \emph{Commutative}: $(g_1 \vee g_2)(k) = (g_2 \vee g_1)(k)$
+          \item \emph{Associative}: $((g_1 \vee g_2) \vee g_3)(k) = (g_1 \vee (g_2 \vee g_3))(k)$
+          \item \emph{Idempotent}: $(g \vee g)(k) = g(k)$
         \end{itemize}

### 2022-10-03 paper/paper.tex
l omega semilattice

@@
         We write $x \leq y$ to mean $y = x \vee y$.
+Note that if $x \leq y$ and $y \leq x$, then $x = x \vee y = y$.
+Additionally, $x \leq x$, and if $x \leq y$ and $y \leq z$, then $z = z \vee y = z \vee y \vee x \geq x$.
+Accordingly, $\leq$ forms a partial order.
 
@@
 
-      Note that if $x \leq y$ and $y \leq x$, then $x = x \vee y = y$.
-
       Semilattices form a category $\mathrm{Semilattice}$.
-The objects are semilattices and the morphisms are coproduct-preserving functors between semilattices.
-A functor between two semilattices $L_1, L_2$ consists of a function $f: L_1 \rightarrow L_2$ mapping elements of $L_1$ to elements of $L_2$ that preserves joins in the sense that for $x_1 : L_1, x_2 : L_2$, $f(x_1 \vee x_2) = f(x_1) \vee f(x_2)$.
-From this condition it follows that if $x \leq y$, then $f(x) \leq f(x) \vee f(y) = f(x \vee y) = f(y)$.
+The objects are semilattices and the morphisms are join-preserving functors between semilattices.
+A functor between two semilattices $L_1, L_2$ is, by the definition of functor, a function $f: L_1 \rightarrow L_2$ mapping elements of $L_1$ to elements of $L_2$, such that if $x \leq y$, then $f(x) \leq f(y)$.
+The functor preserves joins if for $x_1 : L_1, x_2 : L_2$, $f(x_1 \vee x_2) = f(x_1) \vee f(x_2)$.
 The $\mathrm{Semilattice}$ category has products in the obvious way.
 
-      A semilattice fibration on a semilattice $L$ is a functor $g : L \rightarrow \mathrm{Semilattice}$, which assigns to each element $x$ of $L$ a semilattice $g(x)$, and assigns to each ordering $x \leq y$ a semilattice functor $g(x \leq y)$ from $g(x)$ to $g(y)$.
+      Note that any join-preserving function $f : L_1 \rightarrow L_2$ is automatically a functor: if $x \leq y$, then $f(x) \leq f(x) \vee f(y) = f(x \vee y) = f(y)$.
 
-      The total space's elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
+      A semilattice fibration on a semilattice $L$ is a functor $g : L \rightarrow \mathrm{Semilattice}$, which assigns to each element $x$ of $L$ a semilattice $g(x)$, and assigns to each ordering $x \leq y$ a semilattice morphism (i.e.
+join-preserving functor) $g(x \leq y)$ from $g(x)$ to $g(y)$.
+
+      The total space $\mathrm{Tot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
 We consider $(x, x') \leq (y, y')$ when $x \leq y$ and $g(x \leq y)(x') \leq y'$.
@@
             \begin{align*}
-              (x, x') \vee (y, y') &= (x \vee y, g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y')) \\
-              &= (y \vee x, g(y \leq y \vee x)(y') \vee g(x \leq y \vee x)(x')) \\
+              (x, x') \vee (y, y') &= \left(x \vee y, g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y')\right) \\
+              &= \left(y \vee x, g(y \leq y \vee x)(y') \vee g(x \leq y \vee x)(x')\right) \\
               &= (y, y') \vee (x, x')
@@
             \begin{align*} &((x, x') \vee (y, y')) \vee (z, z')  \\
-              =& (x \vee y \vee z, g(x \vee y \leq x \vee y \vee z)(g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y')) \vee g(z \leq x \vee y \vee z)(z')) \\
-              =& (x \vee y \vee z, g(x \leq x \vee y \vee z)(x') \vee g(y \leq x \vee y \vee z)(y') \vee g(z \leq x \vee y \vee z)(z') \\
-              =& (x \vee y \vee z, g(x \leq x \vee y \vee z)(x') \vee g(y \vee z \leq x \vee y \vee z)(g(y \leq y \vee z)(y') \vee g(z \leq y \vee z)(z')) \\
+              =& \left(x \vee y \vee z, g(x \vee y \leq x \vee y \vee z)(g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y')) \vee g(z \leq x \vee y \vee z)(z')\right) \\
+              =& \left(x \vee y \vee z, g(x \leq x \vee y \vee z)(x') \vee g(y \leq x \vee y \vee z)(y') \vee g(z \leq x \vee y \vee z)(z')\right) \\
+              =& \left(x \vee y \vee z, g(x \leq x \vee y \vee z)(x') \vee g(y \vee z \leq x \vee y \vee z)(g(y \leq y \vee z)(y') \vee g(z \leq y \vee z)(z'))\right) \\
               =& (x, x') \vee ((y, y') \vee (z, z'))
@@
             \begin{align*}
-              (x, x') \vee (x, x') &= (x \vee x, g(x \leq x \vee x)(x') \vee g(x \leq x \vee x)(x')) \\
-              &= (x, g(x \leq x)(x')) \\
+              (x, x') \vee (x, x') &= \left(x \vee x, g(x \leq x \vee x)(x') \vee g(x \leq x \vee x)(x')\right) \\
+              &= \left(x, g(x \leq x)(x')\right) \\
               &= (x, x')
@@
 
+      Note that the projection $\pi_1 : \mathrm{Tot}(L, g) \rightarrow L$ is a semilattice morphism.
+
     \section{De-categorizing semilattice fibrations}
 
-      To define a semilattice fibration more directly, we can remove the category-theoretic language, achieving a more direct presentation.
+      To define a semilattice fibration more directly, we can remove the category-theoretic language, replacing category-theoretic entities with their definitions.
 
-      A semilattice fibration consists of a function $g_l : L \rightarrow \mathrm{Semilattice}$, assigning to each element of $L$ a semilattice, and a function $g_t : \Pi (x, y : L, x \leq y, x' : g(x)), g(y)$ (a ``transport function''), satisfying:
+      A semilattice fibration consists of a function $g_l : L \rightarrow \mathrm{Semilattice}$, assigning to each element of $L$ a semilattice, and a function $g_t : \Pi (x, y : L, x \leq y) g_l(x) \rightarrow g_l(y)$ (a ``transport function''), satisfying:
 
@@
 
+      We write $\mathrm{SemiFibration}(L)$ for the set of semilattice fibrations on $L$.
+
       This is equivalent to the above category-theoretic definition.
 $g_l$ corresponds to the semilattice fibration functor's action on objects, and $g_t$ to its action on morphisms (itself a functor).
-The first two conditions ensure that $(g_l, g_t)$ is a functor, while the third ensures that $g_t(x, y, \cdot)$ is a semilattice functor.
+The first two conditions ensure that $(g_l, g_t)$ is a functor, while the third ensures that $g_t(x, y, \cdot)$ is a semilattice morphism.
 
@@
 
-      The total space is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : g_l(x)$.
+      The total space $\mathrm{SemiTot}(L, g_l, g_t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : g_l(x)$.
 The join is defined as $(x, x') \vee (y, y') := (x \vee y, g_t(x, x \vee y, x') \vee g_t(y, x \vee y, y'))$.
 This is equivalent to the above category-theoretic definition, proving it is a semilattice.
+
+    \section{Composing semilattice fibrations}
+
+      Semilattice fibrations can be composed iteratively. Composing finitely is straightforward, while composing infinitely requires more work.
+
+      Define $T_0$ to be a semilattice with only a single element, $()$. Now suppose we have:
+
+      \begin{align*}
+        (g_{n, l}, g_{n, t}) &: \mathrm{SemiFibration}(T_{n-1}) & \text{ for natural $n \geq 1$ } \\
+        T_n &:= \mathrm{SemiTot}(T_{n-1}, g_{n, l}, g_{n, t})    & \text{ for natural $n \geq 1$ }
+      \end{align*}
+
+      We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$.
+This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : g_{n, l}(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : T_{n-1}$.
+The joins are defined as follows:
+
+      $$(\overline{x} \vee \overline{y})_n := g_{n, t}(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n})} g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n)$$
+
+      \begin{theorem}
+        $L_{\omega}$ is a semilattice.
+      \end{theorem}
+      \begin{proof}
+
+        To verify that $L_\omega$ is a semilattice, we must verify three properties:
+        \begin{itemize}
+          \item \emph{Commutativity}:
+            \begin{align*}
+              (\overline{x} \vee \overline{y})_n
+              &= g_{n, t}(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n})} g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n) \\
+              &= g_{n, t}(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{g_{n, l}(y_{<n} \vee x_{<n})} g_{n, t}(x_{<n}, y_{<n} \vee x_{<n}, x_n) \\
+              &= (\overline{y} \vee \overline{x})_n
+            \end{align*}
+          \item \emph{Associativity}:
+            \begin{align*}
+              & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
+              =& g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{x} \vee \overline{y})_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})}
+                 g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
+                          g_{n, t}(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n})}
+                          g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \\
+               &\vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(x_{<n}, x_{<n} \vee y_{<n}, x_n)) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, y_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(y_{<n}, y_{<n} \vee z_{<n}, y_n)) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(z_{<n}, y_{<n} \vee z_{<n}, z_n))  \\
+              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
+                          g_{n, t}(y_{<n}, y_{<n} \vee z_{<n}, y_n) \vee_{g_{n, l}(y_{<n} \vee z_{<n})}
+                          g_{n, t}(z_{<n}, y_{<n} \vee z_{<n}, z_n)) \\
+              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{y} \vee \overline{z})_n) \\
+              =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
+            \end{align*}
+          \item \emph{Idempotence}:
+            \begin{align*}
+              (\overline{x} \vee \overline{x})_n &= g_{n, t}(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{g_{n, l}(x_{<n} \vee x_{<n}))} g_{n, t}(x_{<n}, x_{<n} \vee x_{<n}, x_n) \\
+              &= g_{n, t}(x_{<n}, x_{<n}, x_n) \\
+              &= x_n
+            \end{align*}
+        \end{itemize}
+      \end{proof}
 
@@
             \begin{align*}
-              (g_1 \vee g_2)(k) &= t_{k, g_1, (g_1 \vee g_2)}(g_1(k)) \vee_{k, (g_1 \vee g_2)} t_{k, g_2, (g_1 \vee g_2)}(g_2(k)) \\
-              &= t_{k, g_2, (g_2 \vee g_1)}(g_2(k)) \vee_{k, (g_2 \vee g_1)} t_{k, g_1, (g_2 \vee g_1)}(g_1(k)) \\
+              (g_1 \vee g_2)(k) &= t_{k, g_1, g_1 \vee g_2}(g_1(k)) \vee_{k, g_1 \vee g_2} t_{k, g_2, g_1 \vee g_2}(g_2(k)) \\
+              &= t_{k, g_2, g_2 \vee g_1}(g_2(k)) \vee_{k, g_2 \vee g_1} t_{k, g_1, g_2 \vee g_1}(g_1(k)) \\
               &= (g_2 \vee g_1)(k)
@@
               &((g_1 \vee g_2) \vee g_3))(k) \\
-              =& t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}((g_1 \vee g_2)(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
-              =& t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}(t_{k, g_1, (g_1 \vee g_2)}(g_1(k)) \vee_{k, (g_1 \vee g_2)} t_{k, g_2, (g_1 \vee g_2)}(g_2(k))) \\
-               & \vee_{k, (g_1 \vee g_2 \vee g_3)} t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
-              =& t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}(t_{k, g_1, (g_1 \vee g_2)}(g_1(k))) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
-               & t_{k, (g_1 \vee g_2), (g_1 \vee g_2 \vee g_3)}(t_{k, g_2, (g_1 \vee g_2)}(g_2(k))) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
-               & t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
-              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)}
-                 t_{k, g_2, (g_1 \vee g_2 \vee g_3)}(g_2(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)}
-                 t_{k, g_3, (g_1 \vee g_2 \vee g_3)}(g_3(k)) \\
-              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
-               & t_{k, (g_2 \vee g_3), (g_1 \vee g_2 \vee g_3)}(t_{k, g_2, (g_2 \vee g_3)}(g_2(k))) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
-               & t_{k, (g_2 \vee g_3), (g_1 \vee g_2 \vee g_3)}(t_{k, g_3, (g_2 \vee g_3)}(g_3(k))) \\
-              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} \\
-               & t_{k, (g_2 \vee g_3)}(t_{k, g_2, (g_2 \vee g_3)}(g_2(k)) \vee_{k, (g_2 \vee g_3)} t_{k, g_3, (g_2 \vee g_3)}(g_3(k))) \\
-              =& t_{k, g_1, (g_1 \vee g_2 \vee g_3)}(g_1(k)) \vee_{k, (g_1 \vee g_2 \vee g_3)} t_{k, (g_2 \vee g_3)}((g_2 \vee g_3)(k)) \\
+              =& t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}((g_1 \vee g_2)(k)) \vee_{k, g_1 \vee g_2 \vee g_3} t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
+              =& t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}\left(t_{k, g_1, g_1 \vee g_2}(g_1(k)) \vee_{k, g_1 \vee g_2} t_{k, g_2, g_1 \vee g_2}(g_2(k))\right) \\
+               & \vee_{k, g_1 \vee g_2 \vee g_3} t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
+              =& t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}(t_{k, g_1, g_1 \vee g_2}(g_1(k))) \vee_{k, g_1 \vee g_2 \vee g_3} \\
+               & t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}(t_{k, g_2, g_1 \vee g_2}(g_2(k))) \vee_{k, g_1 \vee g_2 \vee g_3} \\
+               & t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
+              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3}
+                 t_{k, g_2, g_1 \vee g_2 \vee g_3}(g_2(k)) \vee_{k, g_1 \vee g_2 \vee g_3}
+                 t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
+              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3} \\
+               & t_{k, g_2 \vee g_3, g_1 \vee g_2 \vee g_3}(t_{k, g_2, g_2 \vee g_3}(g_2(k))) \vee_{k, g_1 \vee g_2 \vee g_3} \\
+               & t_{k, g_2 \vee g_3, g_1 \vee g_2 \vee g_3}(t_{k, g_3, g_2 \vee g_3}(g_3(k))) \\
+              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3} \\
+               & t_{k, g_2 \vee g_3}(t_{k, g_2, g_2 \vee g_3}(g_2(k)) \vee_{k, g_2 \vee g_3} t_{k, g_3, g_2 \vee g_3}(g_3(k))) \\
+              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3} t_{k, g_2 \vee g_3}((g_2 \vee g_3)(k)) \\
               =& (g_1 \vee (g_2 \vee g_3))(k)
@@
             \begin{align*}
-              (g \vee g)(k) &= t_{k, g, (g \vee g)}(g(k)) \vee_{k, (g \vee g)} t_{k, g, (g \vee g)}(g(k)) \\
+              (g \vee g)(k) &= t_{k, g, g \vee g}(g(k)) \vee_{k, g \vee g} t_{k, g, g \vee g}(g(k)) \\
               &= t_{k, g, g}(g(k)) \vee_{k, g} t_{k, g, g}(g(k)) \\

### 2022-10-03 paper/paper.tex
cleanup

@@
 
-      The total space $\mathrm{Tot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
+      The total space $\mathrm{SemiTot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
 We consider $(x, x') \leq (y, y')$ when $x \leq y$ and $g(x \leq y)(x') \leq y'$.
-The total space is a semilattice, as shown:
+Consider two elements of the total space, $(x, x'), (y, y')$.
+Their join is defined as $$(x, x') \vee (y, y') := (x \vee y', g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y'))$$.
 
       \begin{theorem}
-        The total space of a semilattice fibration $g$ is a semilattice.
+        The total space $\mathrm{SemiTot}(L, g)$ of a semilattice $L$ and a semilattice fibration $g$ is a semilattice.
       \end{theorem}
       \begin{proof}
-        Consider two elements of the total space, $(x, x'), (y, y')$.
-Their join is defined as $(x, x') \vee (y, y') := (x \vee y', g(x \leq x \vee y)(x') \vee g(y \leq x \vee y)(y'))$.
         We must show that this join is commutative, associative, and idempotent.
@@
 
-      A semilattice fibration consists of a function $g_l : L \rightarrow \mathrm{Semilattice}$, assigning to each element of $L$ a semilattice, and a function $g_t : \Pi (x, y : L, x \leq y) g_l(x) \rightarrow g_l(y)$ (a ``transport function''), satisfying:
+      A semilattice fibration consists of a function $l : L \rightarrow \mathrm{Semilattice}$, assigning to each element of $L$ a semilattice, and a function $t : \Pi (x, y : L, x \leq y) l(x) \rightarrow l(y)$ (a ``transport function''), satisfying:
 
       \begin{itemize}
-        \item For any $x : L, x' : g_l(x)$: $$g_t(x, x, x') = x'$$.
-        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : g_l(x)$: $$g_t(y, z, g_t(x, y, x')) = g_t(x, z, x')$$.
-        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : g_l(x)$: $$g_t(x, y, x'_1 \vee x'_2) = g_t(x, y, x'_1) \vee g_t(x, y, x'_2)$$.
+        \item For any $x : L, x' : l(x)$: $$t(x, x, x') = x'$$.
+        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : l(x)$: $$t(y, z, t(x, y, x')) = t(x, z, x')$$.
+        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : l(x)$: $$t(x, y, x'_1 \vee x'_2) = t(x, y, x'_1) \vee t(x, y, x'_2)$$.
       \end{itemize}
@@
       This is equivalent to the above category-theoretic definition.
-$g_l$ corresponds to the semilattice fibration functor's action on objects, and $g_t$ to its action on morphisms (itself a functor).
-The first two conditions ensure that $(g_l, g_t)$ is a functor, while the third ensures that $g_t(x, y, \cdot)$ is a semilattice morphism.
+$l$ corresponds to the semilattice fibration functor's action on objects, and $t$ to its action on morphisms (itself a functor).
+The first two conditions ensure that $(l, t)$ is a functor, while the third ensures that $t(x, y, \cdot)$ is a semilattice morphism.
 
@@
 
-      The total space $\mathrm{SemiTot}(L, g_l, g_t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : g_l(x)$.
-The join is defined as $(x, x') \vee (y, y') := (x \vee y, g_t(x, x \vee y, x') \vee g_t(y, x \vee y, y'))$.
+      The total space $\mathrm{SemiTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$.
+The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y, x') \vee t(y, x \vee y, y'))$.
 This is equivalent to the above category-theoretic definition, proving it is a semilattice.
@@
       \begin{align*}
-        (g_{n, l}, g_{n, t}) &: \mathrm{SemiFibration}(T_{n-1}) & \text{ for natural $n \geq 1$ } \\
-        T_n &:= \mathrm{SemiTot}(T_{n-1}, g_{n, l}, g_{n, t})    & \text{ for natural $n \geq 1$ }
+        (l_n, t_n) &: \mathrm{SemiFibration}(T_{n-1}) & \text{ for natural $n \geq 1$ } \\
+        T_n &:= \mathrm{SemiTot}(T_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
       \end{align*}
@@
       We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$.
-This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : g_{n, l}(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : T_{n-1}$.
+This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : T_{n-1}$.
 The joins are defined as follows:
 
-      $$(\overline{x} \vee \overline{y})_n := g_{n, t}(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n})} g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n)$$
+      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)$$
 
@@
               (\overline{x} \vee \overline{y})_n
-              &= g_{n, t}(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n})} g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n) \\
-              &= g_{n, t}(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{g_{n, l}(y_{<n} \vee x_{<n})} g_{n, t}(x_{<n}, y_{<n} \vee x_{<n}, x_n) \\
+              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n) \\
+              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n}, x_n) \\
               &= (\overline{y} \vee \overline{x})_n
@@
               & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
-              =& g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{x} \vee \overline{y})_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})}
-                 g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
-                          g_{n, t}(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n})}
-                          g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \\
-               &\vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(x_{<n}, x_{<n} \vee y_{<n}, x_n)) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, y_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(y_{<n}, y_{<n} \vee z_{<n}, y_n)) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, g_{n, t}(z_{<n}, y_{<n} \vee z_{<n}, z_n))  \\
-              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
-                          g_{n, t}(y_{<n}, y_{<n} \vee z_{<n}, y_n) \vee_{g_{n, l}(y_{<n} \vee z_{<n})}
-                          g_{n, t}(z_{<n}, y_{<n} \vee z_{<n}, z_n)) \\
-              =& g_{n, t}(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{g_{n, l}(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & g_{n, t}(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{y} \vee \overline{z})_n) \\
+              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
+                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
+                          t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
+                          t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \\
+               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(y_{<n}, y_{<n} \vee z_{<n}, y_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(z_{<n}, y_{<n} \vee z_{<n}, z_n))  \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
+                          t_n(y_{<n}, y_{<n} \vee z_{<n}, y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
+                          t_n(z_{<n}, y_{<n} \vee z_{<n}, z_n)) \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{y} \vee \overline{z})_n) \\
               =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
@@
             \begin{align*}
-              (\overline{x} \vee \overline{x})_n &= g_{n, t}(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{g_{n, l}(x_{<n} \vee x_{<n}))} g_{n, t}(x_{<n}, x_{<n} \vee x_{<n}, x_n) \\
-              &= g_{n, t}(x_{<n}, x_{<n}, x_n) \\
+              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n}, x_n) \\
+              &= t_n(x_{<n}, x_{<n}, x_n) \\
               &= x_n

### 2022-10-03 paper/paper.tex
we need more parens
circ for readability
T is now L

@@
       \begin{itemize}
-        \item For any $x : L, x' : l(x)$: $$t(x, x, x') = x'$$.
-        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : l(x)$: $$t(y, z, t(x, y, x')) = t(x, z, x')$$.
-        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : l(x)$: $$t(x, y, x'_1 \vee x'_2) = t(x, y, x'_1) \vee t(x, y, x'_2)$$.
+        \item For any $x : L, x' : l(x)$: $$t(x, x) = \mathsf{id}$$.
+        \item For any $x, y, z : L$ such that $x \leq y \leq z$, and $x' : l(x)$: $$t(y, z) \circ t(x, y) = t(x, z)$$.
+        \item For any $x, y : L$ such that $x \leq y$, and $x'_1, x'_2 : l(x)$: $$t(x, y)(x'_1 \vee x'_2) = t(x, y)(x'_1) \vee t(x, y)(x'_2)$$.
       \end{itemize}
@@
       The total space $\mathrm{SemiTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$.
-The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y, x') \vee t(y, x \vee y, y'))$.
+The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y)(x') \vee t(y, x \vee y)(y'))$.
 This is equivalent to the above category-theoretic definition, proving it is a semilattice.
@@
 
-      Define $T_0$ to be a semilattice with only a single element, $()$. Now suppose we have:
+      Let $L_0$ be an arbitrary semilattice. Now suppose we have:
 
       \begin{align*}
-        (l_n, t_n) &: \mathrm{SemiFibration}(T_{n-1}) & \text{ for natural $n \geq 1$ } \\
-        T_n &:= \mathrm{SemiTot}(T_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
+        (l_n, t_n) &: \mathrm{SemiFibration}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
+        L_n &:= \mathrm{SemiTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
       \end{align*}
@@
       We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$.
-This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : T_{n-1}$.
+This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : L_{n-1}$.
 The joins are defined as follows:
 
-      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)$$
+      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)$$
 
@@
               (\overline{x} \vee \overline{y})_n
-              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n) \\
-              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n}, x_n) \\
+              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n) \\
+              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n})(x_n) \\
               &= (\overline{y} \vee \overline{x})_n
@@
               & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
-              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
-                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
-                          t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
-                          t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \\
-               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(x_{<n}, x_{<n} \vee y_{<n}, x_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(y_{<n}, x_{<n} \vee y_{<n}, y_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, z_n)  \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(y_{<n}, y_{<n} \vee z_{<n}, y_n)) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, t_n(z_{<n}, y_{<n} \vee z_{<n}, z_n))  \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n},
-                          t_n(y_{<n}, y_{<n} \vee z_{<n}, y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
-                          t_n(z_{<n}, y_{<n} \vee z_{<n}, z_n)) \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}, (\overline{y} \vee \overline{z})_n) \\
+              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
+                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
+              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
+                          t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
+                          t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)) \\
+               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
+              =& (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(x_{<n}, x_{<n} \vee y_{<n}))(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, x_{<n} \vee y_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, y_{<n} \vee z_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(z_{<n}, y_{<n} \vee z_{<n}))(z_n)  \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
+                          t_n(y_{<n}, y_{<n} \vee z_{<n})(y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
+                          t_n(z_{<n}, y_{<n} \vee z_{<n})(z_n)) \\
+              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
+               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{y} \vee \overline{z})_n) \\
               =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
@@
             \begin{align*}
-              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n}, x_n) \\
-              &= t_n(x_{<n}, x_{<n}, x_n) \\
+              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \\
+              &= t_n(x_{<n}, x_{<n})(x_n) \\
               &= x_n

### 2022-10-03 paper/paper.tex
more stacking!

@@
 
+      We can further stack to higher ordinals than $\omega$. Let $n_u \geq 1$ be the number of universes. Let $L_{0, 0}$ be an arbitrary semilattice. Assume we have:
+
+      \begin{align*}
+        (l_{u, n}, t_{u, n}) &: \mathrm{SemiFibration}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        L_{u, n} &:= \mathrm{SemiTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        L_{u+1, 0} &:= L_{u, \omega} & \text{ for natural $0 \leq u < n_u - 1$ }
+      \end{align*}
+
+      where $L_{u, \omega}$ is constructed the same way as $L_{\omega}$ above, as the limit of $L_{u, n}$ as $n$ approaches $\infty$.
+Inductively, each $L_{u, n}$ is a semilattice for natural $0 \leq u < n_u, n \geq 0$; this lattice's ordinal corresponds to $u \omega + n$.
+While in theory it may be possible to reach higher ordinals $\omega^2$ and beyond, we consider this beyond the scope of this work.
+
     \section{Semilattice Graphs}

### 2022-10-03 paper/paper.tex
cleanup

@@
 
-      Let $L_0$ be an arbitrary semilattice. Now suppose we have:
+      Suppose we have:
 
       \begin{align*}
+        L_0 &: \mathrm{Semilattice} & \\
         (l_n, t_n) &: \mathrm{SemiFibration}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
@@
 
-      We can further stack to higher ordinals than $\omega$. Let $n_u \geq 1$ be the number of universes. Let $L_{0, 0}$ be an arbitrary semilattice. Assume we have:
+      We can further stack to higher ordinals than $\omega$. Let $n_u \geq 1$ be the number of universes. Assume we have:
 
       \begin{align*}
+        L_{0, 0} &: \mathrm{Semilattice} & \\
         (l_{u, n}, t_{u, n}) &: \mathrm{SemiFibration}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
         L_{u, n} &:= \mathrm{SemiTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
-        L_{u+1, 0} &:= L_{u, \omega} & \text{ for natural $0 \leq u < n_u - 1$ }
+        L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
       \end{align*}

### 2022-10-03 paper/paper.tex
graphs

@@
 
+      In terms of implementation, it is by default infeasible to construct an element of the semilattice $L_{0, \omega}$, as it is infinite.
+To ameliorate this problem, we consider semilattices with a bottom element $\bot$ satisfying $\bot \vee x = x$ for each $x$ in the semilattice.
+We write $\mathrm{Semilattice}_\bot$ for the type of semilattices with $\bot$; $\mathrm{SemiFibration}_\bot(L)$ for the type of bottom-preserving semilattice fibrations on $L : \mathrm{Semilattice}_\bot$; and $\mathrm{SemiTot}_\bot(L, l, t) : \mathrm{Semilattice}_\bot$ for the total space the semilattice $L : \mathrm{Semilattice}_\bot$ and fibration $(l, t) : \mathrm{SemiFibration}_\bot$.
+
+      Using the above definition of semilattice universes, we can now explicitly represent elements of $L_{u, n}$ for which only a finite number of entries are not $\bot$.
+Joins are straightforward since if one element of the joined universe structure is $\bot$, the join equals the other element.
+
+    \section{Semilattice Graphs}
+
         A common problem in networked computing is dependency managament.

### 2022-10-04 src/lattice.rs
new lattice stuff???

@@ -2,133 +2,63 @@ use std::collections::{BTreeSet, BTreeMap};
 
 use serde::{Serialize, de::DeserializeOwned};
 
-/// A graph where each item is a semilattice, which may depend on other lattice max values.
-pub trait LatGraph : Send + Sync {
+// function from D to semilattice
+pub trait SemiL : Send + Sync {
 
-    /// Keys identifying lattice value keys.
-    type K : Eq + Ord + Clone + Serialize + DeserializeOwned + Send + Sync;
+    type Elem : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
 
-    /// The values themselves.
-    type V : Eq + Clone + Serialize + DeserializeOwned + Send + Sync;
+    fn is_elem(&self, e: &Self::Elem) -> bool;
 
-    /// Dependencies of a given key.
-    fn dependencies(&self, key: &Self::K, value: Option<&Self::V>) -> Result<BTreeSet<Self::K>, String>;
+    fn join(&self, a: &Self::Elem, b: &Self::Elem) -> Self::Elem;
 
-    /// Joins two values given dependencies.
-    fn join(&self, key: &Self::K, v1: &Self::V, v2: &Self::V, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
+    // laws:
+    //   assume self.is_elem(a), self.is_elem(b), self.is_elem(c)
+    //   self.join(a, a) = a
+    //   self.join(a, b) = self.join(b, a)
+    //   self.join(a, self.join(b, c)) = self.join(self.join(a, b), c)
 
-    /// The default element of the lattice.
-    fn autofill(&self, key: &Self::K, deps: &BTreeMap<Self::K, Self::V>) -> Result<Self::V, String>;
-}
+    fn bottom(&self) -> Self::Elem;
 
-pub struct SerializeLatGraph<'a, L: LatGraph>(&'a L);
-
-impl<'a, L: LatGraph> SerializeLatGraph<'a, L> {
-    pub fn new(l: &'a L) -> Self {
-        SerializeLatGraph(l)
-    }
-}
-
-impl<'a, L: LatGraph + 'static> LatGraph for SerializeLatGraph<'a, L> {
-
-    type K = Vec<u8>;
-
-    type V = Vec<u8>;
-
-    fn dependencies(&self, key: &Vec<u8>, value: Option<&Vec<u8>>) -> Result<BTreeSet<Vec<u8>>, String> {
[…]

### 2022-10-04 src/main.rs

@@ -1,5 +1,4 @@
 mod lattice;
-mod mapping;
 
 fn main() {
     println!("Hello, world!");

### 2022-10-04 src/lattice.rs
either semil

@@ -1,6 +1,6 @@
 use std::collections::{BTreeSet, BTreeMap};
 
-use serde::{Serialize, de::DeserializeOwned};
+use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
 // function from D to semilattice
 pub trait SemiL : Send + Sync {
@@ -26,7 +26,39 @@ pub trait SemiL : Send + Sync {
     fn leq(&self, a: &Self::Elem, b: &Self::Elem) -> bool {
         self.join(a, b) == *b
     }
+}
 
+#[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Serialize, Deserialize)]
+enum Either<A, B> {
+    Left(A),
+    Right(B),
+}
+
+impl<L1 : SemiL, L2: SemiL> SemiL for Either<L1, L2> {
+    type Elem = Either<L1::Elem, L2::Elem>;
+
+    fn is_elem(&self, e: &Self::Elem) -> bool {
+        match (self, e) {
+            (Either::Left(l1), Either::Left(e)) => l1.is_elem(e),
+            (Either::Right(l2), Either::Right(e)) => l2.is_elem(e),
+            _ => false,
+        }
+    }
+
+    fn join(&self, a: &Self::Elem, b: &Self::Elem) -> Self::Elem {
+        match (self, a, b) {
+            (Either::Left(l1), Either::Left(a), Either::Left(b)) => Either::Left(l1.join(a, b)),
+            (Either::Right(l2), Either::Right(a), Either::Right(b)) => Either::Right(l2.join(a, b)),
+            _ => panic!("join: not an element"),
+        }
+    }
+
+    fn bottom(&self) -> Self::Elem {
+        match self {
+            Either::Left(l1) => Either::Left(l1.bottom()),
+            Either::Right(l2) => Either::Right(l2.bottom()),
+        }
+    }
 }
 
 // function from D to semilattice fibration
@@ -62,3 +94,4 @@ pub trait SemiLUniverse<L: SemiL> : Send + Sync + Sized {
 pub trait SemiLUniverseSpec<L: SemiL, U: SemiLUniverse<L, Spec = Self>> : SemiL + Sized {
     fn elem_at(&self, lat: &L, i: usize) -> Result<<<U::Fib as SemiLFibration<Self>>::Lat as SemiL>::Elem, String>;
 }
+

### 2022-10-04 src/lattice.rs
keys!

@@ -84,14 +84,17 @@ pub trait SemiLFibration<L : SemiL> : Send + Sync {
 }
 
 pub trait SemiLUniverse<L: SemiL> : Send + Sync + Sized {
+    type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
+
     type Spec : SemiLUniverseSpec<L, Self>;
 
     type Fib : SemiLFibration<Self::Spec>;
 
-    fn fibration(&self, lat: &L, spec: &Self::Spec) -> Self::Fib;
+    fn fibration(&self, lat: &L, key: &Self::Key, spec: &Self::Spec) -> Self::Fib;
 }
 
 pub trait SemiLUniverseSpec<L: SemiL, U: SemiLUniverse<L, Spec = Self>> : SemiL + Sized {
-    fn elem_at(&self, lat: &L, i: usize) -> Result<<<U::Fib as SemiLFibration<Self>>::Lat as SemiL>::Elem, String>;
+
+    fn elem_at(&self, lat: &L, key: &U::Key) -> Result<<<U::Fib as SemiLFibration<Self>>::Lat as SemiL>::Elem, String>;
 }
 

### 2022-10-04 src/lattice.rs
multiverses!

@@ -95,6 +95,23 @@ pub trait SemiLUniverse<L: SemiL> : Send + Sync + Sized {
 
 pub trait SemiLUniverseSpec<L: SemiL, U: SemiLUniverse<L, Spec = Self>> : SemiL + Sized {
 
-    fn elem_at(&self, lat: &L, key: &U::Key) -> Result<<<U::Fib as SemiLFibration<Self>>::Lat as SemiL>::Elem, String>;
+    fn base_elem(&self, lat: &L) -> L::Elem;
+
+    fn elem_at(&self, lat: &L, key: &U::Key) -> <<U::Fib as SemiLFibration<Self>>::Lat as SemiL>::Elem;
 }
 
+
+pub trait SemiLMultiverse<L: SemiL, U: SemiLUniverse<Self::MultiSpec>> : Send + Sync + Sized {
+    type MultiKey : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
+
+    type MultiSpec : SemiLMultiverseSpec<L, U, Self>;
+
+    fn universe(&self, lat: &L, key: &Self::MultiKey, spec: &Self::MultiSpec) -> U;
+}
+
+pub trait SemiLMultiverseSpec<L: SemiL, U: SemiLUniverse<Self>, M: SemiLMultiverse<L, U, MultiSpec = Self>> : SemiL + Sized {
+
+    fn base_elem(&self, lat: &L) -> L::Elem;
+
+    fn universe_at(&self, lat: &L, mkey: &M::MultiKey) -> U;
+}

### 2022-10-04 src/lattice.rs
total space

@@ -67,7 +67,7 @@ pub trait SemiLFibration<L : SemiL> : Send + Sync {
 
     fn lattice(&self, lat: &L, x: &L::Elem) -> Self::Lat;
 
-    fn transport(&self, lat: &L, x: &L::Elem, y: &L::Elem, a: <Self::Lat as SemiL>::Elem) -> <Self::Lat as SemiL>::Elem;
+    fn transport(&self, lat: &L, x: &L::Elem, y: &L::Elem, a: &<Self::Lat as SemiL>::Elem) -> <Self::Lat as SemiL>::Elem;
 
     // laws:
     //   assume lat.is_elem(x), lat.is_elem(y), lat.is_elem(z)
@@ -83,6 +83,33 @@ pub trait SemiLFibration<L : SemiL> : Send + Sync {
     //         self.transport(lat, x, y, a), self.transport(lat, x, y, b))
 }
 
+struct SemiLTot<'a, L: SemiL, F: SemiLFibration<L>> {
+    lat: &'a L,
+    fib: &'a F,
+}
+
+impl<'a, L: SemiL, F: SemiLFibration<L>> SemiL for SemiLTot<'a, L, F> {
+    type Elem = (L::Elem, <F::Lat as SemiL>::Elem);
+
+    fn is_elem(&self, e: &Self::Elem) -> bool {
+        self.lat.is_elem(&e.0) && self.fib.lattice(self.lat, &e.0).is_elem(&e.1)
+    }
+
+    fn join(&self, a: &Self::Elem, b: &Self::Elem) -> Self::Elem {
+        let c0 = self.lat.join(&a.0, &b.0);
+        let ta1 = self.fib.transport(self.lat, &a.0, &c0, &a.1);
+        let tb1 = self.fib.transport(self.lat, &b.0, &c0, &b.1);
+        let lat = self.fib.lattice(self.lat, &c0);
+        (c0, lat.join(&ta1, &tb1))
+    }
+
+    fn bottom(&self) -> Self::Elem {
+        let bot = self.lat.bottom();
+        let lat_bot = self.fib.lattice(self.lat, &bot).bottom();
+        (bot, lat_bot)
+    }
+}
+
 pub trait SemiLUniverse<L: SemiL> : Send + Sync + Sized {
     type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
 
[…]

### 2022-10-04 paper/paper.tex
remove autofills, specify graphs

@@
 To ameliorate this problem, we consider semilattices with a bottom element $\bot$ satisfying $\bot \vee x = x$ for each $x$ in the semilattice.
-We write $\mathrm{Semilattice}_\bot$ for the type of semilattices with $\bot$; $\mathrm{SemiFibration}_\bot(L)$ for the type of bottom-preserving semilattice fibrations on $L : \mathrm{Semilattice}_\bot$; and $\mathrm{SemiTot}_\bot(L, l, t) : \mathrm{Semilattice}_\bot$ for the total space the semilattice $L : \mathrm{Semilattice}_\bot$ and fibration $(l, t) : \mathrm{SemiFibration}_\bot$.
+We write $\mathrm{Semilattice}_\bot$ for the type of semilattices with $\bot$; $\mathrm{SemiLFibration}_\bot(L)$ for the type of bottom-preserving semilattice fibrations on $L : \mathrm{Semilattice}_\bot$; and $\mathrm{SemiLTot}_\bot(L, l, t) : \mathrm{Semilattice}_\bot$ for the total space the semilattice $L : \mathrm{Semilattice}_\bot$ and fibration $(l, t) : \mathrm{SemiLFibration}_\bot$.
 
@@
 Joins are straightforward since if one element of the joined universe structure is $\bot$, the join equals the other element.
+The bottoms of the total space and $\omega$ total spaces are trivial, containing only $\bot$.
+
+      Let $K$ be a countable, well-ordered set of keys whose ordinal is no more than $\omega^2$.
+Let $V$ be a set of values.
+Since keys correspond with ordinals in $\omega^2$, we can write any key as $u\omega + n$; let $u(k)$ and $n(k)$ be $u$ and $n$ in this term respectively.
+We define a \emph{semilattice graph} to consist of:
+      \begin{align*}
+        L_{0, 0} &:= \{ \bot \} \\
+        (l_k, t_k) &: \mathrm{SemiLFibration}_\bot(L_{u(k), n(k)}) & \text{ for each $k \in K$ } \\
+        L_{u, n} &:= \mathrm{SemiLTot}_\bot(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
+      \end{align*}
+
+      Note that each key corresponds to a semilattice fibration from a multiverse containing lower keys.
 
@@
 
-    \section{Autofills}
-
-        Mappings may be infinite in general, and even finite ones contain a large number of keys.
-It is therefore computationally infeasible to store a mapping as a set of $K \times V$ pairs.
-Accordingly, we define \emph{autofills} to fill in lattice values automatically from a set of other values.
-
-        % Define $d^*(k, g) := \{k\} \cup \bigcup_{k' \in d(k, g(k))} d^*(k', g)$, the transitive closure of $d$, where $g$ is a mapping.
-The recursion is well-founded (and $d^*(k, g)$ is finite) because $d$ is well-founded.
-
-        % Note that $d^*(k, g_1 \vee g_2) \subseteq d^*(k, g_1) \cup d^*(k, g_2)$.
-
-        An \emph{autofill} for a semilattice graph consists of a dependency function $\hat{d}: K \rightarrow 2^V$ and an autofill function $a : \forall (k:K), \hat{d}(k) \rightarrow V$ that satisfy:
-
-        \begin{itemize}
-          \item
-            For each $k \in K, v \in V$, $\hat{d}(k) \subseteq d(k, v)$.
-          \item
-            $a(k, g_1) \vee_{k, g_1 \vee g_2} a(k, g_2) = a(k, g_1 \vee g_2)$; that is, the join of two autofill values on two mappings must be an autofill value on the joined mapping.
-        \end{itemize}
-
-        where we write $a(k, g) := a(k, g|_{\hat{d}(k)})$ for brevity.
-
-        A mapping $g : K \rightarrow V$ is an \emph{autofill mapping} if:
-
-        \begin{itemize}
-          \item $g(k) = a(k, g)$ for all but a finite number of $k$ values.
-          \item $g(k) \vee_{k, g} a(k, g) = g(k)$ for all $k \in K$.
-        \end{itemize}
-
-        An autofill mapping can be represented as a finite set of $K \times V$ pairs; the remaining values are filled in automatically.
-
-        \begin{theorem}
-          Let $g_1, g_2 : K \rightarrow V$ be autofill mappings. Then $g_1 \vee g_2$ is an autofill mapping.
-        \end{theorem}
-
-        \begin{proof}
-
-          Whenever $g_1(k) = a(k, g_1)$ and $g_2(k) = a(k, g_2)$ (which is true for all but a finite number of $k$ values):
-           $$(g_1 \vee g_2)(k) = g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) = a(k, g_1) \vee_{k, g_1 \vee g_2} a(k, g_2) = a(k, g_1 \vee g_2)$$
-
-          Additionally, for each $k \in K$:
-          \begin{align*}
-            (g_1 \vee g_2)(k) \vee_{k, (g_1 \vee g_2)} a(k, g) 
-            &= (g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k)) \vee_{k, (g_1 \vee g_2)} a(k, g) \\
-            &= g_1(k) \vee_{k, (g_1 \vee g_2)} (g_2(k) \vee_{k, (g_1 \vee g_2)} a(k, g)) \\
-            &= g_1(k) \vee_{k, (g_1 \vee g_2)} g_2(k) \\
-            &= (g_1 \vee g_2)(k)
-          \end{align*}
-        \end{proof}
-
-        For an autofill mapping $g$, we can define the \emph{concrete dependencies} of a key $k \in K$, which are a set of non-autofill keys of $g$ that are sufficient to derive $k$'s dependencies:
-
-        $$c(k, g) := \bigcup_{k' \in d(k, g(k))} \begin{cases}
-          \{k'\} & \text{if } g(k') \neq a(k, g) \\
-          c(k', g) & \text{otherwise}
-        \end{cases}  $$
-
-        Merging two mappings can be done recursively.
-Given mappings $g_1$ and $g_2$ and a key $k$, we compute the values $g_1(k), g_2(k)$ and their dependency sets $d(k, g_1(k)), d(k, g_2(k))$.
-We recursively merge $g_1$ and $g_2$ on each key in either dependency set, caching intermediate values to avoid repetated recursive calls.
-Finally, we compute the join and cache it in the database.
-
-        We may consider merging a mapping with a singleton mapping, which specifies the value of a single key and otherwise contains only default values.
-In this case, merging will change this value and all its direct or indirect dependencies.
-Merging with a singleton is similar to a ``database insert'', updating the value of a given key.
-
 \end{document}

### 2022-10-04 paper/paper.tex
dependencies

@@
         L_{0, 0} &:= \{ \bot \} \\
-        (l_k, t_k) &: \mathrm{SemiLFibration}_\bot(L_{u(k), n(k)}) & \text{ for each $k \in K$ } \\
+        (l_k, t_k) &: \mathrm{SemiLFibration}_\bot(L_{u(k), n(k)-1}) & \text{ for each $k \in K$ } \\
+        l_k(x) \subseteq V &\text{ for each $k \in K$ and $x \in L_{u(k), n(k)}$} \\
         L_{u, n} &:= \mathrm{SemiLTot}_\bot(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
@@
       Note that each key corresponds to a semilattice fibration from a multiverse containing lower keys.
+We can re-interpret each element of the lattice $L_{u(k), n(k)}$ as mappings from keys less no higher than $k$ (written as $K_{\leq k}$) to $V$, which doesn't affect the semilattice structure.
+Accordingly, $l_k$ and $t_k$ are re-interpreted as taking mappings as arguments.
+
+      It is somewhat unwieldy to define $l_k$ and $t_k$ in implementation, since they depend on a possibly-infinite mapping.
+Accordingly, we define a notion of \emph{dependencies} of a function.
+For a key $k : K$ sets $X, Y$, a function $f : X \times (K_{<k} \rightarrow V) \rightarrow S$ has dependencies according to a function $d : X \rightarrow 2^{K_{<k}}$ when for any $x : X, m, m' : K_{<k} \rightarrow V$, if $m(k') = m'(k')$ for each $k' : d(x)$, we have $f(x, m) = f(x, m')$.
+
+      Now we apply this to $l_k, t_k$.
+Let $d_k^\star : 2^{K_{<k}}$ be a set of lattice dependencies, and $d_{k}(v) : 2^{K_{<k}}$ be a set of value dependencies for $v \in V$.
+We say $(l_k, d_k)$ have dependencies according to $d_{k, l}, d_{k, v}$ if:
+
+      \begin{itemize}
+        \item $\mathrm{Elems}(l_k(m))$ only depends on $m(k')$ for each $k' \in d_^\star$.
+        \item $\bot_{l_k(m})$ only depends on $m(k')$ for each $k' \in d_k^\star$.
+        \item $t_k(m, m')(v)$ only depends on $m(k')$ and $m'(k')$ for each $k' \in d_k^\star \cup d_k(v)$.
+        \item $x \vee_{l_k(m)} y$ only depends on $m(k')$ for each $k' \in d_k^\star \cup d_k(x) \cup d_k(y)$.
+      \end{itemize}
 

### 2022-10-04 src/lattice.rs
now have latgraph

@@ -1,146 +1,24 @@
 use std::collections::{BTreeSet, BTreeMap};
 
-use serde::{Serialize, Deserialize, de::DeserializeOwned};
+use serde::{Serialize, de::DeserializeOwned};
 
-// function from D to semilattice
-pub trait SemiL : Send + Sync {
+pub trait LatGraph : Send + Sync {
 
-    type Elem : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
-
-    fn is_elem(&self, e: &Self::Elem) -> bool;
-
-    fn join(&self, a: &Self::Elem, b: &Self::Elem) -> Self::Elem;
-
-    // laws:
-    //   assume self.is_elem(a), self.is_elem(b), self.is_elem(c)
-    //   self.join(a, a) = a
-    //   self.join(a, b) = self.join(b, a)
-    //   self.join(a, self.join(b, c)) = self.join(self.join(a, b), c)
-
-    fn bottom(&self) -> Self::Elem;
-
-    // laws:
-    //   self.is_elem(self.bottom(d))
-    //   self.join(a, self.bottom(d)) = a
-    
-    fn leq(&self, a: &Self::Elem, b: &Self::Elem) -> bool {
-        self.join(a, b) == *b
-    }
-}
-
-#[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Serialize, Deserialize)]
-enum Either<A, B> {
-    Left(A),
-    Right(B),
-}
-
-impl<L1 : SemiL, L2: SemiL> SemiL for Either<L1, L2> {
-    type Elem = Either<L1::Elem, L2::Elem>;
-
-    fn is_elem(&self, e: &Self::Elem) -> bool {
-        match (self, e) {
-            (Either::Left(l1), Either::Left(e)) => l1.is_elem(e),
-            (Either::Right(l2), Either::Right(e)) => l2.is_elem(e),
-            _ => false,
-        }
-    }
-
-    fn join(&self, a: &Self::Elem, b: &Self::Elem) -> Self::Elem {
-        match (self, a, b) {
-            (Either::Left(l1), Either::Left(a), Either::Left(b)) => Either::Left(l1.join(a, b)),
-            (Either::Right(l2), Either::Right(a), Either::Right(b)) => Either::Right(l2.join(a, b)),
-            _ => panic!("join: not an element"),
[…]

### 2022-10-04 paper/paper.tex
fix typo
serialize and enum lat graphs

@@
       It is somewhat unwieldy to define $l_k$ and $t_k$ in implementation, since they depend on a possibly-infinite mapping.
-Accordingly, we define a notion of \emph{dependencies} of a function.
-For a key $k : K$ sets $X, Y$, a function $f : X \times (K_{<k} \rightarrow V) \rightarrow S$ has dependencies according to a function $d : X \rightarrow 2^{K_{<k}}$ when for any $x : X, m, m' : K_{<k} \rightarrow V$, if $m(k') = m'(k')$ for each $k' : d(x)$, we have $f(x, m) = f(x, m')$.
-
-      Now we apply this to $l_k, t_k$.
+Accordingly, we assume that these operations only depend on a defined set of values.
 Let $d_k^\star : 2^{K_{<k}}$ be a set of lattice dependencies, and $d_{k}(v) : 2^{K_{<k}}$ be a set of value dependencies for $v \in V$.
+Both dependency sets must only include keys less than $k$.
 We say $(l_k, d_k)$ have dependencies according to $d_{k, l}, d_{k, v}$ if:
@@
       \begin{itemize}
-        \item $\mathrm{Elems}(l_k(m))$ only depends on $m(k')$ for each $k' \in d_^\star$.
+        \item $\mathrm{Elems}(l_k(m))$ only depends on $m(k')$ for each $k' \in d_k^\star$.
         \item $\bot_{l_k(m})$ only depends on $m(k')$ for each $k' \in d_k^\star$.
@@
         \item $x \vee_{l_k(m)} y$ only depends on $m(k')$ for each $k' \in d_k^\star \cup d_k(x) \cup d_k(y)$.
+      \end{itemize}
+
+      To ensure dependency sets don't grow unexpectedly, we require:
+      \begin{itemize}
+        \item $d_k(t_k(m, m')(v)) \subseteq d_k(v)$.
+        \item $d_k(x \vee_{l_k(m)} y) \subseteq d_k(x) \cup d_k(y)$.
       \end{itemize}

### 2022-10-04 src/lattice.rs

@@ -8,17 +8,147 @@ pub trait LatGraph : Send + Sync {
 
     type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
 
-    fn lat_deps(&self, key: &Self::Key) -> BTreeSet<Self::Key>;
+    fn lat_deps(&self, key: &Self::Key) -> Result<BTreeSet<Self::Key>, String>;
 
-    fn value_deps(&self, key: &Self::Key, value: &Self::Value) -> BTreeSet<Self::Key>;
+    fn value_deps(&self, key: &Self::Key, value: &Self::Value) -> Result<BTreeSet<Self::Key>, String>;
 
-    fn is_elem(&self, key: &Self::Key, value: &Self::Value, deps: BTreeMap<Self::Key, Self::Value>) -> bool;
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value, deps: &BTreeMap<Self::Key, Self::Value>) -> Result<(), String>;
 
-    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value, deps: BTreeMap<Self::Key, Self::Value>) -> Self::Value;
+    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value, deps: &BTreeMap<Self::Key, Self::Value>) -> Result<Self::Value, String>;
 
-    fn bottom(&self, key: &Self::Key, deps: BTreeMap<Self::Key, Self::Value>) -> Self::Value;
+    fn bottom(&self, key: &Self::Key, deps: &BTreeMap<Self::Key, Self::Value>) -> Result<Self::Value, String>;
 
-    fn transport(&self, key: &Self::Key, value: &Self::Value, old_deps: BTreeMap<Self::Key, Self::Value>, new_deps: BTreeMap<Self::Key, Self::Value>) -> Self::Value;
+    fn transport(&self, key: &Self::Key, value: &Self::Value, old_deps: &BTreeMap<Self::Key, Self::Value>, new_deps: &BTreeMap<Self::Key, Self::Value>) -> Result<Self::Value, String>;
 
 }
 
+pub struct SerializeLatGraph<'a, L: LatGraph>(&'a L);
+
+impl<'a, L: LatGraph> SerializeLatGraph<'a, L> {
+    pub fn new(l: &'a L) -> Self {
+        SerializeLatGraph(l)
+    }
+}
+
[…]

### 2022-10-04 src/lattice.rs
total graph

@@ -1,6 +1,6 @@
 use std::collections::{BTreeSet, BTreeMap};
 
-use serde::{Serialize, de::DeserializeOwned};
+use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
 pub trait LatGraph : Send + Sync {
 
@@ -150,5 +150,104 @@ where L::Key: IsEnum, L::Value: IsEnum {
     fn transport(&self, key: &L::Key, value: &L::Value, old_deps: &BTreeMap<L::Key, L::Value>, new_deps: &BTreeMap<L::Key, L::Value>) -> Result<L::Value, String> {
         Self::check_branch_eq(key.get_branch(), value.get_branch())?;
         self.lat_by_branch(key.get_branch())?.transport(key, value, old_deps, new_deps)
+    }
+}
+
+#[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Debug, Serialize, Deserialize)]
+enum Either<A, B> {
+    Left(A),
+    Right(B),
+}
+
+pub trait DepLatGraph<L: LatGraph> : Send + Sync {
+
+    type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
+
+    type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
+
+    fn lat_deps(&self, key: &Self::Key) -> Result<BTreeSet<Either<L::Key, Self::Key>>, String>;
+
+    fn value_deps(&self, key: &Self::Key, value: &Self::Value) -> Result<BTreeSet<Either<L::Key, Self::Key>>, String>;
+
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value, deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<(), String>;
+
+    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value, deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<Self::Value, String>;
+
+    fn bottom(&self, key: &Self::Key, deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<Self::Value, String>;
+
[…]

### 2022-10-04 src/lattice.rs
ldeps rdeps

@@ -154,7 +154,7 @@ where L::Key: IsEnum, L::Value: IsEnum {
 }
 
 #[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Debug, Serialize, Deserialize)]
-enum Either<A, B> {
+pub enum Either<A, B> {
     Left(A),
     Right(B),
 }
@@ -169,13 +169,13 @@ pub trait DepLatGraph<L: LatGraph> : Send + Sync {
 
     fn value_deps(&self, key: &Self::Key, value: &Self::Value) -> Result<BTreeSet<Either<L::Key, Self::Key>>, String>;
 
-    fn check_elem(&self, key: &Self::Key, value: &Self::Value, deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<(), String>;
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value, ldeps: &BTreeMap<L::Key, L::Value>, rdeps: &BTreeMap<Self::Key, Self::Value>) -> Result<(), String>;
 
-    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value, deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<Self::Value, String>;
+    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value, ldeps: &BTreeMap<L::Key, L::Value>, rdeps: &BTreeMap<Self::Key, Self::Value>) -> Result<Self::Value, String>;
 
-    fn bottom(&self, key: &Self::Key, deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<Self::Value, String>;
+    fn bottom(&self, key: &Self::Key, ldeps: &BTreeMap<L::Key, L::Value>, rdeps: &BTreeMap<Self::Key, Self::Value>) -> Result<Self::Value, String>;
 
-    fn transport(&self, key: &Self::Key, value: &Self::Value, old_deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>, new_deps: &BTreeMap<Either<L::Key, Self::Key>, Either<L::Value, Self::Value>>) -> Result<Self::Value, String>;
[…]

### 2022-10-04 src/lattice.rs
latresult

@@ -1,261 +1,313 @@
 use std::collections::{BTreeSet, BTreeMap};
+use std::cmp::Ordering;
+use std::sync::Arc;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
-pub trait LatGraph : Send + Sync {
+enum LatResult<G: LatGraph, T> {
+    Done(Result<T, String>),
+    Lookup(G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatResult<G, T>>)
+}
+
+impl<G: LatGraph + 'static, T: 'static> LatResult<G, T> {
+    fn and_then<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatResult<G, U>) -> LatResult<G, U> {
+        match self {
+            LatResult::Done(Ok(t)) => f(t),
+            LatResult::Done(Err(e)) => LatResult::Done(Err(e)),
+            LatResult::Lookup(key, g) => LatResult::Lookup(key, Box::new(move |lat, value| g(lat, value).and_then(f)))
+        }
+    }
+    fn map<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> U) -> LatResult<G, U> {
+        self.and_then(move |t| LatResult::Done(Ok(f(t))))
+    }
+}
+
+pub trait LatGraph : Send + Sync + Sized {
 
     type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
 
     type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
 
-    fn lat_deps(&self, key: &Self::Key) -> Result<BTreeSet<Self::Key>, String>;
+    fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {
+        key1.partial_cmp(key2).ok_or("Keys are not comparable".to_string())
+    }
 
-    fn value_deps(&self, key: &Self::Key, value: &Self::Value) -> Result<BTreeSet<Self::Key>, String>;
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatResult<Self, ()>;
 
-    fn check_elem(&self, key: &Self::Key, value: &Self::Value, deps: &BTreeMap<Self::Key, Self::Value>) -> Result<(), String>;
[…]

### 2022-10-04 src/lattice.rs
add dep

@@ -4,20 +4,20 @@ use std::sync::Arc;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
-enum LatResult<G: LatGraph, T> {
+enum LatResult<G: LatGraph, D, T> {
     Done(Result<T, String>),
-    Lookup(G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatResult<G, T>>)
+    Lookup(D, G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatResult<G, D, T>>)
 }
 
-impl<G: LatGraph + 'static, T: 'static> LatResult<G, T> {
-    fn and_then<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatResult<G, U>) -> LatResult<G, U> {
+impl<G: LatGraph + 'static, D: 'static, T: 'static> LatResult<G, D, T> {
+    fn and_then<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatResult<G, D, U>) -> LatResult<G, D, U> {
         match self {
             LatResult::Done(Ok(t)) => f(t),
             LatResult::Done(Err(e)) => LatResult::Done(Err(e)),
-            LatResult::Lookup(key, g) => LatResult::Lookup(key, Box::new(move |lat, value| g(lat, value).and_then(f)))
+            LatResult::Lookup(key, dep, g) => LatResult::Lookup(key, dep, Box::new(move |lat, value| g(lat, value).and_then(f)))
         }
     }
-    fn map<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> U) -> LatResult<G, U> {
+    fn map<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> U) -> LatResult<G, D, U> {
         self.and_then(move |t| LatResult::Done(Ok(f(t))))
     }
 }
@@ -32,13 +32,13 @@ pub trait LatGraph : Send + Sync + Sized {
         key1.partial_cmp(key2).ok_or("Keys are not comparable".to_string())
     }
 
-    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatResult<Self, ()>;
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatResult<Self, (), ()>;
 
-    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> LatResult<Self, Self::Value>;
[…]

### 2022-10-04 src/crypto.rs
crypto module

@@ -1,0 +1,195 @@
+//! Cryptographic types and operations.
+
+use std::cmp::{Ord, Ordering, PartialOrd};
+use std::convert::TryInto;
+use std::default::Default;
+use std::marker::PhantomData;
+
+// use ed25519_dalek::{Signer, Verifier};
+use rand::prelude::*;
+use serde::{Deserialize, Serialize};
+use sha2::{Digest, Sha256};
+
+/* QUESTION: tell me more about how much of this is table stakes
+for a generic implementation, how much of it is reusable with
+cosmos, and how much of it is somewhere in between? */
+
+/// A SHA256 hash code.
+pub type HashCode = [u8; 32];
+
+// pub type PublicKey = ed25519_dalek::PublicKey;
+// 
+// pub type Keypair = ed25519_dalek::Keypair;
+// 
+// pub type Signature = ed25519_dalek::Signature;
+
+/// A hash code that is tagged as being a hash code of a particular serializable type.
+#[derive(Serialize, Deserialize, Debug)]
+pub struct Hash<T> {
+    /// The hash code.
+    pub code: HashCode,
+    /// Phantom data for the type `T`.
+    pub(crate) phantom: std::marker::PhantomData<fn() -> T>,
+}
+
+impl<T> Default for Hash<T> {
+    fn default() -> Self {
+        Self {
+            code: HashCode::default(),
+            phantom: PhantomData,
+        }
+    }
+}
+
+impl<T> Clone for Hash<T> {
+    fn clone(&self) -> Self {
+        Self {
+            code: self.code,
+            phantom: PhantomData,
+        }
+    }
+}
+
+impl<T> Copy for Hash<T> {}
+
+impl<T> PartialEq for Hash<T> {
+    fn eq(&self, other: &Self) -> bool {
+        self.code == other.code
+    }
+}
+
+impl<T> Eq for Hash<T> {}
+
+impl<T> PartialOrd for Hash<T> {
+    fn partial_cmp(&self, other: &Self) -> Option<Ordering> {
+        self.code.partial_cmp(&other.code)
+    }
+}
+
+impl<T> Ord for Hash<T> {
+    fn cmp(&self, other: &Self) -> Ordering {
+        self.code.cmp(&other.code)
+    }
+}
+
+impl<T> Hash<T> {
+    pub fn to_hex(&self) -> String {
+        hex::encode(self.code)
+    }
+    pub fn from_hex(hexrep: &str) -> Result<Self, hex::FromHexError> {
[…]

### 2022-10-04 src/lattice.rs

@@ -4,20 +4,20 @@ use std::sync::Arc;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
-enum LatResult<G: LatGraph, D, T> {
+pub enum LatResult<G: LatGraph, D, T> {
     Done(Result<T, String>),
     Lookup(D, G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatResult<G, D, T>>)
 }
 
 impl<G: LatGraph + 'static, D: 'static, T: 'static> LatResult<G, D, T> {
-    fn and_then<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatResult<G, D, U>) -> LatResult<G, D, U> {
+    pub fn and_then<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatResult<G, D, U>) -> LatResult<G, D, U> {
         match self {
             LatResult::Done(Ok(t)) => f(t),
             LatResult::Done(Err(e)) => LatResult::Done(Err(e)),
             LatResult::Lookup(key, dep, g) => LatResult::Lookup(key, dep, Box::new(move |lat, value| g(lat, value).and_then(f)))
         }
     }
-    fn map<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> U) -> LatResult<G, D, U> {
+    pub fn map<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> U) -> LatResult<G, D, U> {
         self.and_then(move |t| LatResult::Done(Ok(f(t))))
     }
 }

### 2022-10-04 src/main.rs

@@ -1,3 +1,4 @@
+mod crypto;
 mod lattice;
 
 fn main() {

### 2022-10-04 src/mapping.rs

@@ -14,16 +14,6 @@ pub trait LatMapping<L: LatGraph> : Send + Sync {
 }
 
 #[async_trait]
-pub trait LatMappingExt<L: LatGraph> : LatMapping<L> {
-
-    async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String>;
-
-    async fn join(self: Arc<Self>, key: L::K, v1: L::V, v2: L::V) -> Result<L::V, String>;
-
-    async fn autofill(self: Arc<Self>, key: L::K) -> Result<L::V, String>;
-}
-
-#[async_trait]
 impl<L: LatGraph + 'static, M: LatMapping<L> + 'static> LatMappingExt<L> for M {
     async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String> {
         let value = self.clone().get_lattice_max(key.clone()).await?;

### 2022-10-04 src/lattice.rs
deps stuff

@@ -4,21 +4,25 @@ use std::sync::Arc;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
-pub enum LatResult<G: LatGraph, D, T> {
+pub enum LatLookup<G: LatGraph, T> {
     Done(Result<T, String>),
-    Lookup(D, G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatResult<G, D, T>>)
+    Lookup(G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatLookup<G, T>>)
 }
 
-impl<G: LatGraph + 'static, D: 'static, T: 'static> LatResult<G, D, T> {
-    pub fn and_then<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatResult<G, D, U>) -> LatResult<G, D, U> {
-        match self {
-            LatResult::Done(Ok(t)) => f(t),
-            LatResult::Done(Err(e)) => LatResult::Done(Err(e)),
-            LatResult::Lookup(key, dep, g) => LatResult::Lookup(key, dep, Box::new(move |lat, value| g(lat, value).and_then(f)))
+impl<G: LatGraph, T> LatLookup<G, T> {
+    fn to_set(self, lat: &G, mut lookup: impl FnMut(G::Key) -> Result<G::Value, String>) -> Result<(BTreeSet<G::Key>, T), String> {
+        let mut this = self;
+        let set = BTreeSet::new();
+        loop {
+            match this {
+                LatLookup::Done(Err(e)) => { return Err(e); },
+                LatLookup::Done(Ok(res)) => { return Ok((set, res)); },
+                LatLookup::Lookup(k, f) => {
+                    let v = lookup(k)?;
+                    this = f(lat, v);
+                }
+            }
         }
-    }
-    pub fn map<U>(self, f: impl 'static + Send + Sync + FnOnce(T) -> U) -> LatResult<G, D, U> {
-        self.and_then(move |t| LatResult::Done(Ok(f(t))))
     }
 }
 
@@ -28,18 +32,25 @@ pub trait LatGraph : Send + Sync + Sized {
 
     type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
 
+    type LatDeps : Send + Sync + Clone;
+
+    type ValueDeps : Send + Sync;
+
[…]

### 2022-10-04 src/mapping.rs

@@ -10,7 +10,17 @@ pub trait LatMapping<L: LatGraph> : Send + Sync {
 
     fn get_latgraph(&self) -> &L;
 
-    async fn get_lattice_max(self: Arc<Self>, key: L::K) -> Result<L::V, String>;
+    async fn get_lattice_max(self: Arc<Self>, key: L::K) -> Result<(L::V, HashCode), String>;
+}
+
+#[async_trait]
+pub trait LatMappingExt<L: LatGraph> : LatMapping<L> {
+
+    async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String>;
+
+    async fn join(self: Arc<Self>, key: L::K, v1: L::V, v2: L::V) -> Result<L::V, String>;
+
+    async fn autofill(self: Arc<Self>, key: L::K) -> Result<L::V, String>;
 }
 
 #[async_trait]

### 2022-10-04 src/main.rs
tagged mapping

@@ -1,4 +1,5 @@
 mod crypto;
+mod tagged_mapping;
 mod lattice;
 
 fn main() {

### 2022-10-04 src/tagged_mapping.rs

@@ -1,0 +1,130 @@
+use std::cmp::Ordering;
+use std::marker::PhantomData;
+use std::collections::BTreeMap;
+
+use serde::{Serialize, Deserialize, de::DeserializeOwned};
+
+pub trait TaggedMapping {
+    type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
+
+    type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
+}
+
+#[derive(Debug, Serialize, Deserialize)]
+pub struct Tag<M> {
+    index: usize,
+    phantom: PhantomData<M>
+}
+
+impl<M: TaggedMapping> Clone for Tag<M> {
+    fn clone(&self) -> Self {
+        Self {
+            index: self.index,
+            phantom: PhantomData,
+        }
+    }
+}
+
+impl<M: TaggedMapping> Copy for Tag<M> {}
+
+impl<M: TaggedMapping> PartialEq for Tag<M> {
+    fn eq(&self, other: &Self) -> bool {
+        self.index == other.index
+    }
+}
+
+impl<M: TaggedMapping> Eq for Tag<M> {}
+
+impl<M: TaggedMapping> PartialOrd for Tag<M> {
+    fn partial_cmp(&self, other: &Self) -> Option<Ordering> {
+        self.index.partial_cmp(&other.index)
+    }
+}
+
+impl<M: TaggedMapping> Ord for Tag<M> {
+    fn cmp(&self, other: &Self) -> Ordering {
+        self.index.cmp(&other.index)
+    }
+}
+
+pub struct TagRegistry {
+    n_registered: usize
+}
+
+impl TagRegistry {
+    pub fn register<M: TaggedMapping>(&mut self) -> Tag<M> {
+        let index = self.n_registered;
+        self.n_registered += 1;
+        Tag { index, phantom: PhantomData }
+    }
+}
+
+
+#[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Debug, Serialize, Deserialize)]
+pub struct TaggedKey {
+    index: usize,
+    key: Vec<u8>,
+}
+
+impl TaggedKey {
+    pub fn new<M: TaggedMapping>(tag: Tag<M>, key: &M::Key) -> Self {
+        Self { index: tag.index, key: rmp_serde::to_vec(key).unwrap() }
+    }
+    pub fn get_as<M: TaggedMapping>(&self, tag: Tag<M>) -> Result<M::Key, String> {
[…]

### 2022-10-04 src/lattice.rs
lookup more directly?

@@ -4,6 +4,8 @@ use std::sync::Arc;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
+use crate::tagged_mapping::TaggedMapping;
+
 pub enum LatLookup<G: LatGraph, T> {
     Done(Result<T, String>),
     Lookup(G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatLookup<G, T>>)
@@ -26,11 +28,7 @@ impl<G: LatGraph, T> LatLookup<G, T> {
     }
 }
 
-pub trait LatGraph : Send + Sync + Sized {
-
-    type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
-
-    type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
+pub trait LatGraph : Send + Sync + Sized + TaggedMapping {
 
     type LatDeps : Send + Sync + Clone;
 
@@ -40,17 +38,16 @@ pub trait LatGraph : Send + Sync + Sized {
         key1.partial_cmp(key2).ok_or("Keys are not comparable".to_string())
     }
 
-    fn lat_deps(&self, key: &Self::Key) -> LatLookup<Self, Self::LatDeps>;
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatLookup<Self, ()>;
 
-    fn value_deps(&self, key: &Self::Key, value: &Self::Value, lat_deps: Self::LatDeps) -> LatLookup<Self, Self::ValueDeps>;
+    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> LatLookup<Self, Self::Value>;
 
-    fn check_elem(&self, key: &Self::Key, value: &Self::Value, deps: Self::ValueDeps) -> Result<(), String>;
+    fn bottom(&self, key: &Self::Key, deps: Self::LatDeps) -> LatLookup<Self, ()>;
 
-    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value, deps1: Self::ValueDeps, deps2: Self::ValueDeps) -> Result<Self::Value, String>;
+    fn transport(&self, key: &Self::Key, value: &Self::Value) -> LatLookup<Self, LatLookup<Self, Self::Value>>;
+}
 
-    fn bottom(&self, key: &Self::Key, deps: Self::LatDeps) -> Result<Self::Value, String>;
-
[…]

### 2022-10-04 src/lattice.rs
tag stuff for latgraph

@@ -4,14 +4,14 @@ use std::sync::Arc;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
-use crate::tagged_mapping::TaggedMapping;
+use crate::tagged_mapping::{TaggedMapping, Tag, TaggedKey, TaggedValue};
 
 pub enum LatLookup<G: LatGraph, T> {
     Done(Result<T, String>),
     Lookup(G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatLookup<G, T>>)
 }
 
-impl<G: LatGraph, T> LatLookup<G, T> {
+impl<G: LatGraph + 'static, T : 'static> LatLookup<G, T> {
     fn to_set(self, lat: &G, mut lookup: impl FnMut(G::Key) -> Result<G::Value, String>) -> Result<(BTreeSet<G::Key>, T), String> {
         let mut this = self;
         let set = BTreeSet::new();
@@ -26,13 +26,32 @@ impl<G: LatGraph, T> LatLookup<G, T> {
             }
         }
     }
+    fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookup<G, T2>) -> LatLookup<G, T2> {
+        match self {
+            LatLookup::Done(Err(e)) => { LatLookup::Done(Err(e)) },
+            LatLookup::Done(Ok(res)) => { f(res) },
+            LatLookup::Lookup(k, g) => {
+                LatLookup::Lookup(k, Box::new(move |lat, v| g(lat, v).and_then(f)))
+            }
+        }
+    }
+    fn map<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> T2) -> LatLookup<G, T2> {
+        self.and_then(move |res| LatLookup::Done(Ok(f(res))))
+    }
+    fn map_latgraph<G2: LatGraph>(self, conv_g: impl 'static + Send + Sync + Fn(&G2) -> &G, conv_k: impl 'static + Send + Sync + Fn(G::Key) -> G2::Key, conv_v: impl 'static + Send + Sync + Fn(G2::Value) -> Result<G::Value, String>) -> LatLookup<G2, T> {
+        match self {
+            LatLookup::Done(res) => LatLookup::Done(res),
+            LatLookup::Lookup(k, f) => LatLookup::Lookup(conv_k(k), Box::new(move |g, v| {
+                match conv_v(v) {
+                    Err(e) => LatLookup::Done(Err(e)),
[…]

### 2022-10-04 src/tagged_mapping.rs

@@ -70,6 +70,9 @@ impl TaggedKey {
     pub fn new<M: TaggedMapping>(tag: Tag<M>, key: &M::Key) -> Self {
         Self { index: tag.index, key: rmp_serde::to_vec(key).unwrap() }
     }
+    pub fn get_index(&self) -> usize {
+        self.index
+    }
     pub fn get_as<M: TaggedMapping>(&self, tag: Tag<M>) -> Result<M::Key, String> {
         if self.index == tag.index {
             rmp_serde::from_slice(&self.key).map_err(|e| format!("{:?}", e))
@@ -88,6 +91,9 @@ pub struct TaggedValue {
 impl TaggedValue {
     pub fn new<M: TaggedMapping>(tag: Tag<M>, value: &M::Value) -> Self {
         Self { index: tag.index, value: rmp_serde::to_vec(value).unwrap() }
+    }
+    pub fn get_index(&self) -> usize {
+        self.index
     }
     pub fn get_as<M: TaggedMapping>(&self, tag: Tag<M>) -> Result<M::Value, String> {
         if self.index == tag.index {

### 2022-10-04 src/lattice.rs
multimap

@@ -6,13 +6,13 @@ use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
 use crate::tagged_mapping::{TaggedMapping, Tag, TaggedKey, TaggedValue};
 
-pub enum LatLookup<G: LatGraph, T> {
+pub enum LatLookup<K, V, T> {
     Done(Result<T, String>),
-    Lookup(G::Key, Box<dyn Send + Sync + FnOnce(&G, G::Value) -> LatLookup<G, T>>)
+    Lookup(K, Box<dyn Send + Sync + FnOnce(V) -> LatLookup<K, V, T>>)
 }
 
-impl<G: LatGraph + 'static, T : 'static> LatLookup<G, T> {
-    fn to_set(self, lat: &G, mut lookup: impl FnMut(G::Key) -> Result<G::Value, String>) -> Result<(BTreeSet<G::Key>, T), String> {
+impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
+    fn to_set(self, mut lookup: impl FnMut(K) -> Result<V, String>) -> Result<(BTreeSet<K>, T), String> {
         let mut this = self;
         let set = BTreeSet::new();
         loop {
@@ -21,49 +21,53 @@ impl<G: LatGraph + 'static, T : 'static> LatLookup<G, T> {
                 LatLookup::Done(Ok(res)) => { return Ok((set, res)); },
                 LatLookup::Lookup(k, f) => {
                     let v = lookup(k)?;
-                    this = f(lat, v);
+                    this = f(v);
                 }
             }
         }
     }
-    fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookup<G, T2>) -> LatLookup<G, T2> {
+    fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookup<K, V, T2>) -> LatLookup<K, V, T2> {
         match self {
             LatLookup::Done(Err(e)) => { LatLookup::Done(Err(e)) },
             LatLookup::Done(Ok(res)) => { f(res) },
             LatLookup::Lookup(k, g) => {
-                LatLookup::Lookup(k, Box::new(move |lat, v| g(lat, v).and_then(f)))
+                LatLookup::Lookup(k, Box::new(move |v| g(v).and_then(f)))
             }
         }
     }
[…]

### 2022-10-04 src/tagged_mapping.rs

@@ -16,6 +16,15 @@ pub struct Tag<M> {
     phantom: PhantomData<M>
 }
 
+impl<M: TaggedMapping> Tag<M> {
+    pub fn new(index: usize) -> Self {
+        Self {
+            index,
+            phantom: PhantomData
+        }
+    }
+}
+
 impl<M: TaggedMapping> Clone for Tag<M> {
     fn clone(&self) -> Self {
         Self {
@@ -55,7 +64,7 @@ impl TagRegistry {
     pub fn register<M: TaggedMapping>(&mut self) -> Tag<M> {
         let index = self.n_registered;
         self.n_registered += 1;
-        Tag { index, phantom: PhantomData }
+        Tag::new(index)
     }
 }
 

### 2022-10-04 src/lattice.rs
change results of lookups

@@ -7,45 +7,48 @@ use serde::{Serialize, Deserialize, de::DeserializeOwned};
 use crate::tagged_mapping::{TaggedMapping, Tag, TaggedKey, TaggedValue};
 
 pub enum LatLookup<K, V, T> {
-    Done(Result<T, String>),
-    Lookup(K, Box<dyn Send + Sync + FnOnce(V) -> LatLookup<K, V, T>>)
+    Done(T),
+    Lookup(K, Box<dyn Send + Sync + FnOnce(V) -> LatLookupResult<K, V, T>>)
 }
 
+type LatLookupResult<K, V, T> = Result<LatLookup<K, V, T>, String>;
+
 impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
-    fn to_set(self, mut lookup: impl FnMut(K) -> Result<V, String>) -> Result<(BTreeSet<K>, T), String> {
-        let mut this = self;
-        let set = BTreeSet::new();
-        loop {
-            match this {
-                LatLookup::Done(Err(e)) => { return Err(e); },
-                LatLookup::Done(Ok(res)) => { return Ok((set, res)); },
-                LatLookup::Lookup(k, f) => {
-                    let v = lookup(k)?;
-                    this = f(v);
-                }
-            }
-        }
-    }
-    fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookup<K, V, T2>) -> LatLookup<K, V, T2> {
+    // fn to_set(self, mut lookup: impl FnMut(K) -> Result<V, String>) -> Result<(BTreeSet<K>, T), String> {
+    //     let mut this = self;
+    //     let set = BTreeSet::new();
+    //     loop {
+    //         match this {
+    //             LatLookup::Done(Err(e)) => { return Err(e); },
+    //             LatLookup::Done(Ok(res)) => { return Ok((set, res)); },
+    //             LatLookup::Lookup(k, f) => {
+    //                 let v = lookup(k)?;
+    //                 this = f(v);
+    //             }
+    //         }
+    //     }
+    // }
+    fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookupResult<K, V, T2>) -> LatLookupResult<K, V, T2> {
         match self {
-            LatLookup::Done(Err(e)) => { LatLookup::Done(Err(e)) },
[…]

### 2022-10-04 paper/paper.tex
latresult
semiL

@@
 
-      The total space $\mathrm{SemiTot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
+      The total space $\mathrm{SemiLTot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
 We consider $(x, x') \leq (y, y')$ when $x \leq y$ and $g(x \leq y)(x') \leq y'$.
@@
       \begin{theorem}
-        The total space $\mathrm{SemiTot}(L, g)$ of a semilattice $L$ and a semilattice fibration $g$ is a semilattice.
+        The total space $\mathrm{SemiLTot}(L, g)$ of a semilattice $L$ and a semilattice fibration $g$ is a semilattice.
       \end{theorem}
@@
 
-      We write $\mathrm{SemiFibration}(L)$ for the set of semilattice fibrations on $L$.
+      We write $\mathrm{SemiLFibration}(L)$ for the set of semilattice fibrations on $L$.
 
@@
 
-      The total space $\mathrm{SemiTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$.
+      The total space $\mathrm{SemiLTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$.
 The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y)(x') \vee t(y, x \vee y)(y'))$.
@@
         L_0 &: \mathrm{Semilattice} & \\
-        (l_n, t_n) &: \mathrm{SemiFibration}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
-        L_n &:= \mathrm{SemiTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
+        (l_n, t_n) &: \mathrm{SemiLFibration}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
+        L_n &:= \mathrm{SemiLTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
       \end{align*}
@@
         L_{0, 0} &: \mathrm{Semilattice} & \\
-        (l_{u, n}, t_{u, n}) &: \mathrm{SemiFibration}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
-        L_{u, n} &:= \mathrm{SemiTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        (l_{u, n}, t_{u, n}) &: \mathrm{SemiLFibration}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        L_{u, n} &:= \mathrm{SemiLTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
         L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
@@
         \item $d_k(x \vee_{l_k(m)} y) \subseteq d_k(x) \cup d_k(y)$.
+        \item $d_k(\bot_{l_k(m)}) \subseteq d_k^\star$.
       \end{itemize}

### 2022-10-05 src/lattice.rs
to set async

@@ -1,6 +1,7 @@
 use std::collections::{BTreeSet, BTreeMap};
 use std::cmp::Ordering;
 use std::sync::Arc;
+use std::future::Future;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
@@ -14,20 +15,32 @@ pub enum LatLookup<K, V, T> {
 type LatLookupResult<K, V, T> = Result<LatLookup<K, V, T>, String>;
 
 impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
-    // fn to_set(self, mut lookup: impl FnMut(K) -> Result<V, String>) -> Result<(BTreeSet<K>, T), String> {
-    //     let mut this = self;
-    //     let set = BTreeSet::new();
-    //     loop {
-    //         match this {
-    //             LatLookup::Done(Err(e)) => { return Err(e); },
-    //             LatLookup::Done(Ok(res)) => { return Ok((set, res)); },
-    //             LatLookup::Lookup(k, f) => {
-    //                 let v = lookup(k)?;
-    //                 this = f(v);
-    //             }
-    //         }
-    //     }
-    // }
+    fn to_set(self, mut lookup: impl FnMut(K) -> Result<V, String>) -> Result<(BTreeSet<K>, T), String> {
+        let mut this = self;
+        let set = BTreeSet::new();
+        loop {
+            match this {
+                LatLookup::Done(res) => { return Ok((set, res)); },
+                LatLookup::Lookup(k, f) => {
+                    let v = lookup(k)?;
+                    this = f(v)?;
+                }
+            }
+        }
+    }
+    async fn to_set_async<F: Future<Output = Result<V, String>>>(self, mut lookup: impl Send + Sync + FnMut(K) -> F) -> Result<(BTreeSet<K>, T), String> {
+        let mut this = self;
+        let set = BTreeSet::new();
+        loop {
+            match this {
+                LatLookup::Done(res) => { return Ok((set, res)); },
+                LatLookup::Lookup(k, f) => {
+                    let v = lookup(k).await?;
+                    this = f(v)?;
+                }
+            }
+        }
+    }
     fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookupResult<K, V, T2>) -> LatLookupResult<K, V, T2> {
         match self {
             LatLookup::Done(res) => { f(res) },

### 2022-10-05 src/mapping.rs

@@ -16,11 +16,22 @@ pub trait LatMapping<L: LatGraph> : Send + Sync {
 #[async_trait]
 pub trait LatMappingExt<L: LatGraph> : LatMapping<L> {
 
-    async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String>;
+    fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {
+        key1.partial_cmp(key2).ok_or("Keys are not comparable".to_string())
+    }
 
-    async fn join(self: Arc<Self>, key: L::K, v1: L::V, v2: L::V) -> Result<L::V, String>;
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, ()> {
+        Ok(LatLookup::Done(()))
+    }
 
-    async fn autofill(self: Arc<Self>, key: L::K) -> Result<L::V, String>;
+    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
+
+    fn bottom(&self, key: &Self::Key) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
+
+    fn transport(&self, key: &Self::Key, value: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, LatLookup<Self::Key, Self::Value, Self::Value>> {
+        Ok(LatLookup::Done(LatLookup::Done(value.clone())))
+    }
+
 }
 
 #[async_trait]

### 2022-10-05 src/lattice.rs
cleanup

@@ -9,34 +9,36 @@ use crate::tagged_mapping::{TaggedMapping, Tag, TaggedKey, TaggedValue};
 
 pub enum LatLookup<K, V, T> {
     Done(T),
-    Lookup(K, Box<dyn Send + Sync + FnOnce(V) -> LatLookupResult<K, V, T>>)
+    Lookup(K, Box<dyn Send + Sync + FnOnce(&V) -> LatLookupResult<K, V, T>>)
 }
 
 type LatLookupResult<K, V, T> = Result<LatLookup<K, V, T>, String>;
 
 impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
-    fn to_set(self, mut lookup: impl FnMut(K) -> Result<V, String>) -> Result<(BTreeSet<K>, T), String> {
+    fn evaluate(self, mut lookup: impl FnMut(&K) -> Result<V, String>) -> Result<(BTreeMap<K, V>, T), String> {
         let mut this = self;
-        let set = BTreeSet::new();
+        let mut map = BTreeMap::new();
         loop {
             match this {
-                LatLookup::Done(res) => { return Ok((set, res)); },
+                LatLookup::Done(res) => { return Ok((map, res)); },
                 LatLookup::Lookup(k, f) => {
-                    let v = lookup(k)?;
-                    this = f(v)?;
+                    let v = lookup(&k)?;
+                    this = f(&v)?;
+                    map.insert(k, v);
                 }
             }
         }
     }
-    async fn to_set_async<F: Future<Output = Result<V, String>>>(self, mut lookup: impl Send + Sync + FnMut(K) -> F) -> Result<(BTreeSet<K>, T), String> {
+    async fn evaluate_async<F: Future<Output = Result<V, String>>>(self, mut lookup: impl Send + Sync + FnMut(&K) -> F) -> Result<(BTreeMap<K, V>, T), String> {
         let mut this = self;
-        let set = BTreeSet::new();
+        let mut map = BTreeMap::new();
         loop {
             match this {
-                LatLookup::Done(res) => { return Ok((set, res)); },
+                LatLookup::Done(res) => { return Ok((map, res)); },
                 LatLookup::Lookup(k, f) => {
-                    let v = lookup(k).await?;
-                    this = f(v)?;
[…]

### 2022-10-05 src/lattice.rs
eval_lookup

@@ -15,7 +15,7 @@ pub enum LatLookup<K, V, T> {
 type LatLookupResult<K, V, T> = Result<LatLookup<K, V, T>, String>;
 
 impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
-    fn evaluate(self, mut lookup: impl FnMut(&K) -> Result<V, String>) -> Result<(BTreeMap<K, V>, T), String> {
+    pub fn evaluate(self, mut lookup: impl FnMut(&K) -> Result<V, String>) -> Result<(BTreeMap<K, V>, T), String> {
         let mut this = self;
         let mut map = BTreeMap::new();
         loop {
@@ -29,7 +29,7 @@ impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
             }
         }
     }
-    async fn evaluate_async<F: Future<Output = Result<V, String>>>(self, mut lookup: impl Send + Sync + FnMut(&K) -> F) -> Result<(BTreeMap<K, V>, T), String> {
+    pub async fn evaluate_async<F: Future<Output = Result<V, String>>>(self, mut lookup: impl Send + Sync + FnMut(&K) -> F) -> Result<(BTreeMap<K, V>, T), String> {
         let mut this = self;
         let mut map = BTreeMap::new();
         loop {
@@ -43,7 +43,7 @@ impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
             }
         }
     }
-    fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookupResult<K, V, T2>) -> LatLookupResult<K, V, T2> {
+    pub fn and_then<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> LatLookupResult<K, V, T2>) -> LatLookupResult<K, V, T2> {
         match self {
             LatLookup::Done(res) => { f(res) },
             LatLookup::Lookup(k, g) => {
@@ -51,7 +51,7 @@ impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
             }
         }
     }
-    fn map<T2>(self, f: impl 'static + Send + Sync + FnOnce(T) -> T2) -> LatLookup<K, V, T2> {
[…]

### 2022-10-05 src/main.rs

@@ -1,6 +1,7 @@
 mod crypto;
 mod tagged_mapping;
 mod lattice;
+mod mapping;
 
 fn main() {
     println!("Hello, world!");

### 2022-10-05 src/mapping.rs

@@ -3,102 +3,108 @@ use std::sync::{Arc, Mutex};
 
 use async_trait::async_trait;
 
-use super::lattice::LatGraph;
+use crate::lattice::{LatGraph, LatLookup};
+use crate::crypto::HashCode;
 
 #[async_trait]
 pub trait LatMapping<L: LatGraph> : Send + Sync {
 
     fn get_latgraph(&self) -> &L;
 
-    async fn get_lattice_max(self: Arc<Self>, key: L::K) -> Result<(L::V, HashCode), String>;
+    async fn get_lattice_max(self: Arc<Self>, key: L::Key) -> Result<(L::Value, HashCode), String>;
 }
 
 #[async_trait]
-pub trait LatMappingExt<L: LatGraph> : LatMapping<L> {
+pub trait LatMappingExt<L: LatGraph + 'static> : LatMapping<L> {
 
-    fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {
-        key1.partial_cmp(key2).ok_or("Keys are not comparable".to_string())
+    async fn eval_lookup<T: Send + 'static>(self: Arc<Self>, lookup: LatLookup<L::Key, L::Value, T>) -> Result<(BTreeMap<L::Key, L::Value>, T), String> {
+        let self2 = self.clone();
+        lookup.evaluate_async(move |k| {
+            let k = k.clone();
+            let self2 = self.clone();
+            async move { Ok(self2.clone().get_lattice_max(k.clone()).await?.0) }
+        }).await
     }
 
-    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, ()> {
-        Ok(LatLookup::Done(()))
-    }
+    // async fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> Result<Self::Key, Self::Value, ()> {
+    //     Ok(LatLookup::Done(()))
+    // }
 
-    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
+    // fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
 
-    fn bottom(&self, key: &Self::Key) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
[…]

### 2022-10-05 src/mapping.rs
lat ext

@@ -26,46 +26,24 @@ pub trait LatMappingExt<L: LatGraph + 'static> : LatMapping<L> {
         }).await
     }
 
-    // async fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> Result<Self::Key, Self::Value, ()> {
-    //     Ok(LatLookup::Done(()))
-    // }
+    async fn check_elem(self: Arc<Self>, key: &L::Key, value: &L::Value) -> Result<(), String> {
+        Ok(self.clone().eval_lookup(self.get_latgraph().check_elem(key, value)?).await?.1)
+    }
 
-    // fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
+    async fn join(self: Arc<Self>, key: &L::Key, value1: &L::Value, value2: &L::Value) -> Result<L::Value, String> {
+        Ok(self.clone().eval_lookup(self.get_latgraph().join(key, value1, value2)?).await?.1)
+    }
 
-    // fn bottom(&self, key: &Self::Key) -> LatLookupResult<Self::Key, Self::Value, Self::Value>;
+    async fn bottom(self: Arc<Self>, key: &L::Key) -> Result<L::Value, String> {
+        Ok(self.clone().eval_lookup(self.get_latgraph().bottom(key)?).await?.1)
+    }
 
-    // fn transport(&self, key: &Self::Key, value: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, LatLookup<Self::Key, Self::Value, Self::Value>> {
-    //     Ok(LatLookup::Done(LatLookup::Done(value.clone())))
-    // }
-
+    async fn transport(self: Arc<Self>, new: Arc<Self>, key: &L::Key, value: &L::Value) -> Result<L::Value, String> {
+        let lookup_new = self.clone().eval_lookup(self.get_latgraph().transport(key, value)?).await?.1;
+        Ok(new.eval_lookup(lookup_new).await?.1)
+    }
 }
 
-// #[async_trait]
-// impl<L: LatGraph + 'static, M: LatMapping<L> + 'static> LatMappingExt<L> for M {
-//     async fn dependencies(self: Arc<Self>, key: L::K) -> Result<BTreeSet<L::K>, String> {
-//         let value = self.clone().get_lattice_max(key.clone()).await?;
-//         self.get_latgraph().dependencies(&key, Some(&value))
-//     }
[…]

### 2022-10-05 src/mapping.rs
join mappings!

@@ -3,15 +3,18 @@ use std::sync::{Arc, Mutex};
 
 use async_trait::async_trait;
 
+use crate::tagged_mapping::TaggedMapping;
 use crate::lattice::{LatGraph, LatLookup};
-use crate::crypto::HashCode;
+use crate::crypto::{Hash, hash};
+
+type DepsHash<L> = Hash<BTreeMap<<L as TaggedMapping>::Key, <L as TaggedMapping>::Value>>;
 
 #[async_trait]
 pub trait LatMapping<L: LatGraph> : Send + Sync {
 
     fn get_latgraph(&self) -> &L;
 
-    async fn get_lattice_max(self: Arc<Self>, key: L::Key) -> Result<(L::Value, HashCode), String>;
+    async fn get_lattice_max(self: Arc<Self>, key: L::Key) -> Result<(L::Value, DepsHash<L>), String>;
 }
 
 #[async_trait]
@@ -38,51 +41,56 @@ pub trait LatMappingExt<L: LatGraph + 'static> : LatMapping<L> {
         Ok(self.clone().eval_lookup(self.get_latgraph().bottom(key)?).await?.1)
     }
 
-    async fn transport(self: Arc<Self>, new: Arc<Self>, key: &L::Key, value: &L::Value) -> Result<L::Value, String> {
-        let lookup_new = self.clone().eval_lookup(self.get_latgraph().transport(key, value)?).await?.1;
-        Ok(new.eval_lookup(lookup_new).await?.1)
+    async fn transport(self: Arc<Self>, key: &L::Key, value: &L::Value) -> Result<LatLookup<L::Key, L::Value, L::Value>, String> {
+        Ok(self.clone().eval_lookup(self.get_latgraph().transport(key, value)?).await?.1)
     }
 }
 
-// struct JoinMapping<L: LatGraph, M1: LatMapping<L>, M2: LatMapping<L>> {
-//     m1: Arc<M1>,
-//     m2: Arc<M2>,
-//     cache: Mutex<BTreeMap<L::K, L::V>>,
-// }
-// 
-// impl<L: LatGraph, M1: LatMapping<L>, M2: LatMapping<L>> JoinMapping<L, M1, M2> {
-//     pub fn new(m1: Arc<M1>, m2: Arc<M2>) -> Self {
-//         JoinMapping { m1, m2, cache: Mutex::new(BTreeMap::new()) }
-//     }
-// }
-// 
[…]

### 2022-10-05 src/lattice.rs
check key ordering

@@ -72,7 +72,7 @@ impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
 pub trait LatGraph : Send + Sync + TaggedMapping {
 
     fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {
-        key1.partial_cmp(key2).ok_or("Keys are not comparable".to_string())
+        key1.partial_cmp(key2).ok_or(format!("Cannot compare keys: {:?} and {:?}", key1, key2))
     }
 
     fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, ()> {

### 2022-10-05 src/mapping.rs

@@ -1,5 +1,6 @@
 use std::collections::{BTreeSet, BTreeMap};
 use std::sync::{Arc, Mutex};
+use std::cmp::Ordering;
 
 use async_trait::async_trait;
 
@@ -20,29 +21,36 @@ pub trait LatMapping<L: LatGraph> : Send + Sync {
 #[async_trait]
 pub trait LatMappingExt<L: LatGraph + 'static> : LatMapping<L> {
 
-    async fn eval_lookup<T: Send + 'static>(self: Arc<Self>, lookup: LatLookup<L::Key, L::Value, T>) -> Result<(BTreeMap<L::Key, L::Value>, T), String> {
+    async fn eval_lookup<T: Send + 'static>(self: Arc<Self>, upper_key: &L::Key, lookup: LatLookup<L::Key, L::Value, T>) -> Result<(BTreeMap<L::Key, L::Value>, T), String> {
         let self2 = self.clone();
+        let upper_key = upper_key.clone();
         lookup.evaluate_async(move |k| {
             let k = k.clone();
             let self2 = self.clone();
-            async move { Ok(self2.clone().get_lattice_max(k.clone()).await?.0) }
+            let upper_key = upper_key.clone();
+            async move { 
+                if self2.get_latgraph().cmp_keys(&k, &upper_key)? != Ordering::Less {
+                    return Err(format!("Key {:?} is not less than {:?}", k, upper_key));
+                }
+                Ok(self2.clone().get_lattice_max(k.clone()).await?.0)
+            }
         }).await
     }
 
     async fn check_elem(self: Arc<Self>, key: &L::Key, value: &L::Value) -> Result<(), String> {
-        Ok(self.clone().eval_lookup(self.get_latgraph().check_elem(key, value)?).await?.1)
+        Ok(self.clone().eval_lookup(key, self.get_latgraph().check_elem(key, value)?).await?.1)
     }
 
     async fn join(self: Arc<Self>, key: &L::Key, value1: &L::Value, value2: &L::Value) -> Result<L::Value, String> {
-        Ok(self.clone().eval_lookup(self.get_latgraph().join(key, value1, value2)?).await?.1)
+        Ok(self.clone().eval_lookup(key, self.get_latgraph().join(key, value1, value2)?).await?.1)
     }
 
     async fn bottom(self: Arc<Self>, key: &L::Key) -> Result<L::Value, String> {
[…]

### 2022-10-05 src/tagged_mapping.rs

@@ -1,13 +1,14 @@
 use std::cmp::Ordering;
 use std::marker::PhantomData;
 use std::collections::BTreeMap;
+use std::fmt::Debug;
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
 pub trait TaggedMapping {
-    type Key : Eq + Ord + Clone + Send + Sync + Serialize + DeserializeOwned;
+    type Key : Eq + Ord + Clone + Send + Sync + Debug + Serialize + DeserializeOwned;
 
-    type Value : Eq + Clone + Send + Sync + Serialize + DeserializeOwned;
+    type Value : Eq + Clone + Send + Sync + Debug + Serialize + DeserializeOwned;
 }
 
 #[derive(Debug, Serialize, Deserialize)]
@@ -139,7 +140,7 @@ impl TaggedMap {
         self.map.insert(TaggedKey::new(tag, key), TaggedValue::new(tag, value));
     }
     fn get<M: TaggedMapping>(&self, tag: Tag<M>, key: &M::Key) -> Result<M::Value, String> {
-        let tag_value = self.map.get(&TaggedKey::new(tag, key)).ok_or("Key not found".to_string())?;
+        let tag_value = self.map.get(&TaggedKey::new(tag, key)).ok_or(format!("Key not found: {:?}", key))?;
         tag_value.get_as(tag)
     }
 }

### 2022-10-05 src/mapping.rs
just insert one elem!!!

@@ -100,6 +100,50 @@ impl<L: LatGraph + 'static, M1: LatMapping<L>, M2: LatMapping<L>> LatMapping<L> 
     }
 }
 
+struct ReplaceMapping<L: LatGraph, M: LatMapping<L>> {
+    m: Arc<M>,
+    key: L::Key,
+    value: L::Value,
+    cache: Mutex<BTreeMap<L::Key, (L::Value, DepsHash<L>)>>,
+}
+
+impl<L: LatGraph, M: LatMapping<L>> ReplaceMapping<L, M> {
+    pub fn new(m: Arc<M>, key: L::Key, value: L::Value) -> Self {
+        ReplaceMapping { m, key, value, cache: Mutex::new(BTreeMap::new()) }
+    }
+}
+
+#[async_trait]
+impl<L: LatGraph + 'static, M: LatMapping<L>> LatMapping<L> for ReplaceMapping<L, M> {
+    fn get_latgraph(&self) -> &L {
+        self.m.get_latgraph()
+    }
+
+    async fn get_lattice_max(self: Arc<Self>, key: L::Key) -> Result<(L::Value, DepsHash<L>), String> {
+        let cmp = self.get_latgraph().cmp_keys(&key, &self.key)?;
+        if cmp == Ordering::Less {
+            return self.m.clone().get_lattice_max(key).await;
+        }
+        {
+            let cache = self.cache.lock().unwrap();
+            if let Some(res) = cache.get(&key) {
+                return Ok(res.clone());
+            }
+        }
+        let (old_value, old_dephash) = self.m.clone().get_lattice_max(key.clone()).await?;
+        let res = if cmp == Ordering::Equal {
+            self.clone().join(&key, &old_value, &self.value).await?
+        } else {
+            self.clone().eval_lookup(&key, self.m.clone().transport(&key, &old_value).await?).await?.1
+        };
+        let (res_deps, ()) = self.clone().eval_lookup(&key, self.get_latgraph().check_elem(&key, &res)?).await?;
+        let mut cache = self.cache.lock().unwrap();
+        let hash_deps = hash(&res_deps);
+        cache.insert(key, (res.clone(), hash_deps));
+        Ok((res, hash_deps))
+    }
+}
+
 // pub struct MapLatticeReadDB<L: LatGraph> {
 //     latgraph: Arc<L>,
 //     values: Arc<BTreeMap<L::LID, L::Value>>,

### 2022-10-05 src/lattice.rs
tag lookup

@@ -69,6 +69,16 @@ impl<K: Ord + 'static, V: 'static, T : 'static> LatLookup<K, V, T> {
     }
 }
 
+type TagLatLookup<T> = LatLookup<TaggedKey, TaggedValue, T>;
+type TagLatLookupResult<T> = LatLookupResult<TaggedKey, TaggedValue, T>;
+
+impl<T> TagLatLookup<T> {
+    pub fn lookup<M: TaggedMapping + Send + Sync + 'static>(tag: Tag<M>, key: &M::Key, cont: impl 'static + Send + Sync + FnOnce(&M::Value) -> TagLatLookupResult<T>) -> TagLatLookup<T> {
+        LatLookup::Lookup(TaggedKey::new(tag, key), Box::new(move |v| Ok(cont(&v.get_as(tag)?)?)))
+    }
+
+}
+
 pub trait LatGraph : Send + Sync + TaggedMapping {
 
     fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {

### 2022-10-05 src/mapping.rs

@@ -120,6 +120,8 @@ impl<L: LatGraph + 'static, M: LatMapping<L>> LatMapping<L> for ReplaceMapping<L
     }
 
     async fn get_lattice_max(self: Arc<Self>, key: L::Key) -> Result<(L::Value, DepsHash<L>), String> {
+        // TODO: this evaluates on too many keys! all greater keys, not just dependents. maybe
+        // topological sort?
         let cmp = self.get_latgraph().cmp_keys(&key, &self.key)?;
         if cmp == Ordering::Less {
             return self.m.clone().get_lattice_max(key).await;

### 2022-10-05 paper/paper.tex
fix tagged lat graphs

@@
       It is somewhat unwieldy to define $l_k$ and $t_k$ in implementation, since they depend on a possibly-infinite mapping.
-Accordingly, we assume that these operations only depend on a defined set of values.
+To make the representation tractable, we assume $l_k$ and $t_k$ are \emph{continuous} in the sense that each of their outputs depend on only the values of a finite number of keys.
+
+      A continuous function $g_k : (K_{<k} \rightarrow V) \rightarrow S$, with $K_{<k}, V, S$ treated as having a discrete topology and the $K_{<k} \rightarrow V$ space treated as having the product topology, is continuous if for each $s \in S$, the preimage $f_k^{-1}(s)$ is open.
+An subset $F \subseteq K_{<k} \rightarrow V$ is open if it is a union of basis sets of the form $\{ f \in K_{<k} \rightarrow V | f(k_1) = v_1 \wedge \ldots \wedge f(k_n) = v_n \}$.
+A given $f : K_{<k} \rightarrow V$ in this set is, accordingly, in at least one of these bases.
+Any $f'$ that agrees with $f$ on the keys of this basis therefore has the same value $g(f') = g(f)$.
+We say the elements of the smallest such basis set are the dependencies of $g$ in $f$.
+
       Let $d_k^\star : 2^{K_{<k}}$ be a set of lattice dependencies, and $d_{k}(v) : 2^{K_{<k}}$ be a set of value dependencies for $v \in V$.

### 2022-10-05 src/lattice.rs

@@ -81,10 +81,6 @@ impl<T> TagLatLookup<T> {
 
 pub trait LatGraph : Send + Sync + TaggedMapping {
 
-    fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {
-        key1.partial_cmp(key2).ok_or(format!("Cannot compare keys: {:?} and {:?}", key1, key2))
-    }
-
     fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> LatLookupResult<Self::Key, Self::Value, ()> {
         Ok(LatLookup::Done(()))
     }
@@ -98,52 +94,68 @@ pub trait LatGraph : Send + Sync + TaggedMapping {
     }
 }
 
-pub struct TaggedLatGraph<G: LatGraph> {
+pub trait TaggableLatGraph : Send + Sync + TaggedMapping {
+    fn check_elem(&self, key: &Self::Key, value: &Self::Value) -> TagLatLookupResult<()> {
+        Ok(LatLookup::Done(()))
+    }
+
+    fn join(&self, key: &Self::Key, value1: &Self::Value, value2: &Self::Value) -> TagLatLookupResult<Self::Value>;
+
+    fn bottom(&self, key: &Self::Key) -> TagLatLookupResult<Self::Value>;
+
+    fn transport(&self, key: &Self::Key, value: &Self::Value) -> TagLatLookupResult<TagLatLookup<Self::Value>> {
+        Ok(LatLookup::Done(LatLookup::Done(value.clone())))
+    }
+
+}
+
+pub struct TaggedLatGraph<G: TaggableLatGraph> {
     latgraph: G,
     tag: Tag<G>,
 }
 
-impl<G: LatGraph + 'static> TaggedLatGraph<G> {
+impl<G: TaggableLatGraph + 'static> TaggedLatGraph<G> {
     fn new(latgraph: G, tag: Tag<G>) -> Self {
         TaggedLatGraph { latgraph, tag }
     }
-    fn map_lookup<T: 'static>(tag: Tag<G>, lookup: LatLookup<G::Key, G::Value, T>) -> LatLookup<TaggedKey, TaggedValue, T> {
-        lookup.map_kv(move |k| TaggedKey::new(tag, &k), move |v: &TaggedValue| v.get_as(tag))
-    }
+    // fn map_lookup<T: 'static>(tag: Tag<G>, lookup: LatLookup<G::Key, G::Value, T>) -> LatLookup<TaggedKey, TaggedValue, T> {
[…]

### 2022-10-05 src/tagged_mapping.rs

@@ -9,6 +9,11 @@ pub trait TaggedMapping {
     type Key : Eq + Ord + Clone + Send + Sync + Debug + Serialize + DeserializeOwned;
 
     type Value : Eq + Clone + Send + Sync + Debug + Serialize + DeserializeOwned;
+
+    fn cmp_keys(&self, key1: &Self::Key, key2: &Self::Key) -> Result<Ordering, String> {
+        key1.partial_cmp(key2).ok_or(format!("Cannot compare keys: {:?} and {:?}", key1, key2))
+    }
+
 }
 
 #[derive(Debug, Serialize, Deserialize)]

### 2022-11-18 src/db.rs
cargo fix

@@ -1,7 +1,7 @@
 use std::collections::{BTreeMap, BTreeSet};
 use std::sync::{Arc, Mutex, MutexGuard};
 
-use core::pin::Pin;
+
 use core::fmt::{Debug};
 
 use futures::Future;
@@ -10,10 +10,10 @@ use async_trait::async_trait;
 use async_recursion::async_recursion;
 
 use crate::error::Res;
-use crate::crypto::{HashCode, hash_of_bytes, Hash, hash};
+use crate::crypto::{HashCode, hash_of_bytes, Hash};
 use crate::tagged_mapping::TaggedMapping;
-use crate::lattice::{HashLookup, hash_lookup_generic, hash_put_generic, LatticeLibrary, ComputationLibrary, ComputationImmutContext, HashPut, LatticeImmutContext, LatticeMutContext, LatMerkleDeps, LatMerkleNode, EmptyContext};
-use serde::{Serialize, Deserialize, de::DeserializeOwned};
+use crate::lattice::{HashLookup, hash_lookup_generic, hash_put_generic, LatticeLibrary, ComputationLibrary, ComputationImmutContext, HashPut, LatticeImmutContext, LatticeMutContext, LatMerkleDeps, LatMerkleNode};
+use serde::{Serialize, Deserialize};
 
 /// a key-value store with key-key dependencies; auto-removes unpinned, non-depended-on keys
 pub trait DepDB<M: TaggedMapping> : Send {
@@ -247,7 +247,6 @@ impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping +
                 let ((), tr_deps) = LatDepsTracker::with_get_deps(self, move |this| async move {
                     lat_lib.check_elem(&key, &tr_value2, this).await
                 }).await?;
-                let merkle_deps = tr_deps.to_merkle_deps();
                 let merkle_hash = hash_put_generic(self, &LatMerkleNode {
                     value: tr_value,
                     deps: tr_deps.to_merkle_deps(),
@@ -266,13 +265,13 @@ impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping +
             }
             for key in dirty {
                 match &key {
-                    LatDBKey::Computation(comp_key) => {
+                    LatDBKey::Computation(_comp_key) => {
                         self.get_db().clear_value_deps(&key)?;
                     }
                     LatDBKey::Lattice(lat_key) => {
                         self.transport_dirty_lat(lat_key).await?;
[…]

### 2022-11-18 src/js/js_library.rs

@@ -2,7 +2,7 @@ use core::cmp::Ordering;
 use core::pin::Pin;
 use core::task::{Poll, Context};
 use std::sync::{Arc, Mutex};
-use std::rc::Rc;
+
 use std::sync::mpsc::{Sender, Receiver, channel, TryRecvError};
 use std::collections::BTreeMap;
 use std::thread::{spawn, JoinHandle};
@@ -15,13 +15,13 @@ use futures::future::{FutureExt, poll_fn};
 use async_trait::async_trait;
 use anyhow::bail;
 use serde::{Serialize, Deserialize};
-use deno_core::error::AnyError;
-use deno_core::{JsRuntime, Extension, RuntimeOptions, op, OpState, Resource};
+
+
 use deno_core::serde_json::{Value as SerdeJsValue, to_string as json_to_string};
 
 use crate::error::Res;
 use crate::tagged_mapping::TaggedMapping;
-use crate::crypto::{Hash, HashCode, hash};
+use crate::crypto::{Hash, HashCode};
 use crate::lattice::{HashLookup, ComputationImmutContext, HashPut, ComputationLibrary, LatticeLibrary, LatticeImmutContext, LatticeMutContext, LatMerkleNode, hash_lookup_generic, hash_put_generic};
 use crate::js::runtime_channel::{RuntimeState, CtxId, QueryId, LibraryQuery, LibraryResult, CtxQuery, CtxResult, MessageToRuntime, MessageFromRuntime};
 
@@ -81,8 +81,8 @@ impl HashLookup for DynContext {
 impl HashPut for DynContext {
      async fn hash_put(self: Arc<Self>, value: Vec<u8>) -> Res<HashCode> {
          match &*self {
-             DynContext::ComputationImmut(ctx) => bail!("Cannot hash_put in ComputationImmutContext"),
-             DynContext::LatticeImmut(ctx) => bail!("Cannot hash_put in LatticeImmutContext"),
+             DynContext::ComputationImmut(_ctx) => bail!("Cannot hash_put in ComputationImmutContext"),
+             DynContext::LatticeImmut(_ctx) => bail!("Cannot hash_put in LatticeImmutContext"),
              DynContext::LatticeMut(ctx) => ctx.clone().hash_put(value).await,
          }
      }
@@ -105,7 +105,7 @@ impl ComputationImmutContext<JsMapping> for DynContext {
 impl LatticeImmutContext<JsMapping, JsMapping, JsMapping> for DynContext {
[…]

### 2022-11-18 src/js/runtime_channel.rs

@@ -3,8 +3,8 @@ use std::ops::DerefMut;
 use std::collections::BTreeMap;
 use std::sync::{Arc, Mutex};
 use std::sync::mpsc::{Sender, Receiver, TryRecvError};
-use std::rc::Rc;
-use core::cmp::Ordering;
+
+
 use core::pin::Pin;
 
 use futures::task::{Poll, Context};
@@ -14,15 +14,15 @@ use futures::channel::oneshot;
 use anyhow::{anyhow, bail};
 use serde::{Serialize, Deserialize};
 use deno_core::serde_v8::{to_v8, from_v8};
-use deno_core::v8::{Value as V8Value, Function as V8Function, Local, Global};
+use deno_core::v8::{Function as V8Function, Local, Global};
 use deno_core::error::AnyError;
 use deno_core::{JsRuntime, Extension, RuntimeOptions, op, OpState, Resource};
-use deno_core::serde_json::{Value as JsValue, to_string as json_to_string};
+use deno_core::serde_json::{Value as JsValue};
 
 use crate::error::Res;
-use crate::tagged_mapping::TaggedMapping;
-use crate::crypto::{Hash, HashCode, hash};
-use crate::lattice::{HashLookup, ComputationImmutContext, HashPut, ComputationLibrary, LatticeLibrary, LatticeImmutContext, LatticeMutContext};
+
+use crate::crypto::{Hash, HashCode};
+
 
 pub type QueryId = u64;
 pub type CtxId = u64;

### 2022-11-18 src/lattice.rs

@@ -3,7 +3,7 @@ use std::collections::BTreeMap;
 use std::sync::Arc;
 use anyhow::bail;
 use async_trait::async_trait;
-use core::pin::Pin;
+
 
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
@@ -15,7 +15,7 @@ use crate::tagged_mapping::{TaggedMapping};
 #[async_trait]
 pub trait HashLookup : Send + Sync {
 
-    async fn hash_lookup(self: Arc<Self>, hash: HashCode) -> Res<Vec<u8>> {
+    async fn hash_lookup(self: Arc<Self>, _hash: HashCode) -> Res<Vec<u8>> {
         bail!("hash_lookup not implemented")
     }
 }
@@ -28,7 +28,7 @@ pub async fn hash_lookup_generic<H: HashLookup + ?Sized, T: DeserializeOwned>(hl
 #[async_trait]
 pub trait HashPut : HashLookup + Send + Sync {
 
-    async fn hash_put(self: Arc<Self>, value: Vec<u8>) -> Res<HashCode> {
+    async fn hash_put(self: Arc<Self>, _value: Vec<u8>) -> Res<HashCode> {
         bail!("hash_put not implemented")
     }
 
@@ -43,7 +43,7 @@ pub async fn hash_put_generic<H: HashPut + ?Sized, T: Serialize>(hp: &Arc<H>, va
 #[async_trait]
 pub trait ComputationImmutContext<C: TaggedMapping> : HashLookup + Send + Sync {
 
-    async fn eval_computation(self: Arc<Self>, key: &C::Key) -> Res<C::Value> {
+    async fn eval_computation(self: Arc<Self>, _key: &C::Key) -> Res<C::Value> {
         bail!("eval_computation not implemented")
     }
 }
@@ -52,7 +52,7 @@ pub trait ComputationImmutContext<C: TaggedMapping> : HashLookup + Send + Sync {
 #[async_trait]
 pub trait ComputationLibrary<C: TaggedMapping + 'static> : Send + Sync {
 
-    async fn eval_computation(self: Arc<Self>, key: &C::Key, ctx: Arc<dyn ComputationImmutContext<C>>) -> Res<C::Value> {
+    async fn eval_computation(self: Arc<Self>, _key: &C::Key, _ctx: Arc<dyn ComputationImmutContext<C>>) -> Res<C::Value> {
[…]

### 2022-11-18 src/transport/signalserver.rs

@@ -1,6 +1,6 @@
 use std::net::SocketAddr;
 
-use futures::{StreamExt, TryStreamExt};
+use futures::{StreamExt};
 
 use tokio::net::{TcpListener, TcpStream};
 use tungstenite::protocol::Message;

### 2022-11-18 src/db.rs
phantom

@@ -1,5 +1,6 @@
 use std::collections::{BTreeMap, BTreeSet};
 use std::sync::{Arc, Mutex, MutexGuard};
+use std::marker::PhantomData;
 
 
 use core::fmt::{Debug};
@@ -98,9 +99,9 @@ impl<CK: Ord + Clone, LK: Ord + Clone, LV, LCK: Ord + Clone, LCV> LatDeps<CK, LK
 }
 
 pub struct LatDBMapping<C: TaggedMapping, L: TaggedMapping, LC: TaggedMapping> {
-    computation: C,
-    lat: L,
-    lat_comp: LC,
+    computation: PhantomData<C>,
+    lat: PhantomData<L>,
+    lat_comp: PhantomData<LC>,
 }
 
 impl<C: TaggedMapping, L: TaggedMapping, LC: TaggedMapping> TaggedMapping for LatDBMapping<C, L, LC> {

### 2022-11-18 src/lattice.rs

@@ -1,10 +1,9 @@
 
 use std::collections::BTreeMap;
 use std::sync::Arc;
+
 use anyhow::bail;
 use async_trait::async_trait;
-
-
 use serde::{Serialize, Deserialize, de::DeserializeOwned};
 
 use crate::error::Res;

### 2022-11-18 src/sql_db.rs
test passes!

@@ -59,7 +59,7 @@ impl<M: TaggedMapping> SqlDepDB<M> {
     fn set_dependents_dirty_raw(&mut self, key: &[u8]) -> Res<()> {
         let mut non_dirty_deps = BTreeSet::new();
         {
-            let mut stmt = self.conn.prepare("SELECT dep FROM key_dep WHERE key = :key AND dirty = true")?
+            let mut stmt = self.conn.prepare("SELECT dep FROM key_dep INNER JOIN key_value ON key_dep.key = :key AND key_value.key = key_dep.dep AND dirty = true")?
                 .bind_by_name(":key", &*key)?;
             while stmt.next()? != State::Done {
                 let dep = stmt.read::<Vec<u8>>(0)?;

### 2022-11-18 tests/test_db.rs
more test

@@ -48,9 +48,13 @@ async fn test_db() {
     let store = Arc::new(LatStore::new(db, EmptyComputationLibrary, MaxTupleLatLibrary(3), EmptyContext));
     let key = "first".to_string();
     assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());
+
     assert_eq!(vec![1, 2, 0], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![1, 2, 0]).await.unwrap().unwrap()).await.unwrap().value);
-    // assert_eq!(vec![1, 2, 0], store.lattice_lookup(&key).unwrap().unwrap().0);
-    // assert_eq!(vec![4, 2, 1], store.lattice_join(&key, &vec![4, 0, 1], &EmptyContext).unwrap());
+    assert_eq!(vec![1, 2, 0], hash_lookup_generic(&store, store.clone().lattice_lookup(&key).await.unwrap().unwrap()).await.unwrap().value);
+
+    assert_eq!(vec![4, 2, 1], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![4, 0, 1]).await.unwrap().unwrap()).await.unwrap().value);
+    assert_eq!(vec![4, 2, 1], hash_lookup_generic(&store, store.clone().lattice_lookup(&key).await.unwrap().unwrap()).await.unwrap().value);
+
     // assert_eq!(vec![4, 2, 1], store.lattice_lookup(&key).unwrap().unwrap().0);
     // 
     // key = "second".to_string();

### 2022-11-18 tests/test_db.rs
more tests

@@ -46,7 +46,8 @@ async fn test_db() {
     let mut db = SqlDepDB::<LatDBMapping<EmptyMapping, MaxTupleMapping, EmptyMapping>>::new(":memory:").unwrap();
     db.initialize().unwrap();
     let store = Arc::new(LatStore::new(db, EmptyComputationLibrary, MaxTupleLatLibrary(3), EmptyContext));
-    let key = "first".to_string();
+
+    let mut key = "first".to_string();
     assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());
 
     assert_eq!(vec![1, 2, 0], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![1, 2, 0]).await.unwrap().unwrap()).await.unwrap().value);
@@ -55,14 +56,15 @@ async fn test_db() {
     assert_eq!(vec![4, 2, 1], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![4, 0, 1]).await.unwrap().unwrap()).await.unwrap().value);
     assert_eq!(vec![4, 2, 1], hash_lookup_generic(&store, store.clone().lattice_lookup(&key).await.unwrap().unwrap()).await.unwrap().value);
 
-    // assert_eq!(vec![4, 2, 1], store.lattice_lookup(&key).unwrap().unwrap().0);
-    // 
-    // key = "second".to_string();
-    // assert!(store.lattice_lookup(&key).await.unwrap().is_none());
-    // assert_eq!(vec![1, 0, 1], store.lattice_join(&key, &vec![1, 0, 1], &EmptyContext).unwrap());
-    // assert_eq!(vec![1, 0, 1], store.lattice_lookup(&key).unwrap().unwrap().0);
-    // assert_eq!(vec![1, 2, 3], store.lattice_join(&key, &vec![0, 2, 3], &EmptyContext).unwrap());
-    // assert_eq!(vec![1, 2, 3], store.lattice_lookup(&key).unwrap().unwrap().0);
+    key = "second".to_string();
+    assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());
 
-    // assert_eq!(vec![4, 2, 1], store.lattice_lookup(&"first".to_string()).unwrap().unwrap().0);
+    assert_eq!(vec![1, 0, 1], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![1, 0, 1]).await.unwrap().unwrap()).await.unwrap().value);
[…]

### 2022-11-18 tests/test_db.rs
tests shorter

@@ -40,6 +40,15 @@ impl LatticeLibrary<EmptyMapping, MaxTupleMapping, EmptyMapping> for MaxTupleLat
     }
 }
 
+async fn test_lookup<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping + 'static>(store: &Arc<LatStore<C, L, LC>>, key: &L::Key, expected: L::Value) {
+    assert_eq!(expected, hash_lookup_generic(&store, store.clone().lattice_lookup(key).await.unwrap().unwrap()).await.unwrap().value);
+}
+
+async fn test_join<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping + 'static>(store: &Arc<LatStore<C, L, LC>>, key: &L::Key, value: L::Value, expected: L::Value) {
+    assert_eq!(expected.clone(), hash_lookup_generic(&store, store.lattice_join_elem(key, value).await.unwrap().unwrap()).await.unwrap().value);
+    test_lookup(store, key, expected).await;
+}
+
 
 #[tokio::test]
 async fn test_db() {
@@ -50,21 +59,14 @@ async fn test_db() {
     let mut key = "first".to_string();
     assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());
 
-    assert_eq!(vec![1, 2, 0], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![1, 2, 0]).await.unwrap().unwrap()).await.unwrap().value);
-    assert_eq!(vec![1, 2, 0], hash_lookup_generic(&store, store.clone().lattice_lookup(&key).await.unwrap().unwrap()).await.unwrap().value);
-
-    assert_eq!(vec![4, 2, 1], hash_lookup_generic(&store, store.lattice_join_elem(&key, vec![4, 0, 1]).await.unwrap().unwrap()).await.unwrap().value);
-    assert_eq!(vec![4, 2, 1], hash_lookup_generic(&store, store.clone().lattice_lookup(&key).await.unwrap().unwrap()).await.unwrap().value);
+    test_join(&store, &key, vec![1, 2, 3], vec![1, 2, 3]).await;
+    test_join(&store, &key, vec![2, 1, 3], vec![2, 2, 3]).await;
 
     key = "second".to_string();
     assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());
 
[…]

### 2022-11-18 tests/test_db.rs
test same

@@ -61,6 +61,8 @@ async fn test_db() {
 
     test_join(&store, &key, vec![1, 2, 3], vec![1, 2, 3]).await;
     test_join(&store, &key, vec![2, 1, 3], vec![2, 2, 3]).await;
+    test_join(&store, &key, vec![1, 1, 1], vec![2, 2, 3]).await;
+    test_join(&store, &key, vec![2, 2, 3], vec![2, 2, 3]).await;
 
     key = "second".to_string();
     assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());

### 2022-11-18 src/db.rs
store constructor uses Arcs

@@ -199,14 +199,14 @@ impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping +
 
 impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping + 'static> LatStore<C, L, LC> {
     pub fn new(db: impl DepDB<LatDBMapping<C, L, LC>> + 'static,
-               comp_lib: impl ComputationLibrary<C> + 'static,
-               lat_lib: impl LatticeLibrary<C, L, LC> + 'static,
-               extra_hashlookup: impl HashLookup + 'static) -> Self {
+               comp_lib: Arc<impl ComputationLibrary<C> + 'static>,
+               lat_lib: Arc<impl LatticeLibrary<C, L, LC> + 'static>,
+               extra_hashlookup: Arc<impl HashLookup + 'static>) -> Self {
         Self {
             db: Arc::new(Mutex::new(db)),
-            comp_lib: Arc::new(comp_lib),
-            lat_lib: Arc::new(lat_lib),
-            extra_hashlookup: Arc::new(extra_hashlookup),
+            comp_lib: comp_lib,
+            lat_lib: lat_lib,
+            extra_hashlookup: extra_hashlookup,
             extra_hashlookup_cache: Arc::new(Mutex::new(BTreeMap::new())),
         }
     }

### 2022-11-18 tests/test_db.rs

@@ -54,7 +54,7 @@ async fn test_join<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: T
 async fn test_db() {
     let mut db = SqlDepDB::<LatDBMapping<EmptyMapping, MaxTupleMapping, EmptyMapping>>::new(":memory:").unwrap();
     db.initialize().unwrap();
-    let store = Arc::new(LatStore::new(db, EmptyComputationLibrary, MaxTupleLatLibrary(3), EmptyContext));
+    let store = Arc::new(LatStore::new(db, Arc::new(EmptyComputationLibrary), Arc::new(MaxTupleLatLibrary(3)), Arc::new(EmptyContext)));
 
     let mut key = "first".to_string();
     assert!(store.clone().lattice_lookup(&key).await.unwrap().is_none());

### 2022-11-18 src/db.rs
dyn instead of impl

@@ -199,9 +199,9 @@ impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping +
 
 impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping + 'static> LatStore<C, L, LC> {
     pub fn new(db: impl DepDB<LatDBMapping<C, L, LC>> + 'static,
-               comp_lib: Arc<impl ComputationLibrary<C> + 'static>,
-               lat_lib: Arc<impl LatticeLibrary<C, L, LC> + 'static>,
-               extra_hashlookup: Arc<impl HashLookup + 'static>) -> Self {
+               comp_lib: Arc<dyn ComputationLibrary<C> + 'static>,
+               lat_lib: Arc<dyn LatticeLibrary<C, L, LC> + 'static>,
+               extra_hashlookup: Arc<dyn HashLookup + 'static>) -> Self {
         Self {
             db: Arc::new(Mutex::new(db)),
             comp_lib: comp_lib,

### 2022-11-18 src/transport/signalclient.rs
rtcsignalclient

@@ -1,10 +1,53 @@
+use core::pin::Pin;
+use std::sync::Mutex;
 
-use futures::StreamExt;
+use futures::{Future, StreamExt};
+use tokio::net::TcpStream;
 use tungstenite::protocol::Message;
-use tokio_tungstenite::connect_async;
+use tokio_tungstenite::{WebSocketStream, MaybeTlsStream, connect_async};
+use webrtc::ice_transport::ice_candidate::RTCIceCandidateInit;
+use webrtc::peer_connection::sdp::session_description::RTCSessionDescription;
 
 use crate::error::Res;
 use crate::transport::signalmessage::SignalMessageToClient;
+use crate::transport::webrtc::RTCSignalClient;
+
+pub struct SignalClient {
+    ws_stream: WebSocketStream<MaybeTlsStream<TcpStream>>,
+    remote_session_description_handler: Mutex<Option<Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
+    remote_ice_candidate_handler: Mutex<Option<Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
+}
+
+impl SignalClient {
+    async fn new(addr: &str) -> Res<Self> {
+        let (ws_stream, _) = connect_async(addr).await?;
+        Ok(Self {
+            ws_stream,
+            remote_session_description_handler: Mutex::new(None),
+            remote_ice_candidate_handler: Mutex::new(None),
+        })
+    }
+}
+
+impl RTCSignalClient for SignalClient {
+    fn send_session_description(&self, sdp: RTCSessionDescription) {
+        unimplemented!()
+    }
+
+    fn on_remote_session_description(&self, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) {
+        let mut handler = self.remote_session_description_handler.lock().unwrap();
+        *handler = Some(fun);
+    }
+
+    fn send_ice_candidate(&self, candidate: RTCIceCandidateInit) {
+        unimplemented!()
+    }
+
+    fn on_remote_ice_candidate(&self, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) {
[…]

### 2022-11-18 src/transport/webrtc.rs

@@ -31,19 +31,19 @@ use crate::lattice::{HashLookup, LatMerkleNode};
 // https://github.com/webrtc-rs/webrtc/blob/master/examples/examples/data-channels-create/data-channels-create.rs
 
 #[async_trait]
-trait RemoteAccessibleContext : HashLookup {
+pub trait RemoteAccessibleContext : HashLookup {
     async fn lattice_lookup(self: Arc<Self>, key: Vec<u8>) -> Res<Option<Hash<LatMerkleNode<Vec<u8>, Vec<u8>, Vec<u8>, Vec<u8>, Vec<u8>>>>>;
 }
 
-trait RTCSignalClient : Send + Sync {
+pub trait RTCSignalClient : Send + Sync {
 
     fn send_session_description(&self, sdp: RTCSessionDescription);
 
-    fn on_remote_session_description(&self, fun: Box<dyn Fn(RTCSessionDescription) -> Pin<Box<dyn Future<Output = Res<()>>>>>);
+    fn on_remote_session_description(&self, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>);
 
     fn send_ice_candidate(&self, candidate: RTCIceCandidateInit);
 
-    fn on_remote_ice_candidate(&self, fun: Box<dyn Fn(RTCIceCandidateInit) -> Pin<Box<dyn Future<Output = Res<()>>>>>);
+    fn on_remote_ice_candidate(&self, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>);
 }
 
 struct DataChannel<T> {

### 2022-11-18 src/transport/signalclient.rs
signal client async

@@ -1,6 +1,7 @@
 use core::pin::Pin;
-use std::sync::Mutex;
+use std::sync::{Arc, Mutex};
 
+use async_trait::async_trait;
 use futures::{Future, StreamExt};
 use tokio::net::TcpStream;
 use tungstenite::protocol::Message;
@@ -19,7 +20,7 @@ pub struct SignalClient {
 }
 
 impl SignalClient {
-    async fn new(addr: &str) -> Res<Self> {
+    pub async fn new(addr: &str) -> Res<Self> {
         let (ws_stream, _) = connect_async(addr).await?;
         Ok(Self {
             ws_stream,
@@ -29,23 +30,26 @@ impl SignalClient {
     }
 }
 
+#[async_trait]
 impl RTCSignalClient for SignalClient {
-    fn send_session_description(&self, sdp: RTCSessionDescription) {
+    async fn send_session_description(self: Arc<Self>, sdp: RTCSessionDescription) -> Res<()> {
         unimplemented!()
     }
 
-    fn on_remote_session_description(&self, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) {
+    async fn on_remote_session_description(self: Arc<Self>, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()> {
         let mut handler = self.remote_session_description_handler.lock().unwrap();
         *handler = Some(fun);
+        Ok(())
     }
 
-    fn send_ice_candidate(&self, candidate: RTCIceCandidateInit) {
+    async fn send_ice_candidate(self: Arc<Self>, candidate: RTCIceCandidateInit) -> Res<()> {
         unimplemented!()
     }
 
-    fn on_remote_ice_candidate(&self, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) {
+    async fn on_remote_ice_candidate(self: Arc<Self>, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()> {
         let mut handler = self.remote_ice_candidate_handler.lock().unwrap();
         *handler = Some(fun);
+        Ok(())
     }
 }
 

### 2022-11-18 src/transport/webrtc.rs

@@ -35,15 +35,16 @@ pub trait RemoteAccessibleContext : HashLookup {
     async fn lattice_lookup(self: Arc<Self>, key: Vec<u8>) -> Res<Option<Hash<LatMerkleNode<Vec<u8>, Vec<u8>, Vec<u8>, Vec<u8>, Vec<u8>>>>>;
 }
 
+#[async_trait]
 pub trait RTCSignalClient : Send + Sync {
 
-    fn send_session_description(&self, sdp: RTCSessionDescription);
+    async fn send_session_description(self: Arc<Self>, sdp: RTCSessionDescription) -> Res<()>;
 
-    fn on_remote_session_description(&self, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>);
+    async fn on_remote_session_description(self: Arc<Self>, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
 
-    fn send_ice_candidate(&self, candidate: RTCIceCandidateInit);
+    async fn send_ice_candidate(self: Arc<Self>, candidate: RTCIceCandidateInit) -> Res<()>;
 
-    fn on_remote_ice_candidate(&self, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>);
+    async fn on_remote_ice_candidate(self: Arc<Self>, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
 }
 
 struct DataChannel<T> {
@@ -258,11 +259,11 @@ impl PeerConnection {
                 rtc_connection.set_remote_description(sdp).await?;
                 Ok(())
             })
-        }));
+        })).await?;
 
         let offer = self.rtc_connection.create_offer(None).await?;
         self.rtc_connection.set_local_description(offer.clone()).await?;
-        self.signal_client.send_session_description(offer);
+        self.signal_client.clone().send_session_description(offer).await?;
 
         let rtc_connection = self.rtc_connection.clone();
 
@@ -272,7 +273,7 @@ impl PeerConnection {
                 rtc_connection.add_ice_candidate(candidate).await?;
                 Ok(())
             })
-        }));
+        })).await?;
 
[…]

### 2022-11-18 src/transport/signalclient.rs
send signal messages

@@ -1,8 +1,10 @@
 use core::pin::Pin;
 use std::sync::{Arc, Mutex};
+use std::collections::BTreeMap;
 
+use async_mutex::Mutex as AsyncMutex;
 use async_trait::async_trait;
-use futures::{Future, StreamExt};
+use futures::{Future, StreamExt, SinkExt};
 use tokio::net::TcpStream;
 use tungstenite::protocol::Message;
 use tokio_tungstenite::{WebSocketStream, MaybeTlsStream, connect_async};
@@ -10,45 +12,51 @@ use webrtc::ice_transport::ice_candidate::RTCIceCandidateInit;
 use webrtc::peer_connection::sdp::session_description::RTCSessionDescription;
 
 use crate::error::Res;
-use crate::transport::signalmessage::SignalMessageToClient;
+use crate::transport::signalmessage::{SignalMessageToClient, SignalMessageToServer, Peer};
 use crate::transport::webrtc::RTCSignalClient;
 
 pub struct SignalClient {
-    ws_stream: WebSocketStream<MaybeTlsStream<TcpStream>>,
-    remote_session_description_handler: Mutex<Option<Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
-    remote_ice_candidate_handler: Mutex<Option<Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
+    ws_stream: AsyncMutex<WebSocketStream<MaybeTlsStream<TcpStream>>>,
+    remote_session_description_handler: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
+    remote_ice_candidate_handler: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
 }
 
 impl SignalClient {
     pub async fn new(addr: &str) -> Res<Self> {
         let (ws_stream, _) = connect_async(addr).await?;
         Ok(Self {
-            ws_stream,
-            remote_session_description_handler: Mutex::new(None),
-            remote_ice_candidate_handler: Mutex::new(None),
+            ws_stream: AsyncMutex::new(ws_stream),
[…]

### 2022-11-18 src/transport/signalmessage.rs

@@ -1,16 +1,22 @@
 
+use webrtc::ice_transport::ice_candidate::RTCIceCandidateInit;
+use webrtc::peer_connection::sdp::session_description::RTCSessionDescription;
 use serde::{Serialize, Deserialize};
 
 use crate::crypto::{Hash, PublicKey};
 
-#[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Debug, Serialize, Deserialize)]
+pub type Peer = Hash<PublicKey>;
+
+#[derive(Clone, Debug, Serialize, Deserialize)]
 pub enum SignalMessageToServer {
     GetPeers,
-    Send(Hash<PublicKey>, Vec<u8>),
+    SessionDescription(Peer, RTCSessionDescription),
+    IceCandidate(Peer, RTCIceCandidateInit),
 }
 
-#[derive(PartialEq, Eq, PartialOrd, Ord, Clone, Debug, Serialize, Deserialize)]
+#[derive(Clone, Debug, Serialize, Deserialize)]
 pub enum SignalMessageToClient {
-    Peers(Vec<Hash<PublicKey>>),
-    Received(Hash<PublicKey>, Vec<u8>),
+    Peers(Vec<Peer>),
+    SessionDescription(Peer, RTCSessionDescription),
+    IceCandidate(Peer, RTCIceCandidateInit),
 }

### 2022-11-18 src/transport/webrtc.rs

@@ -27,6 +27,7 @@ use crate::tagged_mapping::TaggedMapping;
 use crate::crypto::{Hash, HashCode};
 use crate::error::{Res, to_string_result};
 use crate::lattice::{HashLookup, LatMerkleNode};
+use crate::transport::signalmessage::Peer;
 
 // https://github.com/webrtc-rs/webrtc/blob/master/examples/examples/data-channels-create/data-channels-create.rs
 
@@ -38,13 +39,13 @@ pub trait RemoteAccessibleContext : HashLookup {
 #[async_trait]
 pub trait RTCSignalClient : Send + Sync {
 
-    async fn send_session_description(self: Arc<Self>, sdp: RTCSessionDescription) -> Res<()>;
+    async fn send_session_description(self: Arc<Self>, peer: Peer, sdp: RTCSessionDescription) -> Res<()>;
 
-    async fn on_remote_session_description(self: Arc<Self>, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
+    async fn on_remote_session_description(self: Arc<Self>, peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
 
-    async fn send_ice_candidate(self: Arc<Self>, candidate: RTCIceCandidateInit) -> Res<()>;
+    async fn send_ice_candidate(self: Arc<Self>, peer: Peer, candidate: RTCIceCandidateInit) -> Res<()>;
 
-    async fn on_remote_ice_candidate(self: Arc<Self>, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
+    async fn on_remote_ice_candidate(self: Arc<Self>, peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
 }
 
 struct DataChannel<T> {
@@ -219,6 +220,7 @@ async fn poll_mutex_option<T: Clone>(mutex: &Mutex<Option<T>>) -> T {
 // - lattice lookup
 // - computations? lattice computations? get cached values only
 struct PeerConnection {
[…]

### 2022-11-18 src/transport/signalclient.rs
handle messages

@@ -30,6 +30,26 @@ impl SignalClient {
             remote_ice_candidate_handler: Mutex::new(BTreeMap::new()),
         })
     }
+
+    pub async fn handle_messages(self: &Arc<Self>) -> Res<()> {
+        loop {
+            if let Some(msg) = {
+                let mut ws_stream = self.ws_stream.lock().await;
+                ws_stream.next().await
+            } {
+                match msg? {
+                    Message::Binary(bs) => {
+                        let msg: SignalMessageToClient = rmp_serde::from_slice(&bs)?;
+                        println!("Received a message: {:?}", msg);
+                        // ...
+                    }
+                    other => {
+                        println!("Received a message which is not binary: {:?}", other);
+                    }
+                }
+            }
+        }
+    }
 }
 
 #[async_trait]

### 2022-11-18 src/transport/signalclient.rs
actually handle messages

@@ -40,8 +40,21 @@ impl SignalClient {
                 match msg? {
                     Message::Binary(bs) => {
                         let msg: SignalMessageToClient = rmp_serde::from_slice(&bs)?;
-                        println!("Received a message: {:?}", msg);
-                        // ...
+                        match msg {
+                            SignalMessageToClient::SessionDescription(peer, desc) => {
+                                if let Some(handler) = self.remote_session_description_handler.lock().unwrap().get(&peer) {
+                                    handler(desc).await?;
+                                }
+                            },
+                            SignalMessageToClient::IceCandidate(peer, candidate) => {
+                                if let Some(handler) = self.remote_ice_candidate_handler.lock().unwrap().get(&peer) {
+                                    handler(candidate).await?;
+                                }
+                            },
+                            _ => {
+                                println!("Received unexpected message: {:?}", msg);
+                            }
+                        }
                     }
                     other => {
                         println!("Received a message which is not binary: {:?}", other);

### 2022-11-18 src/transport/signalclient.rs
remove dead code

@@ -15,6 +15,8 @@ use crate::error::Res;
 use crate::transport::signalmessage::{SignalMessageToClient, SignalMessageToServer, Peer};
 use crate::transport::webrtc::RTCSignalClient;
 
+// https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-client.rs
+
 pub struct SignalClient {
     ws_stream: AsyncMutex<WebSocketStream<MaybeTlsStream<TcpStream>>>,
     remote_session_description_handler: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
@@ -95,23 +97,3 @@ impl RTCSignalClient for SignalClient {
 }
 
 
-// https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-client.rs
-
-
-async fn runclient(addr: &str) -> Res<()> {
-    let (mut ws_stream, _) = connect_async(addr).await?;
-    while let Some(msg) = ws_stream.next().await {
-        match msg? {
-            Message::Binary(bs) => {
-                let msg: SignalMessageToClient = rmp_serde::from_slice(&bs)?;
-                println!("Received a message: {:?}", msg);
-                // ...
-            }
-            other => {
-                println!("Received a message which is not binary: {:?}", other);
-            }
-        }
-    }
-
-    Ok(())
-}

### 2022-11-18 src/transport/signalserver.rs
server stuff

@@ -1,15 +1,22 @@
 use std::net::SocketAddr;
+use std::collections::BTreeMap;
+use std::sync::Mutex;
 
 use futures::{StreamExt};
 
+use tokio_tungstenite::{WebSocketStream, MaybeTlsStream, connect_async};
 use tokio::net::{TcpListener, TcpStream};
 use tungstenite::protocol::Message;
 
 
 use crate::error::Res;
-use crate::transport::signalmessage::{SignalMessageToServer};
+use crate::transport::signalmessage::{Peer, SignalMessageToClient, SignalMessageToServer};
 
 // https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-server.rs
+
+struct SignalServer {
+    peer_streams: Mutex<BTreeMap<Peer, WebSocketStream<MaybeTlsStream<TcpStream>>>>,
+}
 
 async fn runserver() -> Res<()> {
     // let addr = env::args().nth(1).unwrap_or_else(|| "127.0.0.1:8080".to_string());

### 2022-11-18 src/transport/signalclient.rs
signal client uses spawn

@@ -1,3 +1,4 @@
+use core::mem::drop;
 use core::pin::Pin;
 use std::sync::{Arc, Mutex};
 use std::collections::BTreeMap;
@@ -45,12 +46,24 @@ impl SignalClient {
                         match msg {
                             SignalMessageToClient::SessionDescription(peer, desc) => {
                                 if let Some(handler) = self.remote_session_description_handler.lock().unwrap().get(&peer) {
-                                    handler(desc).await?;
+                                    let fut = handler(desc);
+                                    drop(handler);
+                                    tokio::spawn(async move {
+                                        if let Err(e) = fut.await {
+                                            println!("Error handling remote session description: {}", e);
+                                        }
+                                    });
                                 }
                             },
                             SignalMessageToClient::IceCandidate(peer, candidate) => {
                                 if let Some(handler) = self.remote_ice_candidate_handler.lock().unwrap().get(&peer) {
-                                    handler(candidate).await?;
+                                    let fut = handler(candidate);
+                                    drop(handler);
+                                    tokio::spawn(async move {
+                                        if let Err(e) = fut.await {
+                                            println!("Error handling remote ice candidate: {}", e);
+                                        }
+                                    });
                                 }
                             },
                             _ => {

### 2022-11-18 src/transport/webrtc.rs
on_receive spawns

@@ -95,10 +95,12 @@ impl<T: Serialize + DeserializeOwned + Send + Sync + 'static> DataChannel<T> {
                             return;
                         }
                     };
-                    match fun(value).await {
-                        Ok(()) => (),
-                        Err(e) => println!("Failed to handle message: {:?}", e),
-                    }
+                    tokio::spawn(async move {
+                        match fun(value).await {
+                            Ok(()) => (),
+                            Err(e) => println!("Failed to handle message: {:?}", e),
+                        }
+                    });
                 }
             })
         })).await;

### 2022-11-18 src/transport/signalclient.rs
rename handler to handlers

@@ -20,8 +20,8 @@ use crate::transport::webrtc::RTCSignalClient;
 
 pub struct SignalClient {
     ws_stream: AsyncMutex<WebSocketStream<MaybeTlsStream<TcpStream>>>,
-    remote_session_description_handler: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
-    remote_ice_candidate_handler: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
+    remote_session_description_handlers: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
+    remote_ice_candidate_handlers: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
 }
 
 impl SignalClient {
@@ -29,8 +29,8 @@ impl SignalClient {
         let (ws_stream, _) = connect_async(addr).await?;
         Ok(Self {
             ws_stream: AsyncMutex::new(ws_stream),
-            remote_session_description_handler: Mutex::new(BTreeMap::new()),
-            remote_ice_candidate_handler: Mutex::new(BTreeMap::new()),
+            remote_session_description_handlers: Mutex::new(BTreeMap::new()),
+            remote_ice_candidate_handlers: Mutex::new(BTreeMap::new()),
         })
     }
 
@@ -45,7 +45,7 @@ impl SignalClient {
                         let msg: SignalMessageToClient = rmp_serde::from_slice(&bs)?;
                         match msg {
                             SignalMessageToClient::SessionDescription(peer, desc) => {
-                                if let Some(handler) = self.remote_session_description_handler.lock().unwrap().get(&peer) {
+                                if let Some(handler) = self.remote_session_description_handlers.lock().unwrap().get(&peer) {
                                     let fut = handler(desc);
                                     drop(handler);
                                     tokio::spawn(async move {
@@ -56,7 +56,7 @@ impl SignalClient {
                                 }
                             },
                             SignalMessageToClient::IceCandidate(peer, candidate) => {
[…]

### 2022-11-18 src/transport/webrtc.rs

@@ -46,6 +46,8 @@ pub trait RTCSignalClient : Send + Sync {
     async fn send_ice_candidate(self: Arc<Self>, peer: Peer, candidate: RTCIceCandidateInit) -> Res<()>;
 
     async fn on_remote_ice_candidate(self: Arc<Self>, peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
+
+    async fn remove_peer(self: Arc<Self>, peer: Peer) -> Res<()>;
 }
 
 struct DataChannel<T> {
@@ -297,6 +299,7 @@ impl PeerConnection {
             Box::pin(async move {
                 println!("ICE Candidate: {:?}", candidate);
                 if let Some(candidate) = candidate {
+                    // TODO spawn?
                     signal_client.clone().send_ice_candidate(peer, candidate.to_json().await.unwrap()).await.unwrap();
                 }
             })

### 2022-11-18 src/transport/webrtc.rs
remove peer when disconnected

@@ -282,17 +282,26 @@ impl PeerConnection {
             })
         })).await?;
 
-        self.rtc_connection.on_peer_connection_state_change(Box::new(|s: RTCPeerConnectionState| Box::pin(async move {
-            println!("Peer Connection State has changed: {}", s);
-
-            if s == RTCPeerConnectionState::Failed {
-                panic!("Peer Connection has gone to failed exiting");
-            }
-
-        }))).await;
-
         let signal_client = self.signal_client.clone();
         let peer = self.peer;
+
+        self.rtc_connection.on_peer_connection_state_change(Box::new(move |s: RTCPeerConnectionState| {
+            let signal_client = signal_client.clone();
+
+            Box::pin(async move {
+                println!("Peer Connection State has changed: {}", s);
+
+                if s == RTCPeerConnectionState::Failed ||
+                   s == RTCPeerConnectionState::Disconnected ||
+                   s == RTCPeerConnectionState::Closed {
+                       signal_client.clone().remove_peer(peer).await.unwrap();
+                }
+            })
+
+        })).await;
+
+
+        let signal_client = self.signal_client.clone();
 
         self.rtc_connection.on_ice_candidate(Box::new(move |candidate| {
             let signal_client = signal_client.clone();

### 2022-11-19 src/transport/signalserver.rs
signal server as struct

@@ -1,6 +1,6 @@
 use std::net::SocketAddr;
 use std::collections::BTreeMap;
-use std::sync::Mutex;
+use std::sync::{Arc, Mutex};
 
 use futures::{StreamExt};
 
@@ -14,50 +14,56 @@ use crate::transport::signalmessage::{Peer, SignalMessageToClient, SignalMessage
 
 // https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-server.rs
 
-struct SignalServer {
+pub struct SignalServer {
     peer_streams: Mutex<BTreeMap<Peer, WebSocketStream<MaybeTlsStream<TcpStream>>>>,
 }
 
-async fn runserver() -> Res<()> {
-    // let addr = env::args().nth(1).unwrap_or_else(|| "127.0.0.1:8080".to_string());
-    let addr = "127.0.0.1:2020";
+impl SignalServer {
+    pub fn new() -> Self {
+        let peer_streams = Mutex::new(BTreeMap::new());
+        SignalServer { peer_streams }
+    }
+    pub async fn listen(self: Arc<Self>, addr: &str) -> Res<()> {
+        let addr = "127.0.0.1:2020";
 
-    // Create the event loop and TCP listener we'll accept connections on.
-    let listener = TcpListener::bind(&addr).await?;
-    println!("Listening on: {}", addr);
+        // Create the event loop and TCP listener we'll accept connections on.
+        let listener = TcpListener::bind(&addr).await?;
+        println!("Listening on: {}", addr);
 
-    while let Ok((stream, _)) = listener.accept().await {
-        tokio::spawn(accept_connection(stream));
+        while let Ok((stream, _)) = listener.accept().await {
+            tokio::spawn(self.clone().accept_connection(stream));
+        }
+
+        Ok(())
     }
 
-    Ok(())
-}
-
-async fn accept_connection(stream: TcpStream) {
-    let addr = stream.peer_addr().expect("connected streams should have a peer address");
-    println!("Peer address: {}", addr);
+    async fn accept_connection(self: Arc<Self>, stream: TcpStream) {
+        let addr = stream.peer_addr().expect("connected streams should have a peer address");
+        println!("Peer address: {}", addr);
 
 
-    println!("New WebSocket connection: {}", addr);
+        println!("New WebSocket connection: {}", addr);
 
-    if let Err(e) = handle_connection(addr, stream).await {
[…]

### 2022-11-19 src/transport/signalclient.rs
this peer

@@ -19,15 +19,17 @@ use crate::transport::webrtc::RTCSignalClient;
 // https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-client.rs
 
 pub struct SignalClient {
+    this_peer: Peer,
     ws_stream: AsyncMutex<WebSocketStream<MaybeTlsStream<TcpStream>>>,
     remote_session_description_handlers: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
     remote_ice_candidate_handlers: Mutex<BTreeMap<Peer, Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>>>,
 }
 
 impl SignalClient {
-    pub async fn new(addr: &str) -> Res<Self> {
+    pub async fn new(this_peer: Peer, addr: &str) -> Res<Self> {
         let (ws_stream, _) = connect_async(addr).await?;
         Ok(Self {
+            this_peer,
             ws_stream: AsyncMutex::new(ws_stream),
             remote_session_description_handlers: Mutex::new(BTreeMap::new()),
             remote_ice_candidate_handlers: Mutex::new(BTreeMap::new()),
@@ -82,8 +84,8 @@ impl SignalClient {
 
 #[async_trait]
 impl RTCSignalClient for SignalClient {
-    async fn send_session_description(self: Arc<Self>, peer: Peer, sdp: RTCSessionDescription) -> Res<()> {
-        let msg = SignalMessageToServer::SessionDescription(peer, sdp);
+    async fn send_session_description(self: Arc<Self>, remote_peer: Peer, sdp: RTCSessionDescription) -> Res<()> {
+        let msg = SignalMessageToServer::SessionDescription(self.this_peer, remote_peer, sdp);
         let mut ws_stream = self.ws_stream.lock().await;
         ws_stream.send(Message::Binary(serde_json::to_vec(&msg)?)).await?;
         Ok(())
@@ -95,8 +97,8 @@ impl RTCSignalClient for SignalClient {
         Ok(())
     }
 
-    async fn send_ice_candidate(self: Arc<Self>, peer: Peer, candidate: RTCIceCandidateInit) -> Res<()> {
-        let msg = SignalMessageToServer::IceCandidate(peer, candidate);
[…]

### 2022-11-19 src/transport/signalmessage.rs

@@ -9,9 +9,9 @@ pub type Peer = Hash<PublicKey>;
 
 #[derive(Clone, Debug, Serialize, Deserialize)]
 pub enum SignalMessageToServer {
-    GetPeers,
-    SessionDescription(Peer, RTCSessionDescription),
-    IceCandidate(Peer, RTCIceCandidateInit),
+    GetPeers(Peer),
+    SessionDescription(Peer, Peer, RTCSessionDescription),
+    IceCandidate(Peer, Peer, RTCIceCandidateInit),
 }
 
 #[derive(Clone, Debug, Serialize, Deserialize)]

### 2022-11-19 src/transport/webrtc.rs

@@ -39,15 +39,15 @@ pub trait RemoteAccessibleContext : HashLookup {
 #[async_trait]
 pub trait RTCSignalClient : Send + Sync {
 
-    async fn send_session_description(self: Arc<Self>, peer: Peer, sdp: RTCSessionDescription) -> Res<()>;
+    async fn send_session_description(self: Arc<Self>, remote_peer: Peer, sdp: RTCSessionDescription) -> Res<()>;
 
-    async fn on_remote_session_description(self: Arc<Self>, peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
+    async fn on_remote_session_description(self: Arc<Self>, remote_peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCSessionDescription) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
 
-    async fn send_ice_candidate(self: Arc<Self>, peer: Peer, candidate: RTCIceCandidateInit) -> Res<()>;
+    async fn send_ice_candidate(self: Arc<Self>, remote_peer: Peer, candidate: RTCIceCandidateInit) -> Res<()>;
 
-    async fn on_remote_ice_candidate(self: Arc<Self>, peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
+    async fn on_remote_ice_candidate(self: Arc<Self>, remote_peer: Peer, fun: Box<dyn Send + Sync + Fn(RTCIceCandidateInit) -> Pin<Box<dyn Send + Future<Output = Res<()>>>>>) -> Res<()>;
 
-    async fn remove_peer(self: Arc<Self>, peer: Peer) -> Res<()>;
+    async fn remove_peer(self: Arc<Self>, remote_peer: Peer) -> Res<()>;
 }
 
 struct DataChannel<T> {
@@ -224,7 +224,7 @@ async fn poll_mutex_option<T: Clone>(mutex: &Mutex<Option<T>>) -> T {
 // - lattice lookup
 // - computations? lattice computations? get cached values only
 struct PeerConnection {
-    peer: Peer,
+    remote_peer: Peer,
     accessible_context: Arc<dyn RemoteAccessibleContext>,
[…]

### 2022-11-19 src/transport/signalclient.rs
signal server stuff

@@ -16,6 +16,7 @@ use crate::error::Res;
 use crate::transport::signalmessage::{SignalMessageToClient, SignalMessageToServer, Peer};
 use crate::transport::webrtc::RTCSignalClient;
 
+
 // https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-client.rs
 
 pub struct SignalClient {
@@ -28,6 +29,7 @@ pub struct SignalClient {
 impl SignalClient {
     pub async fn new(this_peer: Peer, addr: &str) -> Res<Self> {
         let (ws_stream, _) = connect_async(addr).await?;
+        // TODO send this peer to server
         Ok(Self {
             this_peer,
             ws_stream: AsyncMutex::new(ws_stream),
@@ -85,7 +87,7 @@ impl SignalClient {
 #[async_trait]
 impl RTCSignalClient for SignalClient {
     async fn send_session_description(self: Arc<Self>, remote_peer: Peer, sdp: RTCSessionDescription) -> Res<()> {
-        let msg = SignalMessageToServer::SessionDescription(self.this_peer, remote_peer, sdp);
+        let msg = SignalMessageToServer::SessionDescription(remote_peer, sdp);
         let mut ws_stream = self.ws_stream.lock().await;
         ws_stream.send(Message::Binary(serde_json::to_vec(&msg)?)).await?;
         Ok(())
@@ -98,7 +100,7 @@ impl RTCSignalClient for SignalClient {
     }
 
     async fn send_ice_candidate(self: Arc<Self>, remote_peer: Peer, candidate: RTCIceCandidateInit) -> Res<()> {
-        let msg = SignalMessageToServer::IceCandidate(self.this_peer, remote_peer, candidate);
+        let msg = SignalMessageToServer::IceCandidate(remote_peer, candidate);
         let mut ws_stream = self.ws_stream.lock().await;
         ws_stream.send(Message::Binary(serde_json::to_vec(&msg)?)).await?;
         Ok(())

### 2022-11-19 src/transport/signalmessage.rs

@@ -9,13 +9,15 @@ pub type Peer = Hash<PublicKey>;
 
 #[derive(Clone, Debug, Serialize, Deserialize)]
 pub enum SignalMessageToServer {
+    SetPeer(Peer),
     GetPeers(Peer),
-    SessionDescription(Peer, Peer, RTCSessionDescription),
-    IceCandidate(Peer, Peer, RTCIceCandidateInit),
+    SessionDescription(Peer, RTCSessionDescription),
+    IceCandidate(Peer, RTCIceCandidateInit),
 }
 
 #[derive(Clone, Debug, Serialize, Deserialize)]
 pub enum SignalMessageToClient {
+    SetPeer,
     Peers(Vec<Peer>),
     SessionDescription(Peer, RTCSessionDescription),
     IceCandidate(Peer, RTCIceCandidateInit),

### 2022-11-19 src/transport/signalserver.rs

@@ -2,6 +2,7 @@ use std::net::SocketAddr;
 use std::collections::BTreeMap;
 use std::sync::{Arc, Mutex};
 
+use async_mutex::Mutex as AsyncMutex;
 use futures::{StreamExt};
 
 use tokio_tungstenite::{WebSocketStream, MaybeTlsStream, connect_async};
@@ -15,7 +16,7 @@ use crate::transport::signalmessage::{Peer, SignalMessageToClient, SignalMessage
 // https://github.com/snapview/tokio-tungstenite/blob/master/examples/autobahn-server.rs
 
 pub struct SignalServer {
-    peer_streams: Mutex<BTreeMap<Peer, WebSocketStream<MaybeTlsStream<TcpStream>>>>,
+    peer_streams: Mutex<BTreeMap<Peer, Arc<AsyncMutex<WebSocketStream<TcpStream>>>>>,
 }
 
 impl SignalServer {
@@ -50,13 +51,24 @@ impl SignalServer {
     }
 
     async fn handle_connection(self: Arc<Self>, peer: SocketAddr, stream: TcpStream) -> Res<()> {
-        let mut ws_stream = tokio_tungstenite::accept_async(stream).await?;
-        while let Some(msg) = ws_stream.next().await {
+        let mut ws_stream = Arc::new(AsyncMutex::new(tokio_tungstenite::accept_async(stream).await?));
+        while let Some(msg) = ws_stream.lock().await.next().await {
             match msg? {
                 Message::Binary(bs) => {
                     let msg: SignalMessageToServer = rmp_serde::from_slice(&bs)?;
-                    println!("Received a message from {}: {:?}", peer, msg);
-                    // ...
+                    match msg {
+                        SignalMessageToServer::SetPeer(this_peer) => {
+                            let mut peer_streams = self.peer_streams.lock().unwrap();
+                            peer_streams.insert(this_peer, ws_stream.clone());
+                        }
+                        SignalMessageToServer::SessionDescription(remote_peer, session_desc) => {
+                            unimplemented!()
+                        }
+                        SignalMessageToServer::IceCandidate(remote_peer, candidate) => {
+                            unimplemented!()
+                        }
+                        _ => println!("Received a message from {}: {:?}", peer, msg),
+                    }
                 }
                 other => {
                     println!("Received a message which is not binary: {:?}", other);

### 2022-11-19 src/transport/signalserver.rs
drop stream

@@ -1,3 +1,4 @@
+use core::mem::drop;
 use std::net::SocketAddr;
 use std::collections::BTreeMap;
 use std::sync::{Arc, Mutex};
@@ -52,26 +53,32 @@ impl SignalServer {
 
     async fn handle_connection(self: Arc<Self>, peer: SocketAddr, stream: TcpStream) -> Res<()> {
         let mut ws_stream = Arc::new(AsyncMutex::new(tokio_tungstenite::accept_async(stream).await?));
-        while let Some(msg) = ws_stream.lock().await.next().await {
-            match msg? {
-                Message::Binary(bs) => {
-                    let msg: SignalMessageToServer = rmp_serde::from_slice(&bs)?;
-                    match msg {
-                        SignalMessageToServer::SetPeer(this_peer) => {
-                            let mut peer_streams = self.peer_streams.lock().unwrap();
-                            peer_streams.insert(this_peer, ws_stream.clone());
+        loop {
+            let mut stream = ws_stream.lock().await;
+            let opt_msg = stream.next().await;
+            drop(stream);
+            match opt_msg {
+                None => return Ok(()),
+                Some(msg) => match msg? {
+                    Message::Binary(bs) => {
+                        let msg: SignalMessageToServer = rmp_serde::from_slice(&bs)?;
+                        match msg {
+                            SignalMessageToServer::SetPeer(this_peer) => {
+                                let mut peer_streams = self.peer_streams.lock().unwrap();
+                                peer_streams.insert(this_peer, ws_stream.clone());
+                            }
+                            SignalMessageToServer::SessionDescription(remote_peer, session_desc) => {
+                                unimplemented!()
+                            }
+                            SignalMessageToServer::IceCandidate(remote_peer, candidate) => {
+                                unimplemented!()
+                            }
+                            _ => println!("Received a message from {}: {:?}", peer, msg),
                         }
-                        SignalMessageToServer::SessionDescription(remote_peer, session_desc) => {
-                            unimplemented!()
-                        }
-                        SignalMessageToServer::IceCandidate(remote_peer, candidate) => {
-                            unimplemented!()
-                        }
-                        _ => println!("Received a message from {}: {:?}", peer, msg),
                     }
-                }
-                other => {
-                    println!("Received a message which is not binary: {:?}", other);
+                    other => {
+                        println!("Received a message which is not binary: {:?}", other);
+                    }
                 }
             }
         }

### 2022-11-19 src/transport/signalserver.rs
this peer

@@ -53,6 +53,7 @@ impl SignalServer {
 
     async fn handle_connection(self: Arc<Self>, peer: SocketAddr, stream: TcpStream) -> Res<()> {
         let mut ws_stream = Arc::new(AsyncMutex::new(tokio_tungstenite::accept_async(stream).await?));
+        let mut this_peer: Option<Peer> = None;
         loop {
             let mut stream = ws_stream.lock().await;
             let opt_msg = stream.next().await;
@@ -63,9 +64,10 @@ impl SignalServer {
                     Message::Binary(bs) => {
                         let msg: SignalMessageToServer = rmp_serde::from_slice(&bs)?;
                         match msg {
-                            SignalMessageToServer::SetPeer(this_peer) => {
+                            SignalMessageToServer::SetPeer(peer) => {
                                 let mut peer_streams = self.peer_streams.lock().unwrap();
-                                peer_streams.insert(this_peer, ws_stream.clone());
+                                peer_streams.insert(peer, ws_stream.clone());
+                                this_peer = Some(peer);
                             }
                             SignalMessageToServer::SessionDescription(remote_peer, session_desc) => {
                                 unimplemented!()

### 2022-11-19 src/transport/signalmessage.rs
session description

@@ -17,7 +17,6 @@ pub enum SignalMessageToServer {
 
 #[derive(Clone, Debug, Serialize, Deserialize)]
 pub enum SignalMessageToClient {
-    SetPeer,
     Peers(Vec<Peer>),
     SessionDescription(Peer, RTCSessionDescription),
     IceCandidate(Peer, RTCIceCandidateInit),

### 2022-11-19 src/transport/signalserver.rs

@@ -3,8 +3,9 @@ use std::net::SocketAddr;
 use std::collections::BTreeMap;
 use std::sync::{Arc, Mutex};
 
+use anyhow::anyhow;
 use async_mutex::Mutex as AsyncMutex;
-use futures::{StreamExt};
+use futures::{StreamExt, SinkExt};
 
 use tokio_tungstenite::{WebSocketStream, MaybeTlsStream, connect_async};
 use tokio::net::{TcpListener, TcpStream};
@@ -70,7 +71,16 @@ impl SignalServer {
                                 this_peer = Some(peer);
                             }
                             SignalMessageToServer::SessionDescription(remote_peer, session_desc) => {
-                                unimplemented!()
+                                let msg = SignalMessageToClient::SessionDescription(this_peer.ok_or(anyhow!("must set peer first"))?, session_desc);
+                                let opt_stream = {
+                                    let peer_streams = self.peer_streams.lock().unwrap();
+                                    peer_streams.get(&remote_peer).cloned()
+                                };
+                                if let Some(stream) = opt_stream {
+                                    let mut stream = stream.lock().await;
+                                    let bs = rmp_serde::to_vec(&msg)?;
+                                    stream.send(Message::Binary(bs)).await?;
+                                }
                             }
                             SignalMessageToServer::IceCandidate(remote_peer, candidate) => {
                                 unimplemented!()

### 2022-11-19 src/transport/signalserver.rs
ice candidate sending

@@ -83,7 +83,16 @@ impl SignalServer {
                                 }
                             }
                             SignalMessageToServer::IceCandidate(remote_peer, candidate) => {
-                                unimplemented!()
+                                let msg = SignalMessageToClient::IceCandidate(this_peer.ok_or(anyhow!("must set peer first"))?, candidate);
+                                let opt_stream = {
+                                    let peer_streams = self.peer_streams.lock().unwrap();
+                                    peer_streams.get(&remote_peer).cloned()
+                                };
+                                if let Some(stream) = opt_stream {
+                                    let mut stream = stream.lock().await;
+                                    let bs = rmp_serde::to_vec(&msg)?;
+                                    stream.send(Message::Binary(bs)).await?;
+                                }
                             }
                             _ => println!("Received a message from {}: {:?}", peer, msg),
                         }

### 2022-11-19 src/transport/signalserver.rs
consolidate send_to

@@ -52,6 +52,19 @@ impl SignalServer {
         }
     }
 
+    async fn send_to(self: &Arc<Self>, remote_peer: Peer, message: SignalMessageToClient) -> Res<()> {
+        let opt_stream = {
+            let peer_streams = self.peer_streams.lock().unwrap();
+            peer_streams.get(&remote_peer).cloned()
+        };
+        if let Some(stream) = opt_stream {
+            let mut stream = stream.lock().await;
+            let bs = rmp_serde::to_vec(&message)?;
+            stream.send(Message::Binary(bs)).await?;
+        }
+        Ok(())
+    }
+
     async fn handle_connection(self: Arc<Self>, peer: SocketAddr, stream: TcpStream) -> Res<()> {
         let mut ws_stream = Arc::new(AsyncMutex::new(tokio_tungstenite::accept_async(stream).await?));
         let mut this_peer: Option<Peer> = None;
@@ -72,27 +85,11 @@ impl SignalServer {
                             }
                             SignalMessageToServer::SessionDescription(remote_peer, session_desc) => {
                                 let msg = SignalMessageToClient::SessionDescription(this_peer.ok_or(anyhow!("must set peer first"))?, session_desc);
-                                let opt_stream = {
-                                    let peer_streams = self.peer_streams.lock().unwrap();
-                                    peer_streams.get(&remote_peer).cloned()
-                                };
-                                if let Some(stream) = opt_stream {
-                                    let mut stream = stream.lock().await;
-                                    let bs = rmp_serde::to_vec(&msg)?;
-                                    stream.send(Message::Binary(bs)).await?;
-                                }
+                                self.send_to(remote_peer, msg).await?;
                             }
                             SignalMessageToServer::IceCandidate(remote_peer, candidate) => {
                                 let msg = SignalMessageToClient::IceCandidate(this_peer.ok_or(anyhow!("must set peer first"))?, candidate);
-                                let opt_stream = {
-                                    let peer_streams = self.peer_streams.lock().unwrap();
-                                    peer_streams.get(&remote_peer).cloned()
-                                };
-                                if let Some(stream) = opt_stream {
-                                    let mut stream = stream.lock().await;
-                                    let bs = rmp_serde::to_vec(&msg)?;
-                                    stream.send(Message::Binary(bs)).await?;
-                                }
+                                self.send_to(remote_peer, msg).await?;
                             }
                             _ => println!("Received a message from {}: {:?}", peer, msg),
                         }

### 2022-11-19 src/transport/signalserver.rs
ref arc

@@ -65,7 +65,7 @@ impl SignalServer {
         Ok(())
     }
 
-    async fn handle_connection(self: Arc<Self>, peer: SocketAddr, stream: TcpStream) -> Res<()> {
+    async fn handle_connection(self: &Arc<Self>, peer: SocketAddr, stream: TcpStream) -> Res<()> {
         let mut ws_stream = Arc::new(AsyncMutex::new(tokio_tungstenite::accept_async(stream).await?));
         let mut this_peer: Option<Peer> = None;
         loop {

### 2022-11-19 src/transport/signalclient.rs
send set peer

@@ -28,8 +28,8 @@ pub struct SignalClient {
 
 impl SignalClient {
     pub async fn new(this_peer: Peer, addr: &str) -> Res<Self> {
-        let (ws_stream, _) = connect_async(addr).await?;
-        // TODO send this peer to server
+        let (mut ws_stream, _) = connect_async(addr).await?;
+        ws_stream.send(Message::Binary(serde_json::to_vec(&SignalMessageToServer::SetPeer(this_peer))?)).await?;
         Ok(Self {
             this_peer,
             ws_stream: AsyncMutex::new(ws_stream),

### 2022-11-20 src/db.rs
join doesnt need to return option

@@ -442,12 +442,7 @@ impl<C: TaggedMapping + 'static, L: TaggedMapping + 'static, LC: TaggedMapping +
                 if tr_other == store_merkle.value {
                     return Ok(Some(store_merkle_hash));
                 }
-                let joined = self.lat_lib.clone().join(key, &store_merkle.value, &tr_other, self.clone()).await?;
-                if joined.is_none() {
-                    // this probably shouldn't happen...
-                    return Ok(Some(store_merkle_hash));
-                }
-                joined.unwrap()
+                self.lat_lib.clone().join(key, &store_merkle.value, &tr_other, self.clone()).await?
             }
         };
         let lat_lib = self.lat_lib.clone();

### 2022-11-20 src/js/js_library.rs

@@ -310,11 +310,11 @@ impl LatticeLibrary<JsMapping, JsMapping, JsMapping> for JsLibrary {
         }
     }
 
-    async fn join(self: Arc<Self>, key: &JsValue, a: &JsValue, b: &JsValue, ctx: Arc<dyn LatticeMutContext<JsMapping, JsMapping, JsMapping>>) -> Res<Option<JsValue>> {
+    async fn join(self: Arc<Self>, key: &JsValue, a: &JsValue, b: &JsValue, ctx: Arc<dyn LatticeMutContext<JsMapping, JsMapping, JsMapping>>) -> Res<JsValue> {
         let dyn_ctx = DynContext::LatticeMut(ctx);
         let ctxid = self.get_ctx_id(&dyn_ctx);
         if let LibraryResult::Join(result) = self.do_query(LibraryQuery::Join(key.0.clone(), a.0.clone(), b.0.clone(), ctxid)).await? {
-            Ok(result.map(JsValue))
+            Ok(JsValue(result))
         } else {
             bail!("Unexpected result from join")
         }

### 2022-11-20 src/js/runtime_channel.rs

@@ -40,7 +40,7 @@ pub enum LibraryQuery {
 pub enum LibraryResult {
     EvalComputation(JsValue),
     CheckElem,
-    Join(Option<JsValue>),
+    Join(JsValue),
     Transport(Option<JsValue>),
     EvalLatComputation(JsValue),
 }
@@ -273,7 +273,7 @@ impl RuntimeState {
                                 self.register_call_function(query_id, "hash_put", vec![key, value, JsValue::from(ctxid)], |_| Ok(LibraryResult::CheckElem));
                             },
                             LibraryQuery::Join(key, value1, value2, ctxid) => {
-                                self.register_call_function(query_id, "lattice_join", vec![key, value1, value2, JsValue::from(ctxid)], |res| Ok(LibraryResult::Join(js_to_option(res)?)));
+                                self.register_call_function(query_id, "lattice_join", vec![key, value1, value2, JsValue::from(ctxid)], |res| Ok(LibraryResult::Join(res)));
                             },
                             LibraryQuery::Transport(key, value, old_ctxid, new_ctxid) => {
                                 self.register_call_function(query_id, "transport", vec![key, value, JsValue::from(old_ctxid), JsValue::from(new_ctxid)], |res| Ok(LibraryResult::Transport(js_to_option(res)?)));

### 2022-11-20 src/lattice.rs

@@ -64,7 +64,7 @@ pub trait LatticeLibrary<C: TaggedMapping + 'static, L: TaggedMapping + 'static,
         bail!("check_elem not implemented")
     }
 
-    async fn join(self: Arc<Self>, _key: &L::Key, _a: &L::Value, _b: &L::Value, _ctx: Arc<dyn LatticeMutContext<C, L, LC>>) -> Res<Option<L::Value>> {
+    async fn join(self: Arc<Self>, _key: &L::Key, _a: &L::Value, _b: &L::Value, _ctx: Arc<dyn LatticeMutContext<C, L, LC>>) -> Res<L::Value> {
         bail!("join not implemented")
     }
 

### 2022-11-20 paper/hl_paper.tex
paper v2

--- 2022-10-04 paper/paper.tex
@@
 
-\title{Semilattice Graphs for Distributed Databases}
+\title{Hashlattice: A distributed system for monotonic peer-to-peer synchronization}
 \author{Jessica Taylor}
+% Astra???
 \date{\today}
@@
   \section{Introduction}
+    The Internet, while it consists of a set of interoperable protocols, is in practice quite centralized.
+People spend most of their time on large websites that serve many users and which must be trusted to properly handle user data.
+This enables censorship and introduces security risks.
+
+    Peer-to-peer (P2P) systems are an alternative to centralized systems.
+In a P2P system, users are connected to each other directly, and there is no central authority.
+This allows users to communicate without trusting a third party.
+However, P2P systems are difficult to build because they must be able to handle a large number of users while preserving security and robustness.
+
+    Prior art for P2P systems includes BitTorrent, a file sharing system, IPFS, a file indexing and transfer protocol, and Bitcoin, a cryptocurrency.
+BitTorrent magnet links and IPFS share the property that data is indexed by its hash code.
+This enables a user not possessing the data to download it from the network and verify its correctness.
+However, hash codes are static, so applications with state changes cannot straightforwardly be implemented on these protocols.
+Bitcoin implements state changes by having every ``full node'' download and check every transaction to ensure its correctness, computing the new state from the previous state and any ensuing transactions.
+However, Bitcoin is not scalable because of the requirement for every full node to download and check every transaction.
+If $n$ users each make $k$ transactions each, then the total bandwidth and data storage requirement is at least $O(kn^2)$.
+As a result, Bitcoin has limited throughput, high transaction fees, and requires a large amount of bandwidth and disk space to run.
+
+    We propose a new P2P system, HashLattice, which combines the hash code indexed data storage capabilities of IPFS with state changes using algebraic semilattices.
+This system allows users to scalably upload new data to the network and update it over time in a way that can be verified by other users.
+
+    Prior work includes conflict-free replicated data types (CRDTs), a class of data structure that allows multiple servers cooperating to run the same application to independently modify the data and resolve their differing updates.
+CRDTs have been used in industry, but have not to our knowledge been used to implement a P2P Internet.
+
+  \section{Notation}
+
+    We use $\mathrm{Unit}$ to indicate the set containinly only one value, $()$.
+
+    We use $\mathrm{Option}(S)$ for some set $S$ to indicate the sum set $\{ \mathrm{Some}(s) ~ | ~ s \in S \} \cup \mathrm{None}\}$, as in Rust.
+
+    \newcommand{\partialto}{\hookrightarrow}
+
+    We use $A \partialto B$ to indicate the type of partial functions from $A$ to $B$. Conceptually, these are functions that may either return a result or throw an error.
+
+  \section{Data and data computations}
+
+    As in IPFS, raw data is stored in the network indexed by its hash code.
+This enables users to download data from the network and verify its correctness.
+\emph{Data computations} consist of code that either produces either a fixed, immutable value or an error.
+These computations may query data by their hash code, and my also request the result of other data computations.
+The results of data computations are in general cached (due to their immutability), enabling multiple data computations to re-use the results of a given data computation without running it twice.
+
+  \section{Semilattices}
+
     A \emph{semilattice} is a set $L$ with a binary operation $\vee$ (``join'') that is commutative ($x \vee y = y \vee x$), associative ($(x \vee y) \vee z = x \vee (y \vee z)$), and idempotent ($x \vee x = x$).
+(This is in standard literature called a ``join semilattice''; we do not consider the dual ``meet semilattices'' and so use ``semilattice'' to mean ``join semilattice'' henceforth.)
 
@@
 
-    \section{Semilattices as a category}
+    Any total order corresponds to a semilattice, with $x \vee y = \mathrm{max}(x, y)$.
+For example, a total order could be created over timestamped messages signed by a particular public key, where later messages are considered greater in the order.
+The corresponding semilattice will select the latest signed message by this public key that is available on the network.
+This enables users to issue ``status updates'' that are distributed to other users, such that the other users can be aware of the original user's latest status update, corresponding to their current status.
+
+    \newcommand{\checkelem}{\ensuremath{\mathsf{check\_elem}}}
+    \newcommand{\join}{\ensuremath{\mathsf{join}}}
+
+    A semilattice is specified in HashLattice by two functions.
+The first, \checkelem, takes an element $x$ and produces an error if the element is not a member of the semilattice.
+The second, \join, takes two valid elements $x$ and $y$ and produces the join $x \vee y$.
+These functions may look up pieces of data by their hash codes and call data computations.
+
+    In a P2P network context, each node stores data, data computation results, and maximum values of semilattices.
+They share with each other data and semilattice maxima.
+They don't share the results of data computations, since these are in general unverifiable without downloading the correspoding data and running the corresponding computation.
+
+  \section{Dependent semilattices}
+
+    As an extension to this system, we may also enable semilattices to depend on the maximum values of other semilattices.
+While this introduces additional complexity, this feature enables greater composability, as it enables P2P applications running on HashLattice may query the semilattice data of other applications.
+
+    \subsection{Category theoretic interpretation of semilattices}
 
       We discuss category theoretic language for semilattices to derive a notion of dependency between semilattices.
-In the next section, we drop the category-theoretic language, so it is not strictly necessary to understand category theory to understand the result, although it helps with the derivation.
+After this subsection, we drop the category-theoretic language, so it is not strictly necessary to understand category theory to understand the result, although it helps with the derivation.
 
@@
 
-      The total space $\mathrm{SemiTot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
+      The total space $\mathrm{SemiLatTot}(L, g)$ is a preorder whose elements consist of pairs $(x, x')$ where $x : L$ and $x' : g(x)$.
 We consider $(x, x') \leq (y, y')$ when $x \leq y$ and $g(x \leq y)(x') \leq y'$.
@@
       \begin{theorem}
-        The total space $\mathrm{SemiTot}(L, g)$ of a semilattice $L$ and a semilattice fibration $g$ is a semilattice.
+        The total space $\mathrm{SemiLatTot}(L, g)$ of a semilattice $L$ and a semilattice fibration $g$ is a semilattice.
       \end{theorem}
@@
 
-      Note that the projection $\pi_1 : \mathrm{Tot}(L, g) \rightarrow L$ is a semilattice morphism.
+      Note that the projection $\pi_1 : \mathrm{SemiLatTot}(L, g) \rightarrow L$ is a semilattice morphism.
 
-    \section{De-categorizing semilattice fibrations}
+    \subsection{De-categorizing semilattice fibrations}
 
@@
 
-      We write $\mathrm{SemiFibration}(L)$ for the set of semilattice fibrations on $L$.
+      We write $\mathrm{SemiLatFib}(L)$ for the set of semilattice fibrations on $L$.
 
@@
 
-      The total space $\mathrm{SemiTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$.
+      The total space $\mathrm{SemiLatTot}(L, l, t)$ is then a semilattice whose elements are pairs $(x, x')$ where $x : L$ and $x' : l(x)$.
 The join is defined as $(x, x') \vee (y, y') := (x \vee y, t(x, x \vee y)(x') \vee t(y, x \vee y)(y'))$.
@@
 
-    \section{Composing semilattice fibrations}
+    \subsection{Composing semilattice fibrations}
 
-      Semilattice fibrations can be composed iteratively. Composing finitely is straightforward, while composing infinitely requires more work.
+      Semilattice fibrations can be composed iteratively. Composing finitely is straightforward, while omposing infinitely requires more work.
 
@@
         L_0 &: \mathrm{Semilattice} & \\
-        (l_n, t_n) &: \mathrm{SemiFibration}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
-        L_n &:= \mathrm{SemiTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
+        (l_n, t_n) &: \mathrm{SemiLatFib}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
+        L_n &:= \mathrm{SemiLatTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
       \end{align*}
@@
         L_{0, 0} &: \mathrm{Semilattice} & \\
-        (l_{u, n}, t_{u, n}) &: \mathrm{SemiFibration}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
-        L_{u, n} &:= \mathrm{SemiTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        (l_{u, n}, t_{u, n}) &: \mathrm{SemiLatFib}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
+        L_{u, n} &:= \mathrm{SemiLatTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
         L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
@@
       where $L_{u, \omega}$ is constructed the same way as $L_{\omega}$ above, as the limit of $L_{u, n}$ as $n$ approaches $\infty$.
-Inductively, each $L_{u, n}$ is a semilattice for natural $0 \leq u < n_u, n \geq 0$; this lattice's ordinal corresponds to $u \omega + n$.
+Inductively, each $L_{u, n}$ is a semilattice for naturals $0 \leq u < n_u$ and $n \geq 0$; this lattice's ordinal corresponds to $u \omega + n$.
 While in theory it may be possible to reach higher ordinals $\omega^2$ and beyond, we consider this beyond the scope of this work.
 
-    \section{Semilattice Graphs}
+  \section{HashLattice libraries}
 
-      In terms of implementation, it is by default infeasible to construct an element of the semilattice $L_{0, \omega}$, as it is infinite.
-To ameliorate this problem, we consider semilattices with a bottom element $\bot$ satisfying $\bot \vee x = x$ for each $x$ in the semilattice.
-We write $\mathrm{Semilattice}_\bot$ for the type of semilattices with $\bot$; $\mathrm{SemiLFibration}_\bot(L)$ for the type of bottom-preserving semilattice fibrations on $L : \mathrm{Semilattice}_\bot$; and $\mathrm{SemiLTot}_\bot(L, l, t) : \mathrm{Semilattice}_\bot$ for the total space the semilattice $L : \mathrm{Semilattice}_\bot$ and fibration $(l, t) : \mathrm{SemiLFibration}_\bot$.
+    Using the above theory of dependent semilattices, we may specify what a HashLattice library consists of.
+HashLattice libraries provide definitions of data computations, semilattices, and semilattice computations.
+Data computations may depend on data and data computations; semilattices and semilattice computations may depend on data, data computations, semilattice values, and semilattice computations.
 
-      Using the above definition of semilattice universes, we can now explicitly represent elements of $L_{u, n}$ for which only a finite number of entries are not $\bot$.
-Joins are straightforward since if one element of the joined universe structure is $\bot$, the join equals the other element.
-The bottoms of the total space and $\omega$ total spaces are trivial, containing only $\bot$.
+    \subsection{Contexts}
 
-      Let $K$ be a countable, well-ordered set of keys whose ordinal is no more than $\omega^2$.
-Let $V$ be a set of values.
-Since keys correspond with ordinals in $\omega^2$, we can write any key as $u\omega + n$; let $u(k)$ and $n(k)$ be $u$ and $n$ in this term respectively.
-We define a \emph{semilattice graph} to consist of:
-      \begin{align*}
-        L_{0, 0} &:= \{ \bot \} \\
-        (l_k, t_k) &: \mathrm{SemiLFibration}_\bot(L_{u(k), n(k)-1}) & \text{ for each $k \in K$ } \\
-        l_k(x) \subseteq V &\text{ for each $k \in K$ and $x \in L_{u(k), n(k)}$} \\
-        L_{u, n} &:= \mathrm{SemiLTot}_\bot(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
-        L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
-      \end{align*}
+      HashLattice library functions depend on \emph{contexts}, which differ depending on the type of library function specified.
 
-      Note that each key corresponds to a semilattice fibration from a multiverse containing lower keys.
-We can re-interpret each element of the lattice $L_{u(k), n(k)}$ as mappings from keys less no higher than $k$ (written as $K_{\leq k}$) to $V$, which doesn't affect the semilattice structure.
-Accordingly, $l_k$ and $t_k$ are re-interpreted as taking mappings as arguments.
+      A \emph{data computation context} provides the following methods:
 
-      It is somewhat unwieldy to define $l_k$ and $t_k$ in implementation, since they depend on a possibly-infinite mapping.
-Accordingly, we assume that these operations only depend on a defined set of values.
-Let $d_k^\star : 2^{K_{<k}}$ be a set of lattice dependencies, and $d_{k}(v) : 2^{K_{<k}}$ be a set of value dependencies for $v \in V$.
-Both dependency sets must only include keys less than $k$.
-We say $(l_k, d_k)$ have dependencies according to $d_{k, l}, d_{k, v}$ if:
+      \newcommand{\hashlookup}{\ensuremath{\mathsf{hash\_lookup}}}
+      \newcommand{\datacomputationlookup}{\ensuremath{\mathsf{data\_computation\_lookup}}}
 
       \begin{itemize}
-        \item $\mathrm{Elems}(l_k(m))$ only depends on $m(k')$ for each $k' \in d_k^\star$.
-        \item $\bot_{l_k(m})$ only depends on $m(k')$ for each $k' \in d_k^\star$.
-        \item $t_k(m, m')(v)$ only depends on $m(k')$ and $m'(k')$ for each $k' \in d_k^\star \cup d_k(v)$.
-        \item $x \vee_{l_k(m)} y$ only depends on $m(k')$ for each $k' \in d_k^\star \cup d_k(x) \cup d_k(y)$.
+        \item $\hashlookup : \mathrm{HashCode} \partialto \mathrm{Bytes}$: Given a hash code, return the corresponding data value. Throws an error if the data value with this hash code is unavailable.
+        \item $\datacomputationlookup : \mathrm{DataComputation} \partialto \mathrm{DataComputationValue}$: Given a data computation, return the corresponding data value.
+Throws an error if evaluating the computation throws an error.
       \end{itemize}
 
-      To ensure dependency sets don't grow unexpectedly, we require:
+      A \emph{immutable lattice context} provides the above methods and the following additional methods:
+
+      \newcommand{\latticelookup}{\ensuremath{\mathsf{lattice\_lookup}}}
+      \newcommand{\latticecomputationlookup}{\ensuremath{\mathsf{lattice\_computation\_lookup}}}
+
       \begin{itemize}
-        \item $d_k(t_k(m, m')(v)) \subseteq d_k(v)$.
-        \item $d_k(x \vee_{l_k(m)} y) \subseteq d_k(x) \cup d_k(y)$.
+        \item $\latticelookup : \mathrm{Semilattice} \to \mathrm{Option}(SemilatticeValue)$: Given a semilattice, return the corresponding maximum semilattice value, or $\mathrm{None}$ if there is no known valid value.
+        \item $\latticecomputationlookup : \mathrm{LatticeComputation} \partialto \mathrm{LatticeComputationValue}$: Given a semilattice computation, return the corresponding result.
+Throws an error if evaluating the computation throws an error.
       \end{itemize}
 
-    \section{Semilattice Graphs}
+      A \emph{mutable lattice context} provides all the above methods and the following additional method:
 
-        A common problem in networked computing is dependency managament.
-A software package may depend on a set of other packages, each of which depends on another set of packages.
-We consider extending semilattices to semilattice graphs to handle dependency management.
-
-        A \emph{semilattice graph} consists of:
+      \newcommand{\hashput}{\ensuremath{\mathsf{hash\_put}}}
 
       \begin{itemize}
-            \item A set of keys $K$.
-            \item A nonempty set of values $V$.
-            \item A dependency function $d : K \times V \rightarrow 2^K$ that maps each key and value to a finite set of keys, its dependencies.
-            \item A transport function $t : \Pi (k : K, v : V), (d(k, v) \rightarrow V)^2 \rightarrow V$ that transports a value when its dependencies change.
-            \item A join function $j : \Pi (k : K, v_1, v_2 : V), (d(k, v_1) \cup d(k, v_2) \rightarrow V) \rightarrow V$ that maps two values to a value given their dependencies.
+        \item $\hashput : \mathrm{Bytes} \to \mathrm{HashCode}$: Stores a given data value in the database, returning the corresponding hash code.
       \end{itemize}
 
-        We write $v_1 \vee_{k, g} v_2 := j(k, v_1, v_2, g|_{d(k, v_1) \cup d(k, v_2)})$, and $t_{k, g_1, g_2}(v) := t(k, v, g_1|_{d(k, v)}, g_2|_{d(k, v)})$.
+    \subsection{Libraries}
 
-        The semilattice graph must satisfy the following properties:
+      A \emph{data computation library} provides the following methods:
+
+      \newcommand{\evaldatacomputation}{\ensuremath{\mathsf{eval\_data\_computation}}}
 
       \begin{itemize}
-            \item $d$ must be well-founded; that is, there are no infinite dependency chains $k_1, k_2, \ldots \in K$ and $v_1, v_2, \ldots \in V$ such that $k_{i+1} \in d(k_i, v_i)$ for $i = 1, 2, \ldots$
-            \item $t$ must not grow dependencies: for $k : K, v : V, g_1, g_2 : d(k, v) \rightarrow V$, we must have $d(k, t(k, v, g_1, g_2)) \subseteq d(k, v)$.
-            \item $j$ must not grow dependencies; for any $k : K, v_1 : V, v_2: V, g: d(k, v_1) \cup d(k, v_2) \rightarrow V$, we must have $d(k, j(k, v_1, v_2, g)) \subseteq d(k, v_1) \cup d(k, v_2)$.
-            \item $t$ must be an identity when dependencies are equal: for $k : K, v : V, g : d(k, v) \rightarrow V$, we must have $t(k, v, g, g) = v$.
-            \item $t$ must not be path-dependent: for $k : K, v : V, g_1, g_2, g_3: d(k, v) \rightarrow V$, we must have $t_{k, g_2, g_3} \circ t_{k, g_1, g_2} = t_{k, g_1, g_3}$.
-            \item $t$ must distribute over $j$: for $k : K, v_1, v_2 : V, g_1, g_2 : d(k, v_1) \cup d(k, v_2) \rightarrow V$, we must have $t_{k, g_1, g_2}(v_1 \vee_{k, g_1} v_2) = t_{k, g_1, g_2}(v_1) \vee_{k, g_2} t_{k, g_1, g_2}(v_2)$.
-            \item For any $k: K, g : K \rightarrow V$, we must have that $\vee_{k, g}$ is commutative, associative, and idempotent.
+        \item $\evaldatacomputation : \mathrm{DataComputationContext} \times \mathrm{DataComputation} \partialto \mathrm{DataComputationValue}$: Given a data computation, return the corresponding data value.
+This may query a data computation context.
       \end{itemize}
 
-        It is now possible to define a semilattice over a semilattice graph.
-This lattice's items are functions $g : K \rightarrow V$ (called \emph{mappings}).
-The join operation is defined as follows:
+      A \emph{semilattice library} provides the following methods:
 
-        $$ (g_1 \vee g_2)(k) := t_{k, g_1, g_1 \vee g_2}(g_1(k)) \vee_{k, g_1 \vee g_2} t_{k, g_2, g_1 \vee g_2}(g_2(k))$$.
-
-        We must now show that $\vee$ is commutative, associative, and idempotent, inductively.
-For any natural $n$ we say $\vee$ (applied to semilattice graphs) forms a semilattice up to $n$ if, for any $k : K$ whose longest dependency chain is no longer than $n$, we have:
+      \newcommand{\transport}{\ensuremath{\mathsf{transport}}}
+      \newcommand{\evallatticecomputation}{\ensuremath{\mathsf{eval\_lattice\_computation}}}
 
       \begin{itemize}
-          \item \emph{Commutative}: $(g_1 \vee g_2)(k) = (g_2 \vee g_1)(k)$
-          \item \emph{Associative}: $((g_1 \vee g_2) \vee g_3)(k) = (g_1 \vee (g_2 \vee g_3))(k)$
-          \item \emph{Idempotent}: $(g \vee g)(k) = g(k)$
-        \end{itemize}
-
-        We note that if $\vee$ forms a semilattice up to all $n$, then it forms a semilattice.
-We now note that if $\vee$ forms a semilattice up to $n$, then it forms a semilattice up to $n+1$, inductively:
-
-        \begin{itemize}
-          \item \emph{Commutative}:
-            \begin{align*}
-              (g_1 \vee g_2)(k) &= t_{k, g_1, g_1 \vee g_2}(g_1(k)) \vee_{k, g_1 \vee g_2} t_{k, g_2, g_1 \vee g_2}(g_2(k)) \\
-              &= t_{k, g_2, g_2 \vee g_1}(g_2(k)) \vee_{k, g_2 \vee g_1} t_{k, g_1, g_2 \vee g_1}(g_1(k)) \\
-              &= (g_2 \vee g_1)(k)
-            \end{align*}
-          \item \emph{Associative}:
-            \begin{align*}
-              &((g_1 \vee g_2) \vee g_3))(k) \\
-              =& t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}((g_1 \vee g_2)(k)) \vee_{k, g_1 \vee g_2 \vee g_3} t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
-              =& t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}\left(t_{k, g_1, g_1 \vee g_2}(g_1(k)) \vee_{k, g_1 \vee g_2} t_{k, g_2, g_1 \vee g_2}(g_2(k))\right) \\
-               & \vee_{k, g_1 \vee g_2 \vee g_3} t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
-              =& t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}(t_{k, g_1, g_1 \vee g_2}(g_1(k))) \vee_{k, g_1 \vee g_2 \vee g_3} \\
-               & t_{k, g_1 \vee g_2, g_1 \vee g_2 \vee g_3}(t_{k, g_2, g_1 \vee g_2}(g_2(k))) \vee_{k, g_1 \vee g_2 \vee g_3} \\
-               & t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
-              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3}
-                 t_{k, g_2, g_1 \vee g_2 \vee g_3}(g_2(k)) \vee_{k, g_1 \vee g_2 \vee g_3}
-                 t_{k, g_3, g_1 \vee g_2 \vee g_3}(g_3(k)) \\
-              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3} \\
-               & t_{k, g_2 \vee g_3, g_1 \vee g_2 \vee g_3}(t_{k, g_2, g_2 \vee g_3}(g_2(k))) \vee_{k, g_1 \vee g_2 \vee g_3} \\
-               & t_{k, g_2 \vee g_3, g_1 \vee g_2 \vee g_3}(t_{k, g_3, g_2 \vee g_3}(g_3(k))) \\
-              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3} \\
-               & t_{k, g_2 \vee g_3}(t_{k, g_2, g_2 \vee g_3}(g_2(k)) \vee_{k, g_2 \vee g_3} t_{k, g_3, g_2 \vee g_3}(g_3(k))) \\
-              =& t_{k, g_1, g_1 \vee g_2 \vee g_3}(g_1(k)) \vee_{k, g_1 \vee g_2 \vee g_3} t_{k, g_2 \vee g_3}((g_2 \vee g_3)(k)) \\
-              =& (g_1 \vee (g_2 \vee g_3))(k)
-            \end{align*}
-          \item \emph{Idempotent}:
-            \begin{align*}
-              (g \vee g)(k) &= t_{k, g, g \vee g}(g(k)) \vee_{k, g \vee g} t_{k, g, g \vee g}(g(k)) \\
-              &= t_{k, g, g}(g(k)) \vee_{k, g} t_{k, g, g}(g(k)) \\
-              &= g(k)
-            \end{align*}
+        \item $\checkelem : \mathrm{ImmutableLatticeContext} \times \mathrm{Semilattice} \times \mathrm{SemilatticeValue} \partialto \mathrm{Unit}$: Given a semilattice and a semilattice value, throw an error if the value is not an element of the semilattice.
+This may query an immutable lattice context.
+        \item $\join : \mathrm{MutableLatticeContext} \times \mathrm{Semilattice} \times \mathrm{SemilatticeValue} \times \mathrm{SemilatticeValue} \partialto \mathrm{SemilatticeValue}$: Given a semilattice and two semilattice values, return the join of the two values.
+This may query and update a mutable lattice context.
+        \item $\transport : \mathrm{ImmutableLatticeContext} \times \mathrm{MutableLatticeContext} \times \mathrm{Semilattice} \times \mathrm{SemilatticeValue} \partialto \mathrm{Option}{\mathrm{SemilatticeValue}}$: Given a semilattice and a semilattice value, transport it from a value in the dependent semilattice in the context of the first context to a value in the dependent semilattice in the context of the second context, or $\mathrm{None}$ if there is no such translation.
+        \item $\evallatticecomputation : \mathrm{MutableLatticeContext} \times \mathrm{SemilatticeComputation} \partialto \mathrm{SemilatticeComputationValue}$: Given a semilattice computation, return the corresponding result.
+This may query and update a mutable lattice context.
       \end{itemize}

### 2022-11-20 paper/hl_paper.tex
paper update

@@
       \begin{itemize}
-        \item $\latticelookup : \mathrm{Semilattice} \to \mathrm{Option}(SemilatticeValue)$: Given a semilattice, return the corresponding maximum semilattice value, or $\mathrm{None}$ if there is no known valid value.
+        \item $\latticelookup : \mathrm{Semilattice} \to \mathrm{Option}(\mathrm{SemilatticeValue})$: Given a semilattice, return the corresponding maximum semilattice value, or $\mathrm{None}$ if there is no known valid value.
         \item $\latticecomputationlookup : \mathrm{LatticeComputation} \partialto \mathrm{LatticeComputationValue}$: Given a semilattice computation, return the corresponding result.
@@
 This may query and update a mutable lattice context.
-        \item $\transport : \mathrm{ImmutableLatticeContext} \times \mathrm{MutableLatticeContext} \times \mathrm{Semilattice} \times \mathrm{SemilatticeValue} \partialto \mathrm{Option}{\mathrm{SemilatticeValue}}$: Given a semilattice and a semilattice value, transport it from a value in the dependent semilattice in the context of the first context to a value in the dependent semilattice in the context of the second context, or $\mathrm{None}$ if there is no such translation.
+        \item $\transport : \mathrm{ImmutableLatticeContext} \times \mathrm{MutableLatticeContext} \times \mathrm{Semilattice} \times \mathrm{SemilatticeValue} \partialto \mathrm{Option}(\mathrm{SemilatticeValue})$: Given a semilattice and a semilattice value, transport it from a value in the dependent semilattice in the context of the first context to a value in the dependent semilattice in the context of the second context, or $\mathrm{None}$ if there is no such translation.
+For non-dependent semilattices, this may simply return the original value, as no transport is necessary.
         \item $\evallatticecomputation : \mathrm{MutableLatticeContext} \times \mathrm{SemilatticeComputation} \partialto \mathrm{SemilatticeComputationValue}$: Given a semilattice computation, return the corresponding result.
@@
 
+  \section{HashLattice databases}
+
+    A HashLattice database stores four types of values in its database: data values, data computations, semilattice values, and semilattice computations.
+The database tracks dependencies between these different values.
+Data computations may depend on data values and data computations; semilattice values and semilattice computation values may depend on values of any type.
+
+    Dependencies are defined as follows:
+    \begin{itemize}
+      \item A data computation depends on a data value if the corresponding \evaldatacomputation function calls \hashlookup for the data value's hash code.
+      \item A data computation depends on another data computation if the corresponding \evaldatacomputation function calls \datacomputationlookup for the other data computation.
+      \item A semilattice value depends on a data value if the corresponding \checkelem function calls \hashlookup for the data value's hash code.
+      \item A semilattice value depends on a data computation if the corresponding \checkelem function calls \datacomputationlookup for the data computation.
+      \item A semilattice value depends on another semilattice value if the corresponding \checkelem function calls \latticelookup for the other semilattice value's semilattice.
+      \item A semilattice value depends on a semilattice computation if the corresponding \checkelem function calls \latticecomputationlookup for the semilattice computation.
+      \item A semilattice computation depends on a data value if the corresponding \evallatticecomputation function calls \hashlookup for the data value's hash code.
+      \item A semilattice computation depends on a data computation if the corresponding \evallatticecomputation function calls \datacomputationlookup for the data computation.
+      \item A semilattice computation depends on a semilattice value if the corresponding \evallatticecomputation function calls \latticelookup for the semilattice value's semilattice.
+      \item A semilattice computation depends on another semilattice computation if the corresponding \evallatticecomputation function calls \latticecomputationlookup for the other semilattice computation.
+    \end{itemize}
+
+    In general, results of computations are cached, so data computations are not re-run (unless the cache is cleared), and semilattice computations are only re-run if their dependencies have changed.
+
+    For semilattice values and semilattice computations, the database stores not only the value, but also a directed acyclic graph Merkle tree of their dependencies on semilattices or semilattice computations.
+This enables synchronization between different databases, since it is possible to check if any dependencies differ before joining the values.
+Synchronization proceeds recursively; within the \join call that merges two semilattice values, the \latticelookup and \latticecomputationlookup functions recursively compute a merged result from the two databases, using the Merkle trees as a frozen pointer to the relevant subset of each database at the time of the merge.
+
+    When one database has a value for a given semilattice and the other doesn't, the only value is used as the join.
+Conceptually, we may extend a semilattice over $L$ to a semilattice over $\mathrm{Option}(L)$ as follows:
+
+    \begin{itemize}
+      \item $\mathrm{None} \vee y = y$
+      \item $x \vee \mathrm{None} = x$
+      \item $\mathrm{Some}(x) \vee \mathrm{Some}(y) = \mathrm{Some}(x \vee y)$
+    \end{itemize}
+
+    This interpretation of option semilattices makes more sense of why the \transport function returns an option: a value in an original semilattice may be transported to $\mathrm{None}$ in the updated semilattice.
+
 \end{document}

### 2022-11-20 paper/hl_paper.tex
fix spacing

@@
 \usepackage{amssymb}
+\usepackage{xspace}
 \newtheorem{theorem}{Theorem}
@@
 \begin{document}
+  \newcommand{\checkelem}{\ensuremath{\mathsf{check\_elem}}}
+  \newcommand{\join}{\ensuremath{\mathsf{join}}}
+  \newcommand{\partialto}{\hookrightarrow}
+  \newcommand{\hashlookup}{\ensuremath{\mathsf{hash\_lookup}}}
+  \newcommand{\datacomputationlookup}{\ensuremath{\mathsf{data\_computation\_lookup}}}
+  \newcommand{\latticelookup}{\ensuremath{\mathsf{lattice\_lookup}}}
+  \newcommand{\latticecomputationlookup}{\ensuremath{\mathsf{lattice\_computation\_lookup}}}
+  \newcommand{\hashput}{\ensuremath{\mathsf{hash\_put}}}
+  \newcommand{\evaldatacomputation}{\ensuremath{\mathsf{eval\_data\_computation}}}
+  \newcommand{\transport}{\ensuremath{\mathsf{transport}}}
+  \newcommand{\evallatticecomputation}{\ensuremath{\mathsf{eval\_lattice\_computation}}}
+
   \maketitle
@@
 
-    \newcommand{\partialto}{\hookrightarrow}
-
     We use $A \partialto B$ to indicate the type of partial functions from $A$ to $B$. Conceptually, these are functions that may either return a result or throw an error.
@@
 This enables users to issue ``status updates'' that are distributed to other users, such that the other users can be aware of the original user's latest status update, corresponding to their current status.
-
-    \newcommand{\checkelem}{\ensuremath{\mathsf{check\_elem}}}
-    \newcommand{\join}{\ensuremath{\mathsf{join}}}
 
@@
 
-      \newcommand{\hashlookup}{\ensuremath{\mathsf{hash\_lookup}}}
-      \newcommand{\datacomputationlookup}{\ensuremath{\mathsf{data\_computation\_lookup}}}
-
       \begin{itemize}
@@
       A \emph{immutable lattice context} provides the above methods and the following additional methods:
-
-      \newcommand{\latticelookup}{\ensuremath{\mathsf{lattice\_lookup}}}
-      \newcommand{\latticecomputationlookup}{\ensuremath{\mathsf{lattice\_computation\_lookup}}}
 
@@
 
-      \newcommand{\hashput}{\ensuremath{\mathsf{hash\_put}}}
-
       \begin{itemize}
@@
 
-      \newcommand{\evaldatacomputation}{\ensuremath{\mathsf{eval\_data\_computation}}}
-
       \begin{itemize}
@@
       A \emph{semilattice library} provides the following methods:
-
-      \newcommand{\transport}{\ensuremath{\mathsf{transport}}}
-      \newcommand{\evallatticecomputation}{\ensuremath{\mathsf{eval\_lattice\_computation}}}
 
@@
     \begin{itemize}
-      \item A data computation depends on a data value if the corresponding \evaldatacomputation function calls \hashlookup for the data value's hash code.
-      \item A data computation depends on another data computation if the corresponding \evaldatacomputation function calls \datacomputationlookup for the other data computation.
-      \item A semilattice value depends on a data value if the corresponding \checkelem function calls \hashlookup for the data value's hash code.
-      \item A semilattice value depends on a data computation if the corresponding \checkelem function calls \datacomputationlookup for the data computation.
-      \item A semilattice value depends on another semilattice value if the corresponding \checkelem function calls \latticelookup for the other semilattice value's semilattice.
-      \item A semilattice value depends on a semilattice computation if the corresponding \checkelem function calls \latticecomputationlookup for the semilattice computation.
-      \item A semilattice computation depends on a data value if the corresponding \evallatticecomputation function calls \hashlookup for the data value's hash code.
-      \item A semilattice computation depends on a data computation if the corresponding \evallatticecomputation function calls \datacomputationlookup for the data computation.
-      \item A semilattice computation depends on a semilattice value if the corresponding \evallatticecomputation function calls \latticelookup for the semilattice value's semilattice.
-      \item A semilattice computation depends on another semilattice computation if the corresponding \evallatticecomputation function calls \latticecomputationlookup for the other semilattice computation.
+      \item A data computation depends on a data value if the corresponding \evaldatacomputation{} function calls \hashlookup{} for the data value's hash code.
+      \item A data computation depends on another data computation if the corresponding \evaldatacomputation{} function calls \datacomputationlookup{} for the other data computation.
+      \item A semilattice value depends on a data value if the corresponding \checkelem{} function calls \hashlookup{} for the data value's hash code.
+      \item A semilattice value depends on a data computation if the corresponding \checkelem{} function calls \datacomputationlookup{} for the data computation.
+      \item A semilattice value depends on another semilattice value if the corresponding \checkelem{} function calls \latticelookup{} for the other semilattice value's semilattice.
+      \item A semilattice value depends on a semilattice computation if the corresponding \checkelem{} function calls \latticecomputationlookup{} for the semilattice computation.
+      \item A semilattice computation depends on a data value if the corresponding \evallatticecomputation{} function calls \hashlookup{} for the data value's hash code.
+      \item A semilattice computation depends on a data computation if the corresponding \evallatticecomputation{} function calls \datacomputationlookup{} for the data computation.
+      \item A semilattice computation depends on a semilattice value if the corresponding \evallatticecomputation{} function calls \latticelookup{} for the semilattice value's semilattice.
+      \item A semilattice computation depends on another semilattice computation if the corresponding \evallatticecomputation{} function calls \latticecomputationlookup{} for the other semilattice computation.
     \end{itemize}
@@
 This enables synchronization between different databases, since it is possible to check if any dependencies differ before joining the values.
-Synchronization proceeds recursively; within the \join call that merges two semilattice values, the \latticelookup and \latticecomputationlookup functions recursively compute a merged result from the two databases, using the Merkle trees as a frozen pointer to the relevant subset of each database at the time of the merge.
+Synchronization proceeds recursively; within the \join{} call that merges two semilattice values, the \latticelookup{} and \latticecomputationlookup{} functions recursively compute a merged result from the two databases, using the Merkle trees as a frozen pointer to the relevant subset of each database at the time of the merge.
 

### 2022-11-20 src/js/runtime_channel.rs
fix check_elem js

@@ -270,7 +270,7 @@ impl RuntimeState {
                                 self.register_call_function(query_id, "eval_computation", vec![key, JsValue::from(ctxid)], |res| Ok(LibraryResult::EvalComputation(res)));
                             },
                             LibraryQuery::CheckElem(key, value, ctxid) => {
-                                self.register_call_function(query_id, "hash_put", vec![key, value, JsValue::from(ctxid)], |_| Ok(LibraryResult::CheckElem));
+                                self.register_call_function(query_id, "check_elem", vec![key, value, JsValue::from(ctxid)], |_| Ok(LibraryResult::CheckElem));
                             },
                             LibraryQuery::Join(key, value1, value2, ctxid) => {
                                 self.register_call_function(query_id, "lattice_join", vec![key, value1, value2, JsValue::from(ctxid)], |res| Ok(LibraryResult::Join(res)));

### 2022-11-20 paper/hl_paper.tex
semilattice graphs

@@
 
+    \subsection{Semilattice graphs}
+
+      We can extend the notion of dependent semilattices to a graph of semilattices depending on each other.
+
+      As a pre-requisite, we define a semilattice over $\mathrm{Option}(L)$ given a semilattice over $L$ as follows:
+
+      \begin{itemize}
+        \item $\mathrm{None} \vee y = y$
+        \item $x \vee \mathrm{None} = x$
+        \item $\mathrm{Some}(x) \vee \mathrm{Some}(y) = \mathrm{Some}(x \vee y)$
+      \end{itemize}
+
+      A \emph{semilattice graph} consists of:
+      \begin{itemize}
+        \item a totally ordered set $K$ of keys, representing semilattices that may depend on previous semilattices
+        \item a set $V$ of values of any of these semilattices
+        \item a function $\checkelem : \prod_{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times V \to \mathrm{Bool}$ assining to each $K$ and corresponding semilattice value a boolean indicating whether the value is an element of the semilattice, which depends on only previous semilattice elements in a way continuous in the $K_{<k} \to \mathrm{Option}(V)$ argument under the product topology.
+        \item a function $\join : \prod_{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times V \times V \to V$ assigning to each $K$ and corresponding semilattice values a join, which depends on only previous semilattice elements in a way continuous in the $K_{<k} \to \mathrm{Option}(V)$ argument under the product topology.
+        \item a function $\transport : \prod_{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times (K_{<k} \to \mathrm{Option}(V)) \times V \to \mathrm{Option}(V)$ assigning to each $K$ and semilattice value a transported value, which depends on previous semilattice elements in both the previous and new contexts, in a way continuous in the $K_{<k} \to \mathrm{Option}(V)$ arguments under the product topology.
+      \end{itemize}
+
+      These must satisfy dependency conditions and dependent semilattice conditions.
+The dependency conditions state that there exists some function $d : \prod{k \in K} (K_{<k} \to \mathrm{Option}(V)) \times V \rightarrow 2^K$ such that:
+
+      \begin{itemize}
+        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V)), v : V$, the function $c' \mapsto \checkelem(k, c', v)$ equals $\checkelem(k, c, v)$ if $c'(k') = c(k')$ for all $k' \in d(k, c, v)$.
+        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V)), v_1, v_2 : V$, the function $c' \mapsto \join(k, c', v_1, v_2)$ equals $\join(k, c, v_1, v_2)$ if $c'(k') = c(k')$ for all $k' \in d(k, c, v_1) \cup d(k, c, v_2)$.
+        \item For $k : K, c_1 : (K_{<k} \to \mathrm{Option}(V)), c_2 : (K_{<k} \to \mathrm{Option}(V)), v : V$, the function $(c_1', c_2') \mapsto \transport(k, c_1', c_2', v)$ equals $\transport(k, c_1, c_2, v)$ if $c_1'(k') = c_1(k')$ and $c_2'(k') = c_2(k')$ for all $k' \in d(k, c_1, v) \cup d(k, c_2, v)$.
+      \end{itemize}
+
+      To define the dependent semilattice conditions, we first define a partial order over partial key-value mappings.
+If $k : K, v : V, c_1, c_2 : (K_{<k} \to \mathrm{Option}(V))$, if $\checkelem(k, c_1, v)$ and $\checkelem(k, c_2, v)$, then:
+
+      $$c_1 \leq_{k, v} c_2 \Leftrightarrow \forall k' \in d(k, c_1) \cup d(k, c_2), c_1(k') \leq c_2(k')$$.
+
+      The dependent semilattice conditions state that:
+
+      \begin{itemize}
+        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V))$, then $(v_1 : V, v_2 : V) \mapsto \join(k, c, v_1, v_2)$ is the join of a semilattice over $\{ v : V ~ | ~ \checkelem(k, c, v) \}$.
+        \item For $k : K, c_1, c_2 : (K_{<k} \to \mathrm{Option}(V)), v_1 : V$, if $\checkelem(k, c_1, v_1)$ and $c_1 \leq_{k, v_1} c_2$, then $\transport(k, c_1, c_2, v_1) = \mathrm{Some}(v_2)$, then $\checkelem(k, c_2, v_2)$.
+        \item For $k : K, c : (K_{<k} \to \mathrm{Option}(V)), v : V$, if $\checkelem(k, c, v)$, then $\transport(k, c, c, v) = \mathrm{Some}(v)$.
+        \item For $k : K, c_1, c_2, c_3 : (K_{<k} \to \mathrm{Option}(V)), v_1 : V$, if $\checkelem(k, c_1, v_1)$ and $c_1 \leq_{k, v_1} c_2$ and $\transport(k, c_1, c_2, v_1) = \mathrm{Some}(v_2)$ and $c_2 \leq_{k, v_2} c_3$, then $\transport(k, c_1, c_3, v_1) = \transport(k, c_2, c_3, v_2)$.
+        \item For $k : K, c_1, c_2 : (K_{<k} \to \mathrm{Option}(V)), v_1, v_1' : V$, if $\checkelem(k, c_1, v_1)$ and $\checkelem(k, c_1, v_1')$, then $\transport(k, c_1, c_2, \join(k, c_1, v_1, v_1')) = \transport(k, c_1, v_1) \vee_{k, c_2} \transport(k, c_1, v_1')$, where $\vee_{k, c_2}$ is the join of the option semilattice implied by the join $(v, v') \mapsto \join(k, c_2, v, v')$.
+      \end{itemize}
+
     \subsection{Composing semilattice fibrations}

### 2022-11-22 paper/hl_paper.tex
graph agda
refactor partial order!

@@
 
-      $$c_1 \leq_{k, v} c_2 \Leftrightarrow \forall k' \in d(k, c_1) \cup d(k, c_2), c_1(k') \leq c_2(k')$$.
+      $$c_1 \leq_{k, v} c_2 \Leftrightarrow \forall k' \in d(k, c_1) \cup d(k, c_2), c_1(k') \leq c_2(k').$$
 
@@
 
-    \subsection{Composing semilattice fibrations}
+      A context $c : (K_{<k} \to \mathrm{Option}(V))$ is called \emph{finite} if it is $\mathrm{None}$ for all but a finite set of keys.
+For $k : K$ we define a join over finite contexts (not yet proven to be a semilattice join):
 
-      Semilattice fibrations can be composed iteratively. Composing finitely is straightforward, while omposing infinitely requires more work.
+      $$ (c_1 \vee_k c_2)(k') := c_1(k') \vee_{k', c_1 \vee c_2} c_2(k').$$
 
-      Suppose we have:
-
-      \begin{align*}
-        L_0 &: \mathrm{Semilattice} & \\
-        (l_n, t_n) &: \mathrm{SemiLatFib}(L_{n-1}) & \text{ for natural $n \geq 1$ } \\
-        L_n &:= \mathrm{SemiLatTot}(L_{n-1}, l_n, t_n)    & \text{ for natural $n \geq 1$ }
-      \end{align*}
-
-      We now wish to define $L_\omega$ to be a semilattice that is, morally speaking, the limit of $L_n$.
-This lattice's elements are infinite lists $x_1, x_2, \ldots$ satisfying $x_n : l_n(x_{<n})$ for each natural $n \geq 1$, where $x_{<n} = x_1, \ldots, x_{n-1} : L_{n-1}$.
-The joins are defined as follows:
-
-      $$(\overline{x} \vee \overline{y})_n := t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)$$
-
-      \begin{theorem}
-        $L_{\omega}$ is a semilattice.
-      \end{theorem}
-      \begin{proof}
-
-        To verify that $L_\omega$ is a semilattice, we must verify three properties:
-        \begin{itemize}
-          \item \emph{Commutativity}:
-            \begin{align*}
-              (\overline{x} \vee \overline{y})_n
-              &= t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})} t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n) \\
-              &= t_n(y_{<n}, y_{<n} \vee x_{<n})(y_n) \vee_{l_n(y_{<n} \vee x_{<n})} t_n(x_{<n}, y_{<n} \vee x_{<n})(x_n) \\
-              &= (\overline{y} \vee \overline{x})_n
-            \end{align*}
-          \item \emph{Associativity}:
-            \begin{align*}
-              & ((\overline{x} \vee \overline{y}) \vee \overline{z})_n \\
-              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{x} \vee \overline{y})_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})}
-                 t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
-              =& t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
-                          t_n(x_{<n}, x_{<n} \vee y_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n})}
-                          t_n(y_{<n}, x_{<n} \vee y_{<n})(y_n)) \\
-               &\vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
-              =& (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(x_{<n}, x_{<n} \vee y_{<n}))(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & (t_n(x_{<n} \vee y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, x_{<n} \vee y_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(z_n)  \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(y_{<n}, y_{<n} \vee z_{<n}))(y_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & (t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n}) \circ t_n(z_{<n}, y_{<n} \vee z_{<n}))(z_n)  \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(
-                          t_n(y_{<n}, y_{<n} \vee z_{<n})(y_n) \vee_{l_n(y_{<n} \vee z_{<n})}
-                          t_n(z_{<n}, y_{<n} \vee z_{<n})(z_n)) \\
-              =& t_n(x_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})(x_n) \vee_{l_n(x_{<n} \vee y_{<n} \vee z_{<n})} \\
-               & t_n(y_{<n} \vee z_{<n}, x_{<n} \vee y_{<n} \vee z_{<n})((\overline{y} \vee \overline{z})_n) \\
-              =& (\overline{x} \vee (\overline{y} \vee \overline{z}))_n
-            \end{align*}
-          \item \emph{Idempotence}:
-            \begin{align*}
-              (\overline{x} \vee \overline{x})_n &= t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \vee_{l_n(x_{<n} \vee x_{<n}))} t_n(x_{<n}, x_{<n} \vee x_{<n})(x_n) \\
-              &= t_n(x_{<n}, x_{<n})(x_n) \\
-              &= x_n
-            \end{align*}
-        \end{itemize}
-      \end{proof}
-
-      We can further stack to higher ordinals than $\omega$. Let $n_u \geq 1$ be the number of universes. Assume we have:
-
-      \begin{align*}
-        L_{0, 0} &: \mathrm{Semilattice} & \\
-        (l_{u, n}, t_{u, n}) &: \mathrm{SemiLatFib}(L_{u, n-1}) & \text{for natural $0 \leq u < n_u$, $n \geq 1$ } \\
-        L_{u, n} &:= \mathrm{SemiLatTot}(L_{u, n-1}, l_{u, n}, t_{u, n})    & \text{ for natural $0 \leq u < n_u$, $n \geq 1$ } \\
-        L_{u, 0} &:= L_{u-1, \omega} & \text{ for natural $1 \leq u < n_u$ }
-      \end{align*}
-
-      where $L_{u, \omega}$ is constructed the same way as $L_{\omega}$ above, as the limit of $L_{u, n}$ as $n$ approaches $\infty$.
-Inductively, each $L_{u, n}$ is a semilattice for naturals $0 \leq u < n_u$ and $n \geq 0$; this lattice's ordinal corresponds to $u \omega + n$.
-While in theory it may be possible to reach higher ordinals $\omega^2$ and beyond, we consider this beyond the scope of this work.
+      The recursion is well-founded because of the ordering on keys.
 
